### 2023-06-12 Temp_paper.Rnw
Temp

\documentclass{article}
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[a4paper,top=3cm,bottom=3cm,left=4cm,right=4cm,marginparwidth=1cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{hyphenat}
\usepackage{url}
\usepackage{enumerate}
\usepackage{natbib}
\bibliographystyle{plain}
\usepackage{hyphenat}
\title{
% Fitting Mechanistic Modeling on Panel Data with PanelPOMP Framework\\
Analysis of Panel Data via Mechanistic Model in PanelPOMP Framework\\
\large A Thesis Submitted in Partial Fulfillment of Bachelor of Science in (Honor) Data Science at the University of Michigan
}
\author{BO YANG\\
\large Thesis Advisor\\
\large Professor Edward Ionides\\
\large Professor Aaron King\\
\large Department of Statistics, University of Michigan}
\date{Apr 2023}
\linespread{1.25}
\usepackage{Sweave}
\begin{document}

\maketitle

<<Setup, include=FALSE,echo=FALSE,results='hide'>>=
library(ggplot2)
library(plyr)
library(reshape2)
library(magrittr)
library(foreach)
library(readxl)
library(doParallel)
registerDoParallel(cores=36)
library(pomp)
library(panelPomp)
stopifnot(packageVersion("pomp") >= "1.4.7")

set.seed(0001)

lum_data = read_excel("~/Desktop/Honor_thesis/Report_Rnw_code/Mesocosmdata.xlsm")
lumNoPara <- lum_data[1:100, ] 
lumNoPara <- subset(lumNoPara, select = c(rep, day, lum.adult))
lumNoPara <- lumNoPara[100: 1, ]

lumNoPara$day = (lumNoPara$day - 1) * 5 + 7
data_lum = list()
trails = c("A","B","C","D","E","F","G","H","I","J")
for (i in 1: length(trails)){
  data_lum[[i]]=subset(lumNoPara, select = c("day", "lum.adult"),
                   lumNoPara$rep == trails[i])
}

Mesocosm_data = read_excel("~/Desktop/Honor_thesis/Report_Rnw_code/Mesocosmdata.xls")
dentNoPara <- Mesocosm_data[1:100, ]
dentNoPara <- subset(dentNoPara, select = c(rep, day, dent.adult))
dentNoPara <- dentNoPara[100: 1, ]
dentNoPara$day = (dentNoPara$day - 1) * 5 + 7
data_dent = list()
trails = c("A","B","C","D","E","F","G","H","I","J")
for (i in 1: length(trails)){
  data_dent[[i]]=subset(dentNoPara, select = c("day", "dent.adult"),
                   dentNoPara$rep == trails[i])
}
@
\begin{abstract}
\linespread{1}

In ecological research, panel data is a typical format  of data to record various statistics. The panel data comprises multi-magnitude time series data \cite{8}. As a result of its high dimension, panel data analysis presents significant difficulties. In the meantime, mechanistic modeling is an attractive yet difficult issue in ecological research. Mechanistic models enable quantitative analysis, which provides a deeper understanding of the latent process of the dynamic system, and provide inference and prediction based on panel data. However, it is limited by the non-Gaussian nature of the vast majority of ecological data. The objective of this study is to showcase the practicality and advancement of mechanistic model development for ecological panel datasets using the panelPOMP structure and the panel iterated filtering optimization approach (PIF). PIF in the research process is used as a probability maximization and parameter estimation technique. We choose two classical models: logistic growth model and Lotka-Volterra model as an example to interpret the validity of mechanistic models. After fitting two models mentioned above on the Daphnia panel data from the research group of Searle et al. \cite{6}, which includes 10 independent time series with 10 data points in each unit, we apply PIF to get the parameter swarms that result in the lowest Akaike information criterion (AIC) in different models. After comparing the AIC of each model, the result demonstrates that the Lotka-Volterra model has the lowest AIC, indicating that alga, which is the Daphnia's prey, has a substantial impact on the population dynamics of each species of Daphnia. Throughout the panel data examination, we illustrate the ability of a panelPOMP framework to tackle the complex nature of panel data analysis, emphasizing its capacity to provide meaningful explanations and in-depth understanding of latent processes not captured by existing data. Nevertheless, discussion on the limitations and interpretations of mechanistic models in dealing with ecological data will be included.

\end{abstract}

\newpage

\linespread{1.25}
\section{Introduction}

Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary. As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \cite{7}. Utilizing the established process, they elucidate the actions of hidden mechanisms \cite{2}. The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems. This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters. Mechanistic modeling has been widely employed in social sciences, industrial engineering, and physics for an extended period. Improved computational capabilities now make it more effective. Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information. Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task. As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \cite{2}.

Panel data, sometimes referred to as longitudinal data, is composed of multiple time series. Each of these time series, which may be multivariate, contains a series of observations collected from a separate entity \cite{1}. In ecology, panel data is a traditional data form that researchers gather to mitigate the impact of random variations throughout experiments, so that it can more accurately measure changes in experimental objectives and control the variables needed to achieve experimental goals. Due to the non-linear characteristics of ecological data, the analysis of panel data entails a high-dimensional structure. Such features diminish the explanatory power of traditional time series analysis and certain Monte Carlo inference techniques. The goal of this study is to investigate the practicality of employing mechanistic models on ecological panel data using the panelPOMP structure.

The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component. This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering. Stemming from iterated filtering \cite{4} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood. This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.

In this research, we will show the advancement of the panelPOMP model by analyzing the Daphnia panel data generated in the experiment of Searle et al. \cite{6}. The research group of Searle et al. collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \cite{6}. In their lab experiment, they generated standard ecological panel data featuring various independent trials under the same experimental conditions, with each consisting of a brief, nonlinear, and nonstationary time series. Conventional mechanistic theories cannot directly interpret these observations. Utilizing the panelPOMP approach allows for the combination of theoretical models and empirical data, enabling the extraction of the comprehensive information found in each individual time series.

For a more in-depth examination of the concealed processes in the panel data, this research incorporates two classic ecological models - the Logistic Growth model and the Lotka-Volterra model - within the panelPOMP framework. After applying both models to the Daphnia panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach. These findings are subsequently compared to discover a more accurate and appropriate model for the given data. The superior performance of the Lotka-Volterra model implies that the population density of alga \textit{Ankistrodesmus falcatus} \cite{6}, which is Daphnia's prey, impact the dynamics of both species of Daphnia. It is worth noting that the population density of alga is neither observable nor recorded in the experiment. That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes.

Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the Daphnia panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system. By analyzing Daphnia panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data. Additionally, the later section will further discuss how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models.

\section{Data description}
\subsection{Daphnia-alga Data}

Searle and colleagues conducted two experiments to investigate the interaction between native and invasive Daphnia species and a parasitic fungus called \textit{Metschnikowia bicuspidate} \cite{6}. The first experiment was an individual-level study that compared the host competence of the two Daphnia species. The second experiment was a mesocosm experiment that examined the community-level species interaction and the role of population density in disease patterns. The mesocosm experiment involved six different treatments, including both single-species and mixed-species treatments, with and without parasites. The experiment was conducted indoors, with 45 Daphnia added to each 15L COMBO media at the start of the four single-species treatments and 10 invasive and 35 host species added in the treatments of mixed-species. The same treatment will be repeated for ten times, and samples were taken every five days for 52 days after the experiment began. The experiment also refills food by adding $2.0 \times 10^7$ cells of alga \textit{falcatus} twice a week (adding $1.3 \times 10^6$ cells $L^{-1}$). And on the timing of sampling, 1 liter COMBO solution was chosen as a sample for each of the 10 units. The genus, disease status, generation stage, and sex of each individual were documented, and the temperature in the lab was kept constant at 23.27℃ (SE $\mp$ 0.2) with a 16L:8D photoperiod and all the other conditions remained the same for all units during the experiment time.

In the analysis of this dataset, our attention is directed towards the interplay between the Daphnia species (Daphnia \textit{dentifera} and Daphnia \textit{lumholtzi}) and their surroundings in the absence of parasites. Specifically, we analyze a panel data consisting of ten independent time series representing the population density of \textit{dentifera} adults. These observations were gathered simultaneously at ten time points, as shown in Figure 1 and 2, and offer the opportunity to evaluate mechanistic models and evaluate the relationship between alga and Daphnia dynamics.\\

\begin{figure}[h]
  % \includegraphics[width=\linewidth]{data_pic.png}  
<<Dent_density, echo=FALSE, fig.height=5>>=  
ggplot(dentNoPara,aes(x=day,y=dentadult))->ggp
ggp + 
  geom_line(data=dentNoPara,aes(x=day,y=dent.adult)) + 
  facet_wrap(~rep,scales="free")+labs(y = "Daphnia dentifera density")
@

  \caption{Data visulization for Daphnia dentifera: Ten replications of Daphnia population density panel}
  \label{figure 1}
\end{figure}

\begin{figure}
  % \includegraphics[width=\linewidth]{lum.png}
  <<Lum_density2, echo=FALSE, fig.height=5>>=
  ggplot(lumNoPara,aes(x=day,y=lumadult))->ggp
ggp + 
  geom_line(data=lumNoPara,aes(x=day,y=lum.adult)) + 
  facet_wrap(~rep,scales="free")+labs(y = "Daphnia lumholtzi density")
@  
  \caption{Data visulization for Daphnia dentifera: Ten replications of Daphnia population density panel}
  \label{figure 2}
\end{figure}

\newpage

\section{Methodology}

\subsection{Partially Observed Markov Processes}

Frequently, we lack a comprehensive understanding of the underlying mechanisms that drive the evolution of a natural system. For example, the precise number of individuals who are exposed to a disease or the density of alga at any given time is typically unknown. To address these issues of missing information, we can utilize partially observed Markov (POMP) models. These models combine both latent and observable processes which are reliant on each other.

With the random variables $Y_{0:N}$ which are the observable states process, the measurement model is $f_{Y_n|X_n}(y_n|x_n)$, and the collected data $y_{0:N}$ are observations of this process. We suppose random variables $X_{0:N}$ represent the latent states, where $X_0$ represents the initial state of the latent process, $f_{X|X_{n-1}}(x_n|x_{n-1})$. We assume that each $Y_n$ depends only upon the latent process at time $n$, $X_n$, and is conditionally independent of the other variables representing the measurement and latent processes, $Y_{n-1}$ and $X_{n-1}$, where $1 \leq n$. Together, $X_{0:N}$ and $Y_{0:N}$ form our POMP model.

\subsection{Likelihood of models and PIF}

In statistical inference challenges, likelihood frequently serves as the basis for estimating parameters and selecting models. Considering a model parameterized by vector $\theta$ within the $m$-dimensional parameter space $\Theta_m$, the likelihood function represents the combined probability density of the dataset, $y_{1:N}$, at $\theta$:
        \begin{equation}
        \mathcal{L}(\theta) = f_{Y_{1:N}}(y_{1:N}; \theta)
        \end{equation}
The goal is to get the estimation of $\theta$, $\hat{\theta}$ that maximizes the likelihood function $\mathcal{L}(\hat{\theta})$, or log-likelihood function $\mathbf{\ell}(\hat{\theta})$.

An ecological model's effectiveness in tracking disease spread hinges on its capability to predict cases or incidence rates. This capability is inherently tied to our trust in the model's precision in forecasting and our grasp of the temporal changes in latent states. As a result, we encounter two interrelated issues: ascertaining the distribution of $X_{n}$ at time $n$ based on $y_{1:n}$, and pinpointing parameter values, $\hat{\theta}$, that maximize the data likelihood. These processes are commonly respectively concluded as the filtering process and the inference process.

Working with POMP models can prove difficult due to the complexity of formulating both the likelihood function and transition density in intricate settings. To address this issue, several techniques have been devised, including the particle filter. Utilizing the particle filter method, it is only necessary to supply data, initial density simulators, single time-step transition density simulators for the latent process, and an evaluator for the observation process density reliant on the hidden process in order to achieve maximum likelihood estimates.

To implement the particle filter, we start by initializing $M$ particles at time 1, where $\{ X_1^m; m \in 1:M \}$. All with the required latent state detail and a vector of parameter values, $\theta$. As a result, for every time step n, where $n \in $ [$N$], we advance the each particle by doing the one-step random walk based on transition density, resulting in a set of particles that represents the prediction distribution at time n, which can be represented as $f_{X_n|X_{n-1}}(\cdot | X_{n-1}^m; \theta)$. Next, we assign weights to the particles based on the measurement density to obtain an ensemble that represents the filtering distribution at time n, where the new weight is calculated via $w_{n,m} = f_{Y_n|X_n}(y_n | x_n^m)$. Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \cite{5}. The particle filter is particularly useful in situations where both the likelihood function and transition density are difficult to calculate analytically.

As we know, the measurement process variables are mutually independent. In the mean time, the latent process states are dependent to one other. We also notice that the weight we assign to the particles will vary among different particles, which makes the calculation hard. As M to be large enough, we may approximate the weight of each particle to be the mean value of the to calculate the log-likelihood. Based on such fact, we can summarize as following:
         \begin{flalign}
          \mathcal{L}(\theta) &= f_{Y_{1:n}}(y_{1:N}; \theta) \\
         & = \prod_{n=1}^N f_{Y_{n} | Y_{1:n-1}}(y_n | y_{n-1}; \theta) \\
         & = \prod_{n=1}^N f_{Y_n | X_n}(y_n | x_n ; \theta)\\
         & \approx \prod_{n=1}^N \frac{1}{M} \sum_{m = 1}^M f_{Y_n|X_n}(y_n | x_n^m) 
         \end{flalign}

The particle filter technique offers a straightforward approach for estimating the likelihood of data in relation to a model and approximating the latent process variable's distribution at a specific time based on the available observations \cite{3}. The Monte Carlo technique enables us to approximate the conditional likelihood using the weight assigned to each particle. This is a useful tool for analyzing complex models with partially observed Markov processes, as it simplifies the estimation process for the distribution of latent variables and the likelihood of the data.

\subsection{PanelPomp Framework}
In a panel data set containing K independent time series, K entities are represented by the symbols $1,2,...,K$, with N periodic observations collected for each unit $(t_1,t_2,...,t_{N})$. A stochastic observable process, denoted as $y_{k,1},...,y_{k,N}$, generates these measurements. PanelPOMP is an adaptation of the partially observable Markov process (POMP) tailored for panel data analysis \cite{1}. It entails developing a POMP model for each entity and employing the Panel Iterated Filtering (PIF) method to achieve the highest log-likelihood. Based on the POMP framework, a process and a measurement model were devised for every time series unit \cite{4}. The measurement model elucidates how the potential real population density generates observable time series data with the measurement density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$, while the process model portrays the actual population's dynamics through the one-step transition density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$. Ultimately, the joint density equals the initial density multiplied by the product of the process and measurement models for each state. This study will apply and examine three ecological models for a single-species experimental container without parasites: a logistic model, a Lotka-Volterra model with sufficient predator food supply, and a Lotka-Volterra model incorporating food supply management.

\subsection{Benchmarks}
We will choose two different models as benchmarks for this research. The three benchmarks will be: logistic model by Searles et al. \cite{6}, Lotka-Volterra model with enough food supply for predator.
\subsubsection{Logistic Model}
First, we can check the logistic model be Searles et al., 
    \begin{flalign}
    & dS=\left(r + \zeta \right)S\left(1-\frac{S}{K}\right) \cdot dt-\delta S \cdot dt\\
    &\zeta  \sim N\left(0, \sigma_{G}\right)
    \end{flalign}
    This model contains one state variable Daphnia population density(S) and 2 parameters: $r, K$ and $\sigma_{G}$. Among those parameters, $r$ represents the growth rate of Daphnia. $K$ is the carrying capacity of this ecosystem in terms of density for certain species of Daphnia.$\delta$ denotes the sampling rate in the experiment. In order to minimize the dimension of parameters, we fixed the sample rate $\delta$ as a constant equal 0.013 as in Searles et al. \cite{6}. $\zeta$ is a noise term added to population growth to allows for random variation in the dynamic. In this case, we assure $\zeta$ follows a Normal distribution with 0 mean and a standard deviation of $\sigma_{G}$.

The model encompasses a single state variable, Daphnia population density (S), and three parameters: $r, K$, and $\sigma_{G}$. The parameter $r$ signifies the Daphnia growth rate, while $K$ denotes the ecosystem's carrying capacity in terms of density for specific Daphnia species. The sampling rate in the experiment is represented by $\delta$. To reduce parameter dimensionality, we set the sample rate $\delta$ as a constant equal to 0.013, following Searles et al. \cite{6}. The noise term $\zeta$ is incorporated into the population growth to account for random fluctuations in the dynamic. In this instance, we ensure that $\zeta$ adheres to a normal distribution with the mean of 0 and the standard deviation to be $\sigma_{G}$.

The logistic model is a well-known ecological model that is commonly used to describe the population growth of a species under limited conditions, such as space or food availability. Given that the population growth curve appears to follow a similar pattern to the logistic model, we attempt to fit the data to this model. As the population density (S) approaches the carrying capacity (K), the rate of population growth decreases and becomes zero when S equals K. If S exceeds K, the population will decline. Moreover, as S increases, the noise in the data also increases, which will be accounted for by the model.

Assuming that the logistic model fits the data accurately, it would imply that the population has sufficient food and that the limiting factor is the space available for Daphnia. However, if the model does not fit the data well, then we should consider the possibility of alga, which serves as the food source for Daphnia, being a significant factor affecting population dynamics.

\subsubsection{Lotka-Volterra model with enough food supply}

Second, we will check the Lotka-Volterra model with enough food supply \cite{9}: 
\begin{flalign}
    & dF =\left(\alpha + \zeta_F \right) F\left(1-\frac{F}{k_f}\right) \cdot dt-\beta F S \cdot dt  \\
    & dS= \left(\theta+ \zeta_S \right) S F \cdot dt-\gamma S \cdot dt- \delta S \cdot dt \\
     &\zeta_F  \sim N\left(0, \sigma_F\right) \\
     &\zeta_S  \sim N\left(0, \sigma_S \right) 
    \end{flalign}
    This model includes 8 parameters: $\gamma,\alpha,\sigma_F,\sigma_S,\theta,\beta,\delta$ and $k_f$. Among those parameters, $\alpha$ represents the rate of growth of algae. $k_f$ shows the carrying capacity of algae in each bucket. $\beta$ is the consumption rate of algae by certain species of Daphnia, $\theta$ shows the rate of growth of Daphnia with certain type of food and $\gamma$ is the rate of death of certain species of Daphnia. Same as the logistic model, we treat $\delta$ as a constant equal to 0.013 according to the experimental settings as in Searles et al. \cite{6}. $\zeta_F$ and $\zeta_S$ are noise terms of algae (F) growth and Daphnia (S) growth in order to show randomness of the dynamic system. They are generated from a normal distribution with 0 mean, and $\sigma_F, \sigma_S$ to be the variance respectively. It is worth noting that whenever the density of algea equals zero, we will reset the density of alga back to 1000 as a way to ensure the adequacy of food.

Using the Lotka-Volterra model with enough food supply, we can investigate how the population dynamics of Daphnia are influenced by the availability of sufficient algae. If the data are adequately explained by this alternative hypothesis, it can be concluded that the density of algae plays a significant role in Daphnia population dynamics rather than just being a simple environmental factor, as other conditions may not affect this treatment. On the other hand, if the model does not fit the data well, it's possible that the algae supply is not always sufficient. Therefore, it may be necessary to test the Lotka-Volterra model with food supply management, which assumes a limited supply of algae.

\subsection{Lotka-Volterra model with food supply management}

In the model, We denote the density of alga as F ($10^5$ cells per liter), and density of Daphnia as S (individuals per liter), and total amount of Daphnia that is sampled out from the bucket as M(individuals per liter). The model can be expressed as below:
          \begin{flalign}
          & dF =\alpha F\left(1-\frac{F}{k_f}\right)\cdot dt+F \cdot \zeta_F \cdot dt-\beta F S \cdot dt- \delta F \cdot dt+ \mu \cdot dt \\
          & dS = \theta S F \cdot dt+S \cdot \mathrm{\zeta_S} -\gamma S \cdot dt - \delta S \cdot dt\\
          &\zeta_F \sim N \left( 0, \sigma_F \right) \\
          &\zeta_S \sim N \left( 0, \sigma_S  \right) \\
          &d M = \delta S \cdot dt
         \end{flalign}

\subsubsection{Model interpretation}
In this model, we set the initial condition to be $M = 0, S = 3$ and $F = 16.667$ by Searles et al. Also, we can separate the differential equations in to several parts:
\paragraph{Alga}

The change of the density of alga can be viewed as
        \begin{flalign}
        & dF_{birth} = \alpha F\left(1-\frac{F}{k_f}\right) dt+F \cdot \zeta_F dt\\
        &dF_{cons} = \beta F S dt\\
        &dF_{spl} = \delta F dt\\
        &dF_{refill} = \mu dt\\
        &dF = dF_{birth} - dF_{cons} - dF_{spl} + dF_{refill} 
        \end{flalign} where $\alpha$ can be viewed as growth rate, $k_f$ can be viewed as the carrying capacity, $\zeta_F$ is the noise term, $\beta$ can be viewed as the consumption rate, which are all unknown parameters. At the same time, $\delta$ and $\mu$ are fixed parameters with $\delta = 0.013$ and $\mu = 0.37$, which implies the food supply rate. We split the change of alga density (dF) into four parts. The equation (17) represents the change of population due to the birth of alga and noise term. While the equation (18) is shows the amount of alga that is consumed by Daphnia, which will make the density decrease. The equation (19) and (20) indicate that change of alga density due to the process of alga refill and sampling. Such four equations explain the processes and causes of the changes in alga population density.

\paragraph{Daphnia}

The change of the density of Daphnia can be viewed as
        \begin{flalign}
        &dS_{grow} = \theta S F dt+S \cdot \zeta_Sdt\\
        &dS_{death} = \gamma Sdt\\
        &dS_{spl} = \delta Sdt\\
        &dS =  S_{grow} - S_{death} - S_{spl}
        \end{flalign}where $\theta$ can be viewed as growth rate, $\zeta_S$ is the noise term, $\gamma$ can be viewed as the death rate, which are all unknown parameters. At the same time, $\delta$ is fixed parameters with $\delta = 0.013$. We split the change of Daphnia density (dS) into three parts. The equation (22) represents the change of population due to the birth of Daphnia and noise term. While the equation (23) is shows the amount of alga that is consumed by Daphnia, which will make the density decrease. The equation (24) indicates that change of Daphnia density due to the process of  sampling. Such three equations explain the processes and causes of the changes in Daphnia population density.

    This model includes 9 parameters: $\gamma,\alpha,\sigma_F,\sigma_S,\theta,\beta,\delta, \mu$ and $k_f$. Among those parameters, $\alpha$ represents the rate of growth of algae. $k_f$ shows the carrying capacity of algae in each bucket. $\beta$ is the consumption rate of algae by certain species of Daphnia, $\theta$ shows the rate of growth of Daphnia with certain type of food and $\gamma$ is the rate of death of certain species of Daphnia. Similar to previous Lotka-Volterra model, we treat $\delta$ as a constant equal to 0.013 and $\mu = 0.37$ as a constant that implies the food refilling rate according to the experimental settings as in Searles et al. \cite{6}. $\zeta_F$ and $\zeta_S$ are noise terms of algae (F) growth and Daphnia (S) growth terms in order to show randomness of the dynamic system. They are induced by a normal distribution with 0 mean,and $\sigma_F, \sigma_S$ to be variance respectively. Unlike the previous model, we restricted the food supply. Since we observed a precipitous decline in Daphnia population density in the panel data, we suggest that this may be a dynamic change due to food insufficiency. Based on the experimental manipulations, we treated the food supply rate as a constant $\mu = 0.37 \cdot 10^5$ cells per day.

Using this model, we can investigate how the population dynamics of Daphnia are influenced by the density algae with condition that food may not be sufficient. If the data are adequately explained by this alternative hypothesis, it can be concluded that the density of algae plays a significant role in Daphnia population dynamics and the lack of food is potentially the reason of the precipitous decline in Daphnia population, as other factors may not obviously affect this treatment. On the other hand, if the model does not fit the data well, it's possible that the alga may not be the main reason that lead to the change of population of Daphnia, which means we need to consider other possible factors that can affect the population density of Daphnia.

\subsection{Measurement and Initial condition}

Assessment framework illustrates the technique through which noticeable information is derived from potential real populace concentration. It encompasses obtaining samples from the Daphnia population distribution in this case. As Searles and colleagues \cite{6} describe, on every sampling occasion, 1L of specimen is extracted from a 15L mixture post agitation, and the sampling rate $\delta$ is set to 0.013. Due to the constraint that both Daphnia and algae densities must be non-negative values, the procedure was approached as binomial distributions with the likelihood of success (p) equivalent to the sampling rate ($\delta$) and the totalan count (n) as $S/\delta$ for both logistic growth models and Lotka-Volterra models. For logistic growth models, the assessment framework can be summarized as:
        \begin{flalign}
        % &F \sim \operatorname{Poisson}(\lambda = S \times \delta)& \\
        % &\text { Or } &\\
        &F \sim \operatorname{binomial}(n=S / \delta, p=\delta), \delta=0.013
        \end{flalign}

S represents the actual population concentration, and the near integer (1/$\delta$ * S) signifies the estimated genuine count of adult Daphnia in the 15L medium. F symbolizes the observable specimen resulting from the hidden population density. Given that algae lack noticeable samples from the information, within the Lotka-Volterra pomp framework, we exclusively devised two different measurement models for the Daphnia population. One is to follow the measurement model of logistic growth model since alga is not observable, which potentially we can follow the simpler model. On the other hand, we can view the process of sampling as a Poisson process. Where the rate $\lambda = S \cdot \delta$ represents the average number of Daphnia per liter per sampling time. Hence, two possible assessment frameworks are:
        \begin{flalign}
        &F \sim \operatorname{binomial}(n=S / \delta, p=\delta), \delta=0.013 \\
        &F \sim \operatorname{Poisson}(\lambda = S \times \delta) 
        \end{flalign}

\section{Results}

Permitting parameters to be either shared or unique leads to varying dimensions for each model. With 10 units in the panel, adding a shared parameter increases the model's dimension by 1, and a specific parameter will add dimensions by 10. We encountered a balance between fitting accuracy and simplicity when selecting models. As a result, we employed the Akaike information criterion (AIC) as a benchmark to compare models while adjusting for dimension. We get the estimation of parameters via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models. Tables 5 reveal that the Lotka-Volterra model with parameter $\sigma_S$ to be unit specific and all other parameters to be unit shared offers the best AIC of 970.62 for Daphnia dentifera The lowest AIC of all logistic growth models is 2543.04, which we select k to be unit specific only. While Lotka-Volterra models with enough food supply has the minimum AIC of 2263.51. It is safe to conclude that the new Lotka-Volterra model outperforms both benchmark models in fitting Daphnia panel data, with an AIC advantage of 1292.89 units over the best benchmark AIC. This finding implies that the hidden alga process influences Daphnia population dynamics. Daphnia population density fluctuations primarily stem from the interplay between algae and Daphnia. Additionally, significant noise is generated in the Daphnia population throughout the process within these containers.

The log-likelihood iteration process is placed in the appendix. We can observe that among the 250 parts, the vast majority of them are approaching a certain upper bound, and among these log-likelihoods, the maximum value will be a reliable one, since it's close enough to the theoretical maximum log-likelihood. Therefore we believe that we have found a suitable enough set of parameters to describe this dynamic process. What's more, we have the mean value of standard error of the estimated log-likelihoods is 0.114, with over 90 percent having standard error less than 1, which is small enough to conclude that we get a reasonable maximum log-likelihood with the estimated parameters swarm.

Parameter estimates and their units are shown in table 5, 6, and 7in the appendix. $\sigma_S$ vary across the panel while other parameters are constant across units. $\sigma_S$ is the main constraints of Daphnia population dynamics, where  $\sigma_S$ is the noise generated generated in the population of Daphnia. The noise can potentially be generated due to the sampling process. Their variation in different units is probably caused by the different density of algae’s food: higher $\sigma_S$ indicates one unit has potential noise during the process. Level of noise is difficult to control in practice. By affecting the sampling process of Daphnia, they indirectly change the data we used to interpret the performance of Daphnia, resulting in different dynamics across the panel.

\subsection{Simluations}

For the model and parameters swarm that result in the best AIC, we did 20 simulations on each unit of the panel data (Figure 3 and 4). Viewing the figure, almost all simulations capture the data well. And the estimation of parameters swarm used for this simulation is shown below in table 1 and 2.

<<dent_mod, include=FALSE,echo=FALSE,results='hide'>>=
set.seed(0001)

Mesocosm_data = read_excel("~/Desktop/Final model/Mesocosmdata.xls")
dentNoPara <- Mesocosm_data[1:100, ]
dentNoPara <- subset(dentNoPara, select = c(rep, day, dent.adult))
dentNoPara <- dentNoPara[100: 1, ]
#Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
dentNoPara$day = (dentNoPara$day - 1) * 5 + 7
data = list()
trails = c("A","B","C","D","E","F","G","H","I","J")
for (i in 1: length(trails)){
  data[[i]]=subset(dentNoPara, select = c("day", "dent.adult"),
                   dentNoPara$rep == trails[i])
}

dent_rpro <- Csnippet("
                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
                      double noiB, noiG ;
                      double delta = 0.013; //fraction of volume replaced day-1

                      noiB = rnorm(0, sigB * sqrt(dt));
                      noiG = rnorm(0, sigG * sqrt(dt));

                      //------------F-------------

                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
                      Fpred =  S * F * (Beta * dt);
                      Fspl = F * (delta * dt);
                      Ffill = 0.37 * dt;

                      //------------S--------------
                      Sgrow = S * F * (Theta * dt) + S * noiG;
                      Sdeath = S * (gamma * dt);
                      Sspl = S * (delta * dt);

                      //M is the total Daphnia sample out
                      M += Sspl;
                      F += Fbirth - Fpred + Ffill - Fspl;
                      S += Sgrow - Sdeath  - Sspl;

                      if (S <= 0.0) {
                        S = 0.0;
                        error_count += 1;
                      }
                      if (F <= 0) {
                        F = 0;
                        error_count += 1000;
                      }
                      ")

# Initial state. Assume t0 = day 4
dent_init = Csnippet("
                     S = 3; //3= 45/15L
                     F = 16.667;
                     M = 0;
                     error_count = 0;
                     ")

dmeas = Csnippet("
                 double delta = 0.013;
                 double tol = 1e-18;

                 if (error_count > 0.0) {
                   lik = log(tol);
                 } else {
                   lik = dpois(dentadult,M,give_log);
                 }
                 ")

rmeas = Csnippet("
                 //double delta = 0.013;
                 dentadult = rpois(M);
                 ")

pt <- parameter_trans(
  log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
)

pomplist=list()

#Not sure about the consumption rate of alga 
parameters = c(0.5, 0.0005,0.03,0.1, 8,0.6,0.3)
names(parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")

for (i in 1:10){
  colnames(data[[i]]) <- c('day', 'dentadult')
  pomp(data = data[[i]], 
       times = "day",
       t0=1,
       rprocess=euler(dent_rpro,delta.t=1/4),
       rinit = dent_init,
       dmeasure = dmeas, 
       rmeasure = rmeas,
       partrans = pt, 
       obsnames = "dentadult",
       accumvars = c("error_count","M"),
       paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"), 
       statenames = c("S", "error_count", "F","M")
  ) -> pomplist[[i]]
  coef(pomplist[[i]])=parameters
}
names(pomplist)=paste("u", 1:10,sep = "")

shared_parameter =        c(7.773061e-05  ,2.052644e-01   ,1.790036    ,6.940058e-03  ,1.170719e-01, 0.15)
names(shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")

specific_mat = matrix(data = c(rep(0.3, 10)),
                      nrow = 1, nc = 10, byrow = T,
                      dim = list(rownames = c("sigG"),
                                 colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))

panelfood = panelPomp(pomplist, shared=shared_parameter,specific = specific_mat)

load("~/Desktop/Honor_thesis/Report_Rnw_code/sigG.RData")

df_mif <- as.data.frame(mif.estimate)
Beta <- round(df_mif[1,1] * 100000,2)
gamma <- round(df_mif[2,1] * 10,2)
kf <- round(df_mif[3,1],2)
sigF <- round(df_mif[4,1] * 10,2)
alpha <- round(df_mif[5,1] * 100,2)
Theta <- round(df_mif[6,1] * 10,2)
u1<- round(df_mif[7,1] * 10,2)
u2<- round(df_mif[8,1] * 10,2)
u3<- round(df_mif[9,1] * 10,2)
u4<- round(df_mif[10,1] * 10,2)
u5<- round(df_mif[11,1] * 10,2)
u6<- round(df_mif[12,1] * 10,2)
u7<- round(df_mif[13,1] * 10,2)
u8<- round(df_mif[14,1] * 10,2)
u9<- round(df_mif[15,1] * 10,2)
u10<- round(df_mif[16,1] * 10,2)
@

\FloatBarrier
\begin{figure}[h!]
  % \includegraphics[width=\linewidth]{sigG_Thata_simulation.jpg}
<<Dent_sim, echo=FALSE, fig.height=5>>=  
coef(panelfood) <- coef(mf[[best]]$mif)
all_params <- pparams(mf[[best]]$mif)

foreach(u = names(panelfood), .combine = rbind) %do% {
  unit_model <- unitobjects(panelfood)[[u]]
  shared <- all_params$shared
  # unit_specific <- all_params$specific[, u]
  unit_specific <- list("sigG" = all_params$specific[, u])

  pomp::simulate(
    unit_model, 
    nsim = 20,
    format = "data.frame",
    params = c(shared, unit_specific)
  ) -> sims

  sims$unit <- u
  sims
} -> all_sims

data_df <- data[[1]]
data_df$unit <- "u1"
for (i in 2:length(data)) {
  df_unit <- data[[i]]
  df_unit$unit <- paste0('u', i)

  data_df <- rbind(data_df, df_unit)
}
# all_sims %>%
#   filter(error_count == 0) -> large_removed

all_sims[all_sims$error_count == 0,] -> large_removed

ggplot() + 
  geom_line(data = large_removed, aes(x = day, y = dentadult, group = .id), col = 'blue') + 
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red') + 
  facet_wrap(~unit, scales = "free_y")
@
  \caption{Simulations of the Lotka-Voterra model with $\sigma_S$ to be unit specific and all the other  parameters to be shared on each unit of Daphnia dentifera panel data.}
  \label{figure 3}
\end{figure}

<<lum_mod, include=FALSE,echo=FALSE,results='hide'>>=
set.seed(0001)

Mesocosm_data = read_excel("~/Desktop/Final model/Mesocosmdata.xls")
dentNoPara <- Mesocosm_data[1:100, ]
dentNoPara <- subset(dentNoPara, select = c(rep, day, dent.adult))
dentNoPara <- dentNoPara[100: 1, ]
#Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
dentNoPara$day = (dentNoPara$day - 1) * 5 + 7
data = list()
trails = c("A","B","C","D","E","F","G","H","I","J")
for (i in 1: length(trails)){
  data[[i]]=subset(dentNoPara, select = c("day", "dent.adult"),
                   dentNoPara$rep == trails[i])
}

dent_rpro <- Csnippet("
                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
                      double noiB, noiG ;
                      double delta = 0.013; //fraction of volume replaced day-1

                      noiB = rnorm(0, sigB * sqrt(dt));
                      noiG = rnorm(0, sigG * sqrt(dt));

                      //------------F-------------

                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
                      Fpred =  S * F * (Beta * dt);
                      Fspl = F * (delta * dt);
                      Ffill = 0.37 * dt;

                      //------------S--------------
                      Sgrow = S * F * (Theta * dt) + S * noiG;
                      Sdeath = S * (gamma * dt);
                      Sspl = S * (delta * dt);

                      //M is the total Daphnia sample out
                      M += Sspl;
                      F += Fbirth - Fpred + Ffill - Fspl;
                      S += Sgrow - Sdeath  - Sspl;

                      if (S <= 0.0) {
                        S = 0.0;
                        error_count += 1;
                      }
                      if (F <= 0) {
                        F = 0;
                        error_count += 1000;
                      }
                      ")

# Initial state. Assume t0 = day 4
dent_init = Csnippet("
                     S = 3; //3= 45/15L
                     F = 16.667;
                     M = 0;
                     error_count = 0;
                     ")

dmeas = Csnippet("
                 double delta = 0.013;
                 double tol = 1e-18;

                 if (error_count > 0.0) {
                   lik = log(tol);
                 } else {
                   lik = dpois(dentadult,M,give_log);
                 }
                 ")

rmeas = Csnippet("
                 //double delta = 0.013;
                 dentadult = rpois(M);
                 ")

pt <- parameter_trans(
  log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
)

pomplist=list()

#Not sure about the consumption rate of alga 
parameters = c(0.5, 0.0005,0.03,0.1, 8,0.6,0.3)
names(parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")

for (i in 1:10){
  colnames(data[[i]]) <- c('day', 'dentadult')
  pomp(data = data[[i]], 
       times = "day",
       t0=1,
       rprocess=euler(dent_rpro,delta.t=1/4),
       rinit = dent_init,
       dmeasure = dmeas, 
       rmeasure = rmeas,
       partrans = pt, 
       obsnames = "dentadult",
       accumvars = c("error_count","M"),
       paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"), 
       statenames = c("S", "error_count", "F","M")
  ) -> pomplist[[i]]
  coef(pomplist[[i]])=parameters
}
names(pomplist)=paste("u", 1:10,sep = "")

shared_parameter =        c(7.773061e-05  ,2.052644e-01   ,1.790036    ,6.940058e-03  ,1.170719e-01, 0.15)
names(shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")

specific_mat = matrix(data = c(rep(0.3, 10)),
                      nrow = 1, nc = 10, byrow = T,
                      dim = list(rownames = c("sigG"),
                                 colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))

panelfood = panelPomp(pomplist, shared=shared_parameter,specific = specific_mat)

load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invasive_host/One_specific_code/sig_G/sigG_Theta.RData")

df_mif <- as.data.frame(mif.estimate)
Beta <- round(df_mif[1,1] * 100000,2)
gamma <- round(df_mif[2,1] * 10,2)
kf <- round(df_mif[3,1],2)
sigF <- round(df_mif[4,1] * 10,2)
alpha <- round(df_mif[5,1] * 100,2)
Theta <- round(df_mif[6,1] * 10,2)
u1<- round(df_mif[7,1] * 10,2)
u2<- round(df_mif[8,1] * 10,2)
u3<- round(df_mif[9,1] * 10,2)
u4<- round(df_mif[10,1] * 10,2)
u5<- round(df_mif[11,1] * 10,2)
u6<- round(df_mif[12,1] * 10,2)
u7<- round(df_mif[13,1] * 10,2)
u8<- round(df_mif[14,1] * 10,2)
u9<- round(df_mif[15,1] * 10,2)
u10<- round(df_mif[16,1] * 10,2)
@

\begin{figure}[h!]
  % \includegraphics[width=\linewidth]{lum_simulation.png}
  <<Lum_sim, echo=FALSE, fig.height=5>>=  
coef(panelfood) <- coef(mf[[best]]$mif)
all_params <- pparams(mf[[best]]$mif)

foreach(u = names(panelfood), .combine = rbind) %do% {
  unit_model <- unitobjects(panelfood)[[u]]
  shared <- all_params$shared
  # unit_specific <- all_params$specific[, u]
  unit_specific <- list("sigG" = all_params$specific[, u])

  pomp::simulate(
    unit_model, 
    nsim = 20,
    format = "data.frame",
    params = c(shared, unit_specific)
  ) -> sims

  sims$unit <- u
  sims
} -> all_sims

data_df <- data[[1]]
data_df$unit <- "u1"
for (i in 2:length(data)) {
  df_unit <- data[[i]]
  df_unit$unit <- paste0('u', i)

  data_df <- rbind(data_df, df_unit)
}

all_sims[all_sims$error_count == 0,] -> large_removed

ggplot() + 
  geom_line(data = large_removed, aes(x = day, y = dentadult, group = .id), col = 'blue') + 
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red') + 
  facet_wrap(~unit, scales = "free_y")
@
  \caption{Simulations of the Lotka-Voterra model with $\sigma_S$ to be unit specific and all the other  parameters to be shared on each unit of Daphnia lumholtzi panel data.}
  \label{figure 4}
\end{figure}
\FloatBarrier

\newpage

\section{Discussion}

Seales and colleagues \cite{6} utilized a logistic growth model to depict the population dynamics of Daphnia in the absence of parasites. However, our findings suggest that the density of Daphnia is influenced to some extent by algae, which affects its growth rate. Consequently, it appears that the fluctuations in the density of Daphnia can be attributed to the dynamics of algae, even in the presence of parasites. Further investigation through the application of a Lotka-Volterra model on single-species with parasite treatments and mixed-species treatments may provide additional insights into the impact of algae and parasites on Daphnia, given enough time.

There is a concern with the model fitting since, during simulations, the density of Daphnia occasionally exhibits exponential growth. This may be due to the lack of a parameter restricting the upper limit of Daphnia's density, which is a potential avenue to enhance the model's accuracy in future iterations. However, I chose not to incorporate this parameter into the current model as it maintains both simplicity and relevance. Results indicate that $\sigma_S$ could be the probable source of disparities in data among buckets. As this parameter is not involved in the density change of algae, it is plausible that this noise arises from some ecological process specific to Daphnia, which requires further scrutiny and discussion in future research.

We have demonstrated the applicability of a panelPOMP model for mechanistic modeling in ecological studies using Daphnia panel data. We used panelPOMP to successfully explain the large relationship between Daphnia population density and its food in Searle et al.'s experiment. It is worth noting that we only had 100 data points to explain this phenomenon. In the framework of panelPOMP, we used PIF to reduce the dimensionality of the model quickly, which means that this method can be utilized for the analysis of other high-dimensional data and has the advantage of rapid dimensionality reduction. This approach complements the disadvantages of the Mechanistic model and potentially allows ecologists to try to solve more data analysis problems using the Mechanistic model.

The R package panelPomp, which is based on this model and PIF algorithm developed by Bretó et al.\cite{1}, enables access to latent state variables such as algae, offering a quantitative approach that facilitates parameter estimation, hypothesis testing, and the comparison of different models with or without their influence on the dynamic system. This approach provides scientists with better insight into the contribution of each factor to the whole system and the ability to predict its future state. Furthermore, the package allows for simulations, which serve as a valuable tool for testing the goodness-of-fit of chosen models and parameters. Overall, mechanistic modeling under a panelPOMP framework can yield more information from limited ecological data, increasing the efficiency and accuracy of ecological experiments while reducing cost and resource requirements.

\newpage
<<Read_Data, include=FALSE,echo=FALSE,results='hide'>>=
AIC_two <- function(ll){
  temp <- -2 * ll + 50
  return temp
}

AIC_one <- function(ll){
  temp <- -2 * ll + 32
  return temp
}
# --dent--L-V-food_management--
#dent alpha_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/alpha_Beta/alpha_Beta.RData")
d_aB_dim <- 25
d_aB_ll <- round(pf.loglik.of.mif.estimate,2)
d_aB_AIC <- AIC_two(d_aB_ll)

#dent sigG_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigG_gamma/sigG_gamma.RData")
d_sGg_dim <- 25
d_sGg_ll <- round(pf.loglik.of.mif.estimate,2)
d_sGg_AIC <- AIC_two(d_sGg_ll)

#dent alpha_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/alpha_gamma/alpha_gamma.RData")
d_ag_dim <- 25
d_ag_ll <- round(pf.loglik.of.mif.estimate,2)
d_ag_AIC <- AIC_two(d_ag_ll)

#dent alpha_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/alpha_kf/alpha_kf.RData")
d_ak_dim <- 25
d_ak_ll <- round(pf.loglik.of.mif.estimate,2)
d_ak_AIC <- AIC_two(d_ak_ll)

#dent Beta_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/Beta_gamma/Beta_gamma.RData")
d_Bg_dim <- 25
d_Bg_ll <- round(pf.loglik.of.mif.estimate,2)
d_Bg_AIC <- AIC_two(d_Bg_ll)

#dent kf_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/kf_Beta/kf_Beta.RData")
d_kB_dim <- 25
d_kB_ll <- round(pf.loglik.of.mif.estimate,2)
d_kB_AIC <- AIC_two(d_kB_ll)

#dent kf_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/kf_gamma/kf_gamma.RData")
d_kg_dim <- 25
d_kg_ll <- round(pf.loglik.of.mif.estimate,2)
d_kg_AIC <- AIC_two(d_kg_ll)

#dent kf_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/kf_Theta/kf_Theta.RData")
d_kT_dim <- 25
d_kT_ll <- round(pf.loglik.of.mif.estimate,2)
d_kT_AIC <- AIC_two(d_kT_ll)

#dent sigB_alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_alpha/sigB_alpha.RData")
d_sBa_dim <- 25
d_sBa_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBa_AIC <- AIC_two(d_sBa_ll)

#dent sigB_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_Beta/sigB_Beta.RData")
d_sBB_dim <- 25
d_sBB_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBB_AIC <- AIC_two(d_sBB_ll)

#dent sigB_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_gamma/sigB_gamma.RData")
d_sBg_dim <- 25
d_sBg_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBg_AIC <- AIC_two(d_sBg_ll)

#dent sigB_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_kf/sigB_kf.RData")
d_sBk_dim <- 25
d_sBk_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBk_AIC <- AIC_two(d_sBk_ll)

#dent sigB_sigG
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_sigG/sigB_sigG.RData")
d_sBsG_dim <- 25
d_sBsG_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBsG_AIC <- AIC_two(d_sBsG_ll)

#dent sigB_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigB_Theta/sigB_Theta.RData")
d_sBT_dim <- 25
d_sBT_ll <- round(pf.loglik.of.mif.estimate,2)
d_sBT_AIC <- AIC_two(d_sBT_ll)

#dent sigG_alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigG_alpha/sigG_alpha.RData")
d_sGa_dim <- 25
d_sGa_ll <- round(pf.loglik.of.mif.estimate,2)
d_sGa_AIC <- AIC_two(d_sGa_ll)

#dent sigG_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigG_Beta/sigG_Beta.RData")
d_sGB_dim <- 25
d_sGB_ll <- round(pf.loglik.of.mif.estimate,2)
d_sGB_AIC <- AIC_two(d_sGB_ll)

#dent sigG_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigG_kf/sigG_kf.RData")
d_sGk_dim <- 25
d_sGk_ll <- round(pf.loglik.of.mif.estimate,2)
d_sGk_AIC <- AIC_two(d_sGk_ll)

#dent sigG_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/sigG_Theta/sigG_Theta.RData")
d_sGT_dim <- 25
d_sGT_ll <- round(pf.loglik.of.mif.estimate,2)
d_sGT_AIC <- AIC_two(d_sGT_ll)

#dent Theta_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/Theta_Beta/Theta_Beta.RData")
d_TB_dim <- 25
d_TB_ll <- round(pf.loglik.of.mif.estimate,2)
d_TB_AIC <- AIC_two(d_TB_ll)

#dent Theta_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/LL_two_specific_code/Theta_gamma/Theta_gamma.RData")
d_Tg_dim <- 25
d_Tg_ll <- round(pf.loglik.of.mif.estimate,2)
d_Tg_AIC <- AIC_two(d_Tg_ll)

#dent sigG
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/sigG/sigG.RData")
d_sG_dim <- 16
d_sG_ll <- round(pf.loglik.of.mif.estimate,2)
d_sG_AIC <- AIC_one(d_sG_ll)

#dent sigB
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/sigB/sigB.RData")
d_sB_dim <- 16
d_sB_ll <- round(pf.loglik.of.mif.estimate,2)
d_sB_AIC <- AIC_one(d_sB_ll)

#dent alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/alpha/alpha.RData")
d_a_dim <- 16
d_a_ll <- round(pf.loglik.of.mif.estimate,2)
d_a_AIC <- AIC_one(d_a_ll)

#dent Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/Beta/Beta.RData")
d_B_dim <- 16
d_B_ll <- round(pf.loglik.of.mif.estimate,2)
d_B_AIC <- AIC_one(d_B_ll)

#dent Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/Theta/Theta.RData")
d_T_dim <- 16
d_T_ll <- round(pf.loglik.of.mif.estimate,2)
d_T_AIC <- AIC_one(d_T_ll)

#dent gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/gamma/gamma.RData")
d_g_dim <- 16
d_g_ll <- round(pf.loglik.of.mif.estimate,2)
d_g_AIC <- AIC_one(d_g_ll)

#dent kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Native_host/One_specific_code/kf/kf.RData")
d_k_dim <- 16
d_k_ll <- round(pf.loglik.of.mif.estimate,2)
d_k_AIC <- AIC_one(d_k_ll)

# --lum--L-V-food_management--
#lum alpha_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/alpha_Beta/alpha_Beta.RData")
l_aB_dim <- 25
l_aB_ll <- round(pf.loglik.of.mif.estimate,2)
l_aB_AIC <- AIC_two(l_aB_ll)

#lum sigG_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigG_gamma/sigG_gamma.RData")
l_sGg_dim <- 25
l_sGg_ll <- round(pf.loglik.of.mif.estimate,2)
l_sGg_AIC <- AIC_two(l_sGg_ll)

#lum alpha_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/alpha_gamma/alpha_gamma.RData")
l_ag_dim <- 25
l_ag_ll <- round(pf.loglik.of.mif.estimate,2)
l_ag_AIC <- AIC_two(l_ag_ll)

#lum alpha_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/alpha_kf/alpha_kf.RData")
l_ak_dim <- 25
l_ak_ll <- round(pf.loglik.of.mif.estimate,2)
l_ak_AIC <- AIC_two(l_ak_ll)

#lum Beta_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/Beta_gamma/Beta_gamma.RData")
l_Bg_dim <- 25
l_Bg_ll <- round(pf.loglik.of.mif.estimate,2)
l_Bg_AIC <- AIC_two(l_Bg_ll)

#lum kf_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/kf_Beta/kf_Beta.RData")
l_kB_dim <- 25
l_kB_ll <- round(pf.loglik.of.mif.estimate,2)
l_kB_AIC <- AIC_two(l_kB_ll)

#lum kf_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/kf_gamma/kf_gamma.RData")
l_kg_dim <- 25
l_kg_ll <- round(pf.loglik.of.mif.estimate,2)
l_kg_AIC <- AIC_two(l_kg_ll)

#lum kf_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/kf_Theta/kf_Theta.RData")
l_kT_dim <- 25
l_kT_ll <- round(pf.loglik.of.mif.estimate,2)
l_kT_AIC <- AIC_two(l_kT_ll)

#lum sigB_alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_alpha/sigB_alpha.RData")
l_sBa_dim <- 25
l_sBa_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBa_AIC <- AIC_two(l_sBa_ll)

#lum sigB_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_Beta/sigB_Beta.RData")
l_sBB_dim <- 25
l_sBB_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBB_AIC <- AIC_two(l_sBB_ll)

#lum sigB_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_gamma/sigB_gamma.RData")
l_sBg_dim <- 25
l_sBg_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBg_AIC <- AIC_two(l_sBg_ll)

#lum sigB_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_kf/sigB_kf.RData")
l_sBk_dim <- 25
l_sBk_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBk_AIC <- AIC_two(l_sBk_ll)

#lum sigB_sigG
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_sigG/sigB_sigG.RData")
l_sBsG_dim <- 25
l_sBsG_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBsG_AIC <- AIC_two(l_sBsG_ll)

#lum sigB_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigB_Theta/sigB_Theta.RData")
l_sBT_dim <- 25
l_sBT_ll <- round(pf.loglik.of.mif.estimate,2)
l_sBT_AIC <- AIC_two(l_sBT_ll)

#lum sigG_alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigG_alpha/sigG_alpha.RData")
l_sGa_dim <- 25
l_sGa_ll <- round(pf.loglik.of.mif.estimate,2)
l_sGa_AIC <- AIC_two(l_sGa_ll)

#lum sigG_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigG_Beta/sigG_Beta.RData")
l_sGB_dim <- 25
l_sGB_ll <- round(pf.loglik.of.mif.estimate,2)
l_sGB_AIC <- AIC_two(l_sGB_ll)

#lum sigG_kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigG_kf/sigG_kf.RData")
l_sGk_dim <- 25
l_sGk_ll <- round(pf.loglik.of.mif.estimate,2)
l_sGk_AIC <- AIC_two(l_sGk_ll)

#lum sigG_Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/sigG_Theta/sigG_Theta.RData")
l_sGT_dim <- 25
l_sGT_ll <- round(pf.loglik.of.mif.estimate,2)
l_sGT_AIC <- AIC_two(l_sGT_ll)

#lum Theta_Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/Theta_Beta/Theta_Beta.RData")
l_TB_dim <- 25
l_TB_ll <- round(pf.loglik.of.mif.estimate,2)
l_TB_AIC <- AIC_two(l_TB_ll)

#lum Theta_gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/LL_two_specific_code/Theta_gamma/Theta_gamma.RData")
l_Tg_dim <- 25
l_Tg_ll <- round(pf.loglik.of.mif.estimate,2)
l_Tg_AIC <- AIC_two(l_Tg_ll)

#lum sigG
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/sigG/sigG.RData")
l_sG_dim <- 16
l_sG_ll <- round(pf.loglik.of.mif.estimate,2)
l_sG_AIC <- AIC_one(l_sG_ll)

#lum sigB
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/sigB/sigB.RData")
l_sB_dim <- 16
l_sB_ll <- round(pf.loglik.of.mif.estimate,2)
l_sB_AIC <- AIC_one(l_sB_ll)

#lum alpha
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/alpha/alpha.RData")
l_a_dim <- 16
l_a_ll <- round(pf.loglik.of.mif.estimate,2)
l_a_AIC <- AIC_one(l_a_ll)

#lum Beta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/Beta/Beta.RData")
l_B_dim <- 16
l_B_ll <- round(pf.loglik.of.mif.estimate,2)
l_B_AIC <- AIC_one(l_B_ll)

#lum Theta
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/Theta/Theta.RData")
l_T_dim <- 16
l_T_ll <- round(pf.loglik.of.mif.estimate,2)
l_T_AIC <- AIC_one(l_T_ll)

#lum gamma
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/gamma/gamma.RData")
l_g_dim <- 16
l_g_ll <- round(pf.loglik.of.mif.estimate,2)
l_g_AIC <- AIC_one(l_g_ll)

#lum kf
load("~/Desktop/Honor_thesis/Daphnia_alga_model/Invaisve_host/One_specific_code/kf/kf.RData")
l_k_dim <- 16
l_k_ll <- round(pf.loglik.of.mif.estimate,2)
l_k_AIC <- AIC_one(l_k_ll)
@
\section{Appendix}
Tables for each simulations and calculations
% \subsection{Table for Logistic models}
\FloatBarrier
\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c||} 
 \hline
Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex] 
 \hline\hline
   $K$                 & \Sexpr{lg_one_dim}       & \Sexpr{lg_K_ll}           & \Sexpr{lg_K_AIC} \\
   $K + \sigma_S$        & \Sexpr{lg_two_dim}        & \Sexpr{lg_KsG_ll}            & \Sexpr{lg_KsG_AIC}   \\
   all specific        & \Sexpr{lg_zero_dim}         & \Sexpr{lg_KsG_ll}             & \Sexpr{lg_KsG_AIC}   \\
   $\sigma_S$          & \Sexpr{lg_one_dim}        & \Sexpr{lg_sG_ll}           & \Sexpr{lg_sG_AIC} \\
   $r$ + $K$           & \Sexpr{lg_two_dim}        & \Sexpr{lg_rK_ll}            & \Sexpr{lg_rK_AIC} \\
   $r$ + $\sigma_S$    & \Sexpr{lg_two_dim}        & \Sexpr{lg_rsG_ll}            & \Sexpr{lg_rsG_AIC}  \\
   $r$                 & \Sexpr{lg_one_dim}        & \Sexpr{lg_r_ll}          & \Sexpr{lg_r_AIC} \\[1ex]
 \hline
\end{tabular}
\caption{This table includes logistic growth models with different choices of parameters to be either unit specific or shared}
\label{Table:1}
\end{table}

% \subsection{Table for Lotka-Volterra models with enought food supply}
\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c||} 
 \hline
Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex] 
 \hline\hline
 $\alpha$               & \Sexpr{one_dim}        & \Sexpr{x_a_ll}           & \Sexpr{x_a_AIC} \\
   $\beta$                & \Sexpr{one_dim}        & \Sexpr{x_B_ll}          & \Sexpr{x_B_AIC} \\
   $\gamma$               & \Sexpr{one_dim}        & \Sexpr{x_g_ll}           & \Sexpr{x_g_AIC} \\
   $\theta$               & \Sexpr{one_dim}        & \Sexpr{x_T_ll}           & \Sexpr{x_T_AIC} \\
   $\sigma_S$             & \Sexpr{one_dim}        & \Sexpr{x_sG_ll}          & \Sexpr{x_sG_AIC}  \\
   $\sigma_F$             & \Sexpr{one_dim}        & \Sexpr{x_sB_ll}           & \Sexpr{x_sB_AIC} \\
   $k_f$                  & \Sexpr{one_dim}        & \Sexpr{x_k_ll}           & \Sexpr{x_k_AIC}  \\
   $\alpha , k_f$         & \Sexpr{two_dim}        & \Sexpr{x_ak_ll}           & \Sexpr{x_ak_AIC}   \\
   $\alpha , \beta , k_f$ & \Sexpr{thr_dim}        & \Sexpr{x_aBk_ll}           & \Sexpr{x_aBk_AIC}  \\
   $\beta , k_f$          & \Sexpr{two_dim}        & \Sexpr{x_Bk_ll}           & \Sexpr{x_Bk_AIC} \\
   $\alpha,\gamma,k_f$    & \Sexpr{thr_dim}        & \Sexpr{x_agk_ll}          & \Sexpr{x_agk_AIC} \\
   $\theta,k_f$           & \Sexpr{two_dim}        & \Sexpr{x_Tk_ll}          & \Sexpr{x_Tk_AIC} \\
   $\alpha,\theta$        & \Sexpr{two_dim}        & \Sexpr{x_aT_ll}           & \Sexpr{x_aT_AIC} \\
 all specific           & \Sexpr{all_dim}        & \Sexpr{x_all_ll}           & \Sexpr{x_all_AIC} \\[1ex]
 \hline
\end{tabular}
\caption{This table includes Lotka-Volterra model with enough food supply with different choices of parameters to be either unit specific or shared}
\label{Table:2}
\end{table}

% \subsection{Table for Lotka-Volterra models with food supply management}
\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c||} 
 \hline
Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex] 
 \hline\hline
   $\sigma_S$   &   \Sexpr{one_dim}      & \Sexpr{d_sG_ll}          & \Sexpr{d_sG_ll}  \\
 $\sigma_S,\theta$   & \Sexpr{two_dim}        & \Sexpr{d_sGT_ll}          & \Sexpr{d_sGT_ll}    \\
   $\sigma_S,\beta$    & \Sexpr{two_dim}        & \Sexpr{d_sGB_ll}         & \Sexpr{d_sGB_ll}    \\
   $\sigma_S,\alpha$   & \Sexpr{two_dim}        & \Sexpr{d_sGa_ll}         & \Sexpr{d_sGa_ll}  \\
   $\sigma_S,\gamma$   & \Sexpr{two_dim}        & \Sexpr{d_sGg_ll}          & \Sexpr{d_sGg_ll}  \\
   $\sigma_S,\sigma_F$ & \Sexpr{two_dim}        & \Sexpr{d_sBsG_ll}          & \Sexpr{d_sBsG_ll}  \\
 $\sigma_S,k_f$      & \Sexpr{two_dim}        & \Sexpr{d_sGk_ll}          & \Sexpr{d_sGk_ll} \\
   $\alpha, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBa_ll}          & \Sexpr{d_sBa_ll} \\
   $\alpha, \beta$     & \Sexpr{two_dim}        & \Sexpr{d_aB_ll}            & \Sexpr{d_aB_ll}   \\
      $\alpha$   & \Sexpr{one_dim}        & \Sexpr{d_a_ll}          & \Sexpr{d_a_ll}  \\
   $\beta,\gamma$      & \Sexpr{two_dim}        & \Sexpr{d_Bg_ll}         & \Sexpr{d_Bg_ll} \\
  $\gamma, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBg_ll}          & \Sexpr{d_sBg_ll}  \\
   $\beta, \sigma_F$   & \Sexpr{two_dim}        & \Sexpr{d_sBB_ll}          & \Sexpr{d_sBB_ll}  \\
   $\sigma_F$   & \Sexpr{one_dim}        & \Sexpr{d_sB_ll}          & \Sexpr{d_sB_ll}  \\
   $\alpha,\theta$     & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}         & \Sexpr{d_aT_ll}     \\
   $\theta, \gamma$    & \Sexpr{two_dim}        & \Sexpr{d_Tg_ll}          & \Sexpr{d_Tg_ll} \\
   $\alpha,\gamma$     & \Sexpr{two_dim}        & \Sexpr{d_ag_ll}          & \Sexpr{d_ag_ll} \\
  $\beta$   & \Sexpr{one_dim}        & \Sexpr{d_B_ll}         & \Sexpr{d_B_ll}  \\
  $k_f, \sigma_F$     & \Sexpr{two_dim}        & \Sexpr{d_sBk_ll}          & \Sexpr{d_sBk_ll}  \\
   $\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          & \Sexpr{d_TB_ll}  \\
   $\theta$   & \Sexpr{one_dim}        & \Sexpr{d_T_ll}          & \Sexpr{d_T_ll}    \\
 $\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          &  \Sexpr{d_TB_ll}\\
   $\alpha, \theta$    & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}          & \Sexpr{d_aT_ll}  \\
      $k_f$   & \Sexpr{one_dim}        & \Sexpr{d_k_ll}          & \Sexpr{d_k_ll}   \\
   $\alpha , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_ak_ll}            & \Sexpr{d_ak_ll}   \\
   $\gamma , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_gk_ll}          & \Sexpr{d_gk_ll} \\
   $\theta , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_Tk_ll}          & \Sexpr{d_Tk_ll} \\
   $\beta , k_f$       & \Sexpr{two_dim}        & \Sexpr{d_Bk_ll}         & \Sexpr{d_Bk_ll} \\[1ex]
 \hline
\end{tabular}
\caption{This table includes newly introduced Lotka-Volterra models fitted to Daphnia lumniza with different choices of parameters to be either unit specific or shared}
\label{Table:3}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c||}
 \hline
Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex]
 \hline\hline
   $\sigma_S$   & \Sexpr{one_dim}        & \Sexpr{l_sG_ll}          & \Sexpr{l_sG_AIC}  \\
  $\sigma_S,\beta$  & \Sexpr{two_dim}        & \Sexpr{l_sB_ll}          & \Sexpr{l_sB_AIC}    \\
     $\sigma_S,\theta$  & \Sexpr{two_dim}        & \Sexpr{l_sT_ll}         & \Sexpr{l_sT_AIC}   \\
     $\alpha, \beta$ & \Sexpr{two_dim}        & \Sexpr{l_sB_ll}          & \Sexpr{l_sB_AIC}  \\
     $\sigma_S,\sigma_F$ & \Sexpr{two_dim}        & \Sexpr{l_sBsG_ll}          & \Sexpr{l_sBsG_AIC}   \\
   $\sigma_S,\alpha$ & \Sexpr{two_dim}        & \Sexpr{l_sGa_ll}          & \Sexpr{l_sGa_AIC}  \\
  $\alpha, \sigma_F$    & \Sexpr{two_dim}        & \Sexpr{l_sBa_ll}          & \Sexpr{l_sBa_AIC} \\
   $\sigma_S,k_f$   & \Sexpr{two_dim}        & \Sexpr{l_sGk_ll}          & \Sexpr{l_sGk_AIC}  \\
    $\sigma_S,\gamma$    & \Sexpr{two_dim}        & \Sexpr{l_sGg_ll}            & \Sexpr{l_sGg_AIC}   \\
     $\beta,\gamma$   & \Sexpr{one_dim}        & \Sexpr{l_Bg_ll}          & \Sexpr{l_Bg_AIC}  \\
      $\alpha$    & \Sexpr{two_dim}        & \Sexpr{l_a_ll}          & \Sexpr{l_a_AIC} \\
  $\alpha,\theta$ & \Sexpr{two_dim}        & \Sexpr{l_aT_ll}          & \Sexpr{l_aT_AIC} \\
   $\beta, \sigma_F$   & \Sexpr{two_dim}        & \Sexpr{l_sBB_ll}          & \Sexpr{l_sBB_AIC} \\
   $\sigma_F$   & \Sexpr{one_dim}        & \Sexpr{l_sB_ll}          & \Sexpr{l_sB_AIC}   \\
     $\alpha,\gamma$    & \Sexpr{two_dim}        & \Sexpr{l_ag_ll}          & \Sexpr{l_ag_AIC}    \\
     $k_f, \sigma_F$    & \Sexpr{two_dim}        & \Sexpr{l_sBk_ll}         & \Sexpr{l_sBk_AIC} \\
     $\gamma, \sigma_F$   & \Sexpr{two_dim}        & \Sexpr{l_sBg_ll}          & \Sexpr{l_sBg_AIC} \\
   $\theta , \beta$  & \Sexpr{one_dim}        & \Sexpr{l_TB_ll}           & \Sexpr{l_TB_AIC}  \\
   $\theta, \gamma$  & \Sexpr{two_dim}        & \Sexpr{l_Tg_ll}          & \Sexpr{l_Tg_AIC}  \\
     $\alpha , k_f$  & \Sexpr{two_dim}        & \Sexpr{l_ak_ll}          & \Sexpr{l_ak_AIC} \\
   $\theta$   & \Sexpr{one_dim}        & \Sexpr{l_T_ll}           & \Sexpr{l_T_AIC}  \\
 $\alpha, \theta$  & \Sexpr{two_dim}        & \Sexpr{l_aT_ll}           & \Sexpr{l_aT_AIC}\\
      $\theta , k_f$   & \Sexpr{two_dim}        & \Sexpr{l_Tk_ll}           & \Sexpr{l_Tk_AIC} \\
        $\beta , k_f$  & \Sexpr{one_dim}        & \Sexpr{l_Bk_ll}           & \Sexpr{l_Bk_AIC}  \\
     $\beta$    & \Sexpr{two_dim}        & \Sexpr{l_B_ll}             & \Sexpr{l_B_AIC}    \\
   $\gamma , k_f$      & \Sexpr{two_dim}        & \Sexpr{l_gk_ll}          & \Sexpr{l_gk_AIC} \\
      $\theta , \beta$   & \Sexpr{two_dim}        & \Sexpr{l_TB_ll}           & \Sexpr{l_TB_AIC} \\
     $k_f$    & \Sexpr{two_dim}        & \Sexpr{l_k_ll}           & \Sexpr{l_k_AIC}  \\[1ex]
 \hline
\end{tabular}
\caption{This table includes newly introduced Lotka-Volterra models fitted to Daphnia lumniza with different choices of parameters to be either unit specific or shared}
\label{Table:3}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c|c|c|c|c||} 
 \hline
 unit & $\alpha \cdot 10$     & $\beta \cdot 10^4 $      & $\gamma \cdot 10 $     & $\theta \cdot 10$     & $k_f$        & $\sigma_S \cdot 10$   & $\sigma_F \cdot 10$ \\ [0.5ex] 
 \hline\hline
 u1   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u1} & \Sexpr{sigF} \\
   u2   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma}& \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u2} & \Sexpr{sigF} \\
   u3   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u3} & \Sexpr{sigF} \\
   u4   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u4} & \Sexpr{sigF} \\
   u5   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u5} & \Sexpr{sigF} \\
  u6   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u6} & \Sexpr{sigF}\\
   u7   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u7} & \Sexpr{sigF} \\
   u8   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u8} & \Sexpr{sigF} \\
   u9   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta}& \Sexpr{kf} & \Sexpr{u9} & \Sexpr{sigF} \\
   u10  & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u10} & \Sexpr{sigF} \\[1ex]
 \hline
\end{tabular}
\caption{This table includes parameter swarm that gets best AIC with Lotka-Volterra models fitted on Daphnia dentifera panel data}
\label{Table:1}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{||c |c |c |c|c|c|c|c||} 
 \hline
 unit & $\alpha \cdot 10$     & $\beta \cdot 10^4 $      & $\gamma  $     & $\theta \cdot 10$     & $k_f $        & $\sigma_S \cdot 10$   & $\sigma_F \cdot 10$ \\ [0.5ex] 
 \hline\hline
 u1   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u1} & \Sexpr{sigF} \\
   u2   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma}& \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u2} & \Sexpr{sigF} \\
   u3   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u3} & \Sexpr{sigF} \\
   u4   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u4} & \Sexpr{sigF} \\
   u5   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u5} & \Sexpr{sigF} \\
  u6   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u6} & \Sexpr{sigF}\\
   u7   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u7} & \Sexpr{sigF} \\
   u8   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u8} & \Sexpr{sigF} \\
   u9   & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta}& \Sexpr{kf} & \Sexpr{u9} & \Sexpr{sigF} \\
   u10  & \Sexpr{alpha} & \Sexpr{Beta} & \Sexpr{gamma} & \Sexpr{Theta} & \Sexpr{kf} & \Sexpr{u10} & \Sexpr{sigF} \\[1ex]
 \hline
\end{tabular}
\caption{This table includes parameter swarm that gets best AIC with Lotka-Volterra models fitted on Daphnia lumholtzi panel data}
\label{Table:2}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{||c |c||}
 \hline
parameter & unit\\[0.5ex]
\hline\hline
$\alpha$ &$\left(6hr \right)^{-1}$\\
$\beta $& $L \cdot individual^{-1} \cdot \left(6hr \right)^{-1}$\\
$\gamma $ &  $\left(6hr \right)^{-1}$\\
$\theta $ & $ L \cdot \left( 10^5 cells \right) ^{-1} \cdot \left(6hr \right)^{-1}$\\
$k_f$ & $10^5 cells \cdot L $ \\
$\sigma_S $ & $\left(6hr \right)^{-1}$\\
$\sigma_F $ & $\left(6hr \right)^{-1}$\\[1ex]
 \hline
\end{tabular}
\caption{This table shows the units of the parameters}
\label{Table:3}
\end{table}

\FloatBarrier
\bibliography{reference}
\end{document}

### 2023-12-12 Paper.tex
Update of paper

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear,numbers]{natbib}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyphenat}
\tolerance=1000
\usepackage{url}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{hyphenat}
\usepackage{Sweave}
\bibliographystyle{apalike}
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\begin{document}
\input{Paper-concordance}
\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\if1\blind
{
  \title{\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
  \author{Author 1\thanks{
    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
    Department of YYY, University of XXX\\
    and \\
    Author 2 \\
    Department of ZZZ, University of WWW}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
In ecological research, panel data, often comprising multi-magnitude time series, presents unique analytical challenges due to its high dimensionality. This study employs the PanelPOMP framework and the Panel Iterated Filtering (PIF) method to explore the practicality and advancement of mechanistic model development for such complex datasets. Mechanistic models, especially valuable in ecological research, allow for quantitative analysis and deeper understanding of dynamic systems. However, they are often constrained by the non-Gaussian nature of ecological data. Our study applies these methods to the Daphnia panel data, consisting of independent time series with 10 data points each, to derive parameters that minimize the Akaike Information Criterion (AIC). The comparison of AIC values across models reveals the significant impact of alga, the prey of Daphnia, and parasite on their population dynamics which is observed in the SIRPF model highlights this relationship. The panelPOMP framework's ability to handle complex panel data analysis is demonstrated, emphasizing its utility in uncovering latent processes not captured by traditional models. However, the study also discusses limitations and interpretations of mechanistic models and PanelPOMP framework in ecological data analysis.
\end{abstract}

\noindent%
{\it Keywords:}  Panel Iterated Filtering; SIRPF model; PanelPOMP; Daphnia. 
\vfill

\newpage
\spacingset{1.9} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary. As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \citep{Stalidzans2020}. Utilizing the established process, they elucidate the actions of hidden mechanisms \citep{Duarte2003}. The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems. This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters. Mechanistic modeling has been widely employed in social sciences, industrial engineering, and physics for an extended period. Improved computational capabilities now make it more effective. Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information. Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task. As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \citep{Levin_1992}.\\

Panel data, sometimes referred to as longitudinal data, is composed of multiple time series. Each of these time series, which may be multivariate, contains a series of observations collected from a separate entity  \cite{Carles2020}. In ecology, panel data is a traditional data form that researchers gather to mitigate the impact of random variations throughout experiments, so that it can more accurately measure changes in experimental objectives and control the variables needed to achieve experimental goals. Due to the non-linear characteristics of ecological data, the analysis of panel data entails a high-dimensional structure. Such features diminish the explanatory power of traditional time series analysis and certain Monte Carlo inference techniques \cite{johnstone2009statistical}. The goal of this study is to investigate the practicality of employing mechanistic models on ecological panel data using the panelPOMP structure.\\

The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \cite{King2016}. This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering. Stemming from iterated filtering \cite{King2016} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood. This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.\\

In this research, we will show the advancement of the panelPOMP model by analyzing the Daphnia panel data generated in the experiment of Searle et al. \cite{Searle2016}. The research group of Searle et al. collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \cite{Searle2016}. In their lab experiment, they generated standard ecological panel data featuring various independent trials under the same experimental conditions, with each consisting of a brief, nonlinear, and nonstationary time series. Conventional mechanistic theories cannot directly interpret these observations. Utilizing the panelPOMP approach allows for the combination of theoretical models and empirical data, enabling the extraction of the comprehensive information found in each individual time series \cite{Carles2020}.\\

For a more in-depth examination of the concealed processes in the panel data, this research incorporates the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model - within the panelPOMP framework. After applying the model to the Daphnia panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach. These findings are subsequently compared to discover a more accurate and appropriate model for the given data. The superior performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is Daphnia's prey, impact the dynamics of both species of Daphnia. It is worth noting that the population density of alga is neither observable nor recorded in the experiment. That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes \cite{Edwards2012}. At the same time, the panelPOMP framework provides an estimation of the parameters in the model, which helps to understand the hidden mechanisms in the ecological system \cite{Carles2020}. Based on the estimation, a detailed interpretation of relationship between invasive host and native host with presence of parasite is explaned.\\

Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the Daphnia panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system. By analyzing Daphnia panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data. Additionally, the later section will further discuss how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models.

\section{Study System}
\label{sec:studysystem}

The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two Daphnia species: Daphnia \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \cite{searle2016salinization}, and Daphnia \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia. D. lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \cite{havel1993daphnia}. These Daphnia species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines. For their experiments, six clones of each species were selected to eliminate selection bias \cite{Searle2016}.\\

The dynamics also involves \textit{Metschnikowia bicuspidata}, a parasitic fungus predominantly infecting the native host. This parasite, ingested as spores by Daphnia, aggressively affects the host's fecundity and longevity \cite{duffy2010temporal}. The parasite's presence is identifiable by the opaqueness it causes in normally transparent hosts, with spore release upon host death. Searle's study utilized a M. bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \cite{Searle2016}. The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for Daphnia. \textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \cite{arnold1971ingestion}. Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of Daphnia species can be closely monitored and manipulated \cite{martinez1994effect}.\\

Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two Daphnia host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \cite{Searle2016}. This observation suggests a potential competitive advantage for the invasive species. Building on these insights, Searle and colleagues structured the ecological system we analyze here, focusing on dynamics at the community level. Their approach involved a mesocosm experiment that varied species composition and parasite presence. The experimental design was a 3x2x1 factorial arrangement, incorporating three species combinations (native only, invasive only, or both existing species), two parasite exposure scenarios (exposed or unexposed), and a single algal food source. This design enabled an in-depth analysis of the invasive host's impact on native populations, manifesting either directly through competition for resources, indirectly via the amplification of a mutual parasite, or through a reduction in parasite effects.\\

In essence, the impact of an invasive Daphnia \textit{lumholtzi}, particularly in terms of resource competition like food, can be significant. This scenario would lead to a decreased survival rate of the native host in the presence of the invasive species, irrespective of parasite interference. This effect is quantifiable through the ratio of natality to mortality in the native host population. Conversely, should the invasive species primarily influence the native host indirectly via interactions with parasites, a different pattern emerges. Here, the survivability of the native host is expected to be lower in scenarios involving both the parasite and the invasive host, as opposed to situations where the native host is only affected by the parasite. This effect is quantifiable through the ratio of natality to mortality in the native host population as well.\\

\subsection{Non-parasitized Dynamics}
This research stratifies the ecological dynamics above into two analytical categories delineated by parasitic presence. Under the 'Non-parasitized Dynamics' classification, three experimental groups of Daphnia species—solely native, solely invasive, or a combined population—are statistically analyzed. This categorization facilitates a quantitative assessment of the interaction dynamics between native and invasive Daphnia species, particularly in environments not influenced by parasitic interactions.\\

In the Non-parasitized level, indoor mesocosms were used \cite{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media. The experimental design involved 10 replicates for single-species treatments, and 9 replicates for mixed-species treatment. A nitrogen-to-phosphorus ratio of 25:1 was established, and each mesocosm was seeded with 2.5 $\times$ $10^8$ cells of \textit{Ankistrodesmus falcatus}. The laboratory conditions included an average temperature of 23.27$^\circ C$ ($\pm$0.2) and a 16-hour light, 8-hour dark cycle. For single-species treatments, either 45 native or invasive Daphnia hosts were added, representing six genotypes per species to eliminate the selection bias. In mixed-species treatments, 35 native and 10 invasive hosts were introduced, simulating early invasion stages.\\

\subsection{Parasitized Dynamics}
Parasitized dynamics is employed with a key difference, which is the presence of the parasite \textit{Metschnikowia bicuspidata} . Based on the same setup of Non-parasitized Dynamics, after a 4-day acclimation period in the mesocosms, hosts are exposed to parasites, with spore counts set at 25 per mL in parasite-exposed treatments. This category allows for the investigation of the impact of parasites on the interaction dynamics between native and invasive Daphnia species, particularly how parasitism influences competition and survival rates. This approach complements the "Non-parasitized Dynamics" by providing a contrasting scenario where parasitic factors are at play, thereby enriching the understanding of ecological interactions under the presence of parasite.\\

\subsection{Data Collection and Visualization}
Population densities and infection prevalence(if exists) were quantified every five days, beginning a week after the experiment started. On each sampling day, mesocosms were thoroughly mixed, and a 1-liter sample was removed and replaced with fresh COMBO media. The samples were concentrated using a 153-$\mu m$ mesh and the Daphnia were counted under a microscope. For each Daphnia, species (native or invasive), infection status, age, and sex were determined. Regular stirring of mesocosms ensured resuspension of algae and parasites. Algal and nutrient levels were supplemented bi-weekly, with COMBO media added as necessary. The experiment concluded after 52 days, totaling 10 sampling sessions \cite{Searle2016}.\\

As juveniles rarely exhibit clear signs of infection, only adult Daphnia were used for quantifying effects in both Non-parasitized and Parasitized Dynamics to minimize information bias. Figures below illustrates these dynamics, with the figure 1 depicting Non-parasitized Dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}. Figure 2 shows Parasitized Dynamics, where the interplay of susceptible and infected Daphnia of different species is further influenced by parasites, complicating the variation of susceptible and infected Daphnia of different races, which makes it difficult to directly analyze the sources of influence on their variation. Non-parasitized Dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species. And based on this result, the correlation between native host and invasive host in Parasitized Dynamics will be further explored.

\begin{figure}[htbp]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics{Paper-002}
\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\includegraphics{Paper-003}
\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
\end{subfigure}%
\end{figure}

\section{Methodology}
\label{sec:meth}
\subsection{Partially Observed Markov Processes}
Frequently, we lack a comprehensive understanding of the underlying mechanisms that drive the evolution of a natural system. For example, the precise number of individuals who are exposed to a disease or the density of alga at any given time is typically unknown. To address these issues of missing information, we can utilize partially observed Markov (POMP) models. These models combine both latent and observable processes which are reliant on each other \cite{bhadra2011malaria}.\\

The application of POMP models spans various fields, including epidemiology, ecology  \cite{king2022markov}, and finance \cite{welton2005estimation} where they are used to infer the dynamics of complex systems from partial observations. For instance, in epidemiological studies, POMP models are instrumental in estimating disease transmission rates and predicting future outbreaks by modeling the spread of infectious diseases when only a fraction of cases are observed \cite{xue2020data}. Similarly, in ecology, these models help in understanding animal population dynamics and migration patterns based on limited or sporadic sighting data \cite{wood2001partially}.\\

Methodologically, POMP models are analyzed using a suite of computational techniques. Sequential Monte Carlo methods, such as particle filters, are commonly employed for state estimation and parameter inference in these models \cite{bhadra2011malaria}. These methods involve simulating a large number of potential realizations of the latent process and updating these simulations as new observations become available, thereby approximating the posterior distribution of the state given the data \cite{King2016}. Additionally, techniques like iterated filtering and particle Markov chain Monte Carlo are used for parameter estimation and model calibration.\\

With the random variables $Y_{0:N}$ which are the observable states process, the measurement model is $f_{Y_n|X_n}(y_n|x_n)$, and the collected data $y_{0:N}$ are observations of this process. We suppose random variables $X_{0:N}$ represent the latent states, where $X_0$ represents the initial state of the latent process, $f_{X|X_{n-1}}(x_n|x_{n-1})$. We assume that each $Y_n$ depends only upon the latent process at time $n$, $X_n$, and is conditionally independent of the other variables representing the measurement and latent processes, $Y_{n-1}$ and $X_{n-1}$, where $1 \leq n$. Together, $X_{0:N}$ and $Y_{0:N}$ form our POMP model. POMP models are vital in situations where we encounter incomplete or noisy measurements of a latent, unobserved Markov process. These models, also known as hidden Markov models or state space models, are particularly adept at handling time series data where observations are only partial or indirect representations of the underlying state of the system. Their strength lies in the combination of a latent process, which evolves according to Markov dynamics, and an observation process, which relates the latent states to the actual data observed.\\

\subsection{PanelPomp Framework}
In a panel data set containing K independent time series, K entities are represented by the symbols $1,2,...,K$, with N periodic observations collected for each unit $(t_1,t_2,...,t_{N})$. A stochastic observable process, denoted as $y_{k,1},...,y_{k,N}$, generates these measurements. PanelPOMP is an adaptation of the partially observable Markov process (POMP) tailored for panel data analysis. It entails developing a POMP model for each entity and employing the Panel Iterated Filtering (PIF) method to achieve the highest log-likelihood. Based on the POMP framework, a process and a measurement model were devised for every time series unit. The measurement model elucidates how the potential real population density generates observable time series data with the measurement density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$, while the process model portrays the actual population's dynamics through the one-step transition density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$. Ultimately, the joint density equals the initial density multiplied by the product of the process and measurement models for each state.\\

The PanelPOMP framework demonstrates a strong capability in handling ecological panel data due to its ability to model complex, time-varying processes within and across multiple ecological units \cite{Carles2020}. It efficiently integrates mechanistic models that capture underlying ecological dynamics with statistical inference methods, accommodating variability and correlations across units. This framework is adept at dealing with incomplete or noisy observational data, typical in ecological studies, providing robust estimates and predictions that are crucial for understanding and managing ecological systems.\\

\subsection{Likelihood of models and PIF}

In statistical inference challenges, likelihood frequently serves as the basis for estimating parameters and selecting models. Considering a model parameterized by vector $\theta$ within the $m$-dimensional parameter space $\Theta_m$, the likelihood function represents the combined probability density of the dataset, $y_{1:N}$, at $\theta$:
        \begin{equation}
        \mathcal{L}(\theta) = f_{Y_{1:N}}(y_{1:N}; \theta)
        \end{equation}
The goal is to get the estimation of $\theta$, $\hat{\theta}$ that maximizes the likelihood function $\mathcal{L}(\hat{\theta})$, or log-likelihood function $\mathbf{\ell}(\hat{\theta})$.

An ecological model's effectiveness in tracking disease spread hinges on its capability to predict cases or incidence rates. This capability is inherently tied to our trust in the model's precision in forecasting and our grasp of the temporal changes in latent states. As a result, we encounter two interrelated issues: ascertaining the distribution of $X_{n}$ at time $n$ based on $y_{1:n}$, and pinpointing parameter values, $\hat{\theta}$, that maximize the data likelihood. These processes are commonly respectively concluded as the filtering process and the inference process.\\

Working with POMP models can prove difficult due to the complexity of formulating both the likelihood function and transition density in intricate settings. To address this issue, several techniques have been devised, including the particle filter. Utilizing the particle filter method, it is only necessary to supply data, initial density simulators, single time-step transition density simulators for the latent process, and an evaluator for the observation process density reliant on the hidden process in order to achieve maximum likelihood estimates.\\

To implement the particle filter, we start by initializing $M$ particles at time 1, where $\{ X_1^m; m \in 1:M \}$. All with the required latent state detail and a vector of parameter values, $\theta$. As a result, for every time step n, where $n \in $ [$N$], we advance the each particle by doing the one-step random walk based on transition density, resulting in a set of particles that represents the prediction distribution at time n, which can be represented as $f_{X_n|X_{n-1}}(\cdot | X_{n-1}^m; \theta)$. Next, we assign weights to the particles based on the measurement density to obtain an ensemble that represents the filtering distribution at time n, where the new weight is calculated via $w_{n,m} = f_{Y_n|X_n}(y_n | x_n^m)$. Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \cite{5}. The particle filter is particularly useful in situations where both the likelihood function and transition density are difficult to calculate analytically.\\

As we know, the measurement process variables are mutually independent. In the mean time, the latent process states are dependent to one other. We also notice that the weight we assign to the particles will vary among different particles, which makes the calculation hard. As M to be large enough, we may approximate the weight of each particle to be the mean value of the to calculate the log-likelihood. Based on such fact, we can summarize as following:
         \begin{flalign}
          \mathcal{L}(\theta) &= f_{Y_{1:n}}(y_{1:N}; \theta) \\
         & = \prod_{n=1}^N f_{Y_{n} | Y_{1:n-1}}(y_n | y_{n-1}; \theta) \\
         & = \prod_{n=1}^N f_{Y_n | X_n}(y_n | x_n ; \theta)\\
         & \approx \prod_{n=1}^N \frac{1}{M} \sum_{m = 1}^M f_{Y_n|X_n}(y_n | x_n^m) 
         \end{flalign}
The particle filter, a form of Sequential Monte Carlo method, is instrumental in processing partially observed Markov processes within complex statistical models \cite{ionides2006inference}. This technique hinges on the generation of multiple, simulated "particles" that each represent a potential state of the latent process \cite{ionides2011iterated}. As new observational data becomes available, these particles are updated and assigned weights proportional to their likelihood of representing the true latent state, considering the observed data.\\

This weighting system is key to the particle filter's efficacy. It allows for the approximation of the distribution of the latent process at any given point in time, as well as the conditional likelihood of the observed data given the model. Essentially, the particle filter distills a complex, high-dimensional problem into a manageable computation by focusing on the most probable states of the latent process, as indicated by the observed data.\\

Panel Iterated Filtering (PIF) extends the improved iterated filtering algorithm, originally developed for time series data, to panel models \cite{Carles2020}. This method is particularly effective in exploring parameter spaces in complex, nonlinear, non-Gaussian models.  It does this by introducing random perturbations to the model parameters and utilizing a particle filter to explore the parameter space at each iteration. The convergence to a maximum likelihood estimate is achieved through procedures that iterate this search over the parameter space while diminishing the intensity of perturbations.\\

PIF's functionality is especially valuable in analyzing complex dynamical systems where the standard modeling approaches may fall short due to the intricacies of the data and the underlying processes. The ability of PIF to handle high-dimensional parameter spaces and its flexibility in managing different types of data structures make it a robust tool for statistical inference in complex models, particularly those encountered in ecological and epidemiological studies.\\

\section{Mechanistic models}
\label{sec:mecmod}
Mechanistic models are highly effective in understanding ecological dynamics due to their ability to incorporate detailed biological processes and interactions within ecosystems \cite{young1998data}. These models simulate the causal mechanisms underlying ecological phenomena, offering insights into complex ecological processes that are often difficult to observe directly. By integrating empirical data with theoretical principles, mechanistic models enable the prediction and analysis of ecological outcomes under varying conditions. This makes them invaluable tools for addressing challenges in ecology \cite{mouquet2015predictive}.\\

In mechanistic modeling of dynamic systems, the foundational step involves the formulation of differential or difference equations that encapsulate the temporal evolution of the system. This approach is pivotal in time series analysis using mechanistic models, which primarily aims to assess the model's adequacy in capturing the dynamics of the system and, upon validation, estimate unknown parameters that govern these dynamics \cite{breto2009time}. The complexity in such models often arises from intrinsic stochasticity, nonlinearity, and the presence of noisy and incomplete observational data, features that are particularly prevalent in biological sciences \cite{wilkinson2018stochastic}.\\

To analyze the dynamics of Parasitized and Non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \cite{weiss2013sir}, is utilized. This model integrates the aspects of food and parasites, unlike the classical SIR model, to more accurately reflect the interactions between native and invasive hosts in the context of resource availability and parasitic influence. The SIRPF model incorporates these factors as quantitative latent state variables, enabling a more comprehensive understanding of the system's dynamics. In the meantime, the ecological processes of Daphnia can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states. Utilizing panel data on Daphnia, the PanelPOMP framework becomes a valuable analytical tool \cite{Carles2020}. Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\

The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the following graph and model show:

\usetikzlibrary{positioning}
\usetikzlibrary {arrows.meta}
\usetikzlibrary{shapes.geometric}

\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1},
  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

\node (S) at (-0.5,-0.25) [square] {$S$};
\node (I) at (4.5,-0.25) [square] {$I$};
\node (P) at (2.5,0.75)  [rhombus] {$P$};
\node (R) at (3.5,2)[travel] {$R$};

\node (F) at (4.5,4.5) [report] {$F$};

\draw [->] (F) --  (R);
\draw [->] (P) --  (R);
\draw [->] (S) -- (R);
\draw [->] (I) -- (R);

\draw [->] (F) -- (S);
\draw [->] (F) -- (I);

\draw (S) edge[loop left] (S);
\draw (F) edge[loop above] (F);

\draw [->, bendy] (S) to  (I);
\draw [->, bendy] (I) to  (S);
\draw [->, bendy] (P) to (I);
% \path [->] (S) edge (I) 
%            (P) edge (I);
\draw [->, bendy] (I) to (P);
\end{tikzpicture}
}
\end{center}
\vspace{-5mm}
\caption{A flow diagram for the SIRPF model. Apart from the quantified states above, the graph includes $R$ state, which represents deadth status. For Daphnia either dentifera or lumholtzi, susceptible, infected Daphnia and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$. Some susceptible Daphnia get infected by parasites, transitioning from $S$ to $I$. Both susceptible and infected Daphnia contribute to the $R$ state upon death, releasing surviving parasites back into the environment. And the parasites continue to impact other individuals and get into another cycle. Food is consumed by both $S$ and $I$ individuals. Over time, elements in $F$ and $P$ also transit to the $R$ state.}
\label{fig:flow_SIRPF}
\end{figure}

{
\small
\begin{align}
&d S = r F\left(S+\xi I\right)dt - \theta_{S}S dt - p f_{S} S Pdt - \delta Sdt + S d\zeta_{S},\\
&d I = p f_{S} S Pdt - \theta_{I} Idt - \delta Idt + I d\zeta_{I},\\
&dF = \alpha F\left(1-\frac{F}{K_F}\right)dt - \gamma F \left(S+\xi I\right) dt - \delta F dt + \mu dt + F d\zeta_F,\\
&d P = \beta d IPdt +  - f_{S} S Pdt - f_{I} I Pdt - \theta_{\mathrm{p}} Pdt - \delta P dt + P d\zeta_{P},\\
&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}\right),\\
&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}\right).
\end{align}
}

Each equation intricately ties the state variables to depict the complex dynamics of the single species Parasitized system. Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race Parasitized system by introducing another Daphnia species. This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$)hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows. In the SIRPF2 model, the subscript n is used to denote parameters and variables related to the native host, i for those pertaining to the invasive host, and p for the parasite spores. The SIRPF2 model thus offers a more detailed representation of the ecological dynamics by accounting for the complex interplay between different species and environmental factors in a mechanistic framework.
\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
  bendy/.style={bend left=10},
  bendy2/.style={bend left=100},
  bendy3/.style={bend left=-100},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

\node (Sn) at (1,2) [square] {$S_n$};
\node (In) at (4,2) [square] {$I_n$};
\node (Si) at (1,-2) [square] {$S_i$};
\node (Ii) at (4,-2) [square] {$I_i$};
\node (R) at (2.5,0)  [rhombus] {$R$};
\node (P) at (5.5,0)[travel] {$P$};
\node (F) at (-0.5,0) [report] {$F$};

\draw [->, bendy] (Sn) to  (In);
\draw [->, bendy] (Si) to  (Ii);
\draw [->, bendy] (In) to  (Sn);
\draw [->, bendy] (Ii) to  (Si);

\draw [->] (F) --  (R);
\draw [->] (P) --  (R);
\draw [->] (Sn) -- (R);
\draw [->] (In) -- (R);
\draw [->] (Si) -- (R);
\draw [->] (Ii) -- (R);
\draw [->] (F) -- (Sn);
\draw [->] (F) -- (Si);
\draw (Sn) edge[loop below] (Sn);
\draw (Si) edge[loop above] (Si);
\draw (F) edge[loop left] (F);

\draw [->, bendy] (P) to (In);
\draw [->, bendy] (P) to (Ii);
\draw [->, bendy] (In) to (P);
\draw [->, bendy] (Ii) to (P);
\draw [->, bendy2] (F) to (In);
\draw [->, bendy3] (F) to (Ii);

\end{tikzpicture}
}
\end{center}
\vspace{-5mm}
\caption{A flow diagram for the SIRPF2 model. Apart from the quantified states above, the graph includes $R$ state, which represents deadth status. The flow diagram shows the how two species of Daphnia interact with each other indirectly through $F$ and $P$. The marginal processes have been shown in SIRPF mode, and SIRPF2 model is the adaption form in mixed-species dynamcis}
\label{fig:flow_SIRPF2}
\end{figure}

{
\small
\begin{align}
&d S_n = r_n F \left(S_n+\xi I_n\right)dt - \theta_{S_n}S_n dt - p_n f_{S_n} S_n Pdt - \delta S_ndt + S_n d\zeta_{S_n},\\
&d I_n = p_n f_{S_n} S_n Pdt - \theta_{I_n} I_ndt - \delta I_ndt + I_n d\zeta_{I_n},\\
&d S_i = r_i\left(S_i+\xi I_i\right)dt - \theta_{S_i} S_i dt - p_i f_{S_i} S_i Pdt - \delta S_idt + S_i d\zeta_{S_i},\\
&d I_i = p_i f_{S_i} S_i Pdt - \theta_{I_i} I_idt - \delta I_idt + I_i d\zeta_{I_i},\\
&dF = \alpha F\left(1-\frac{F}{K_F}\right)dt - \gamma_n F \left(S_n+\xi I_n\right) dt - \gamma_i F \left(S_i+\xi I_i\right) dt - \delta F dt + \mu dt + F d\zeta_F,\\
&d P = \beta_n d_n I_nPdt + \beta_i d_i I_idt - f_{S_n} S_n Pdt - f_{I_n} I_n Pdt - f_{S_i} S_i Pdt - f_{\mathrm{I}_i} I_i Pdt - \theta_{\mathrm{p}} Pdt - \delta P dt + P d\zeta_{P},\\
&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}\right),\\
&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}\right),\\
&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}\right),\\
&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}\right),\\
&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}\right),\\
&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}\right).
\end{align}
}

In the SIRPF2 model, detailed in Table 1, parameters and variables are distinctly annotated for native hosts (subscript $n$), invasive hosts (subscript $i$), and parasite spores (subscript $p$). Intraspecific competition among hosts is conceptualized via the Lotka-Volterra framework, with $r_n$ and $r_i$ denoting respective reproductive rates. Host density growth, influenced by algal density, is posited to be linear, indicating a direct relationship between food availability and Daphnia reproduction. Mortality rates of susceptible hosts are denoted as $\theta_{S_n}$ and $\theta_{S_i}$. The model encapsulates the dynamics of infection through \textit{Metschnikowia bicuspidata} intake, affecting susceptible hosts. This interaction consequently modulates reproductive rates, represented as $\xi r_n$. Infection rates correlate with spore density ($P$) and host ingestion rates ($f_{S_n}$, $f_{S_i}$). Upon death, infected hosts ($\theta_{I_n}$, $\theta_{I_i}$) release spores, which are subsequently subject to ingestion and environmental degradation. The consumption dynamics of food by susceptible and infected hosts are captured by $\gamma_n$ and $\gamma_i$, and food growth is inversely proportional to its carrying capacity ($K_F$), reflective of resource limitation. The model also incorporates a constant removal rate $\delta$, representing the loss of hosts, spores, and food due to external sampling. The constraint of food supply, indicated by a constant rate $\mu = 0.37 \times 10^5$ cells per day, is based on observed declines in Daphnia populations in the panel data, suggesting a pivotal role of food availability in driving population dynamics.\\

In the realm of ecological modeling, particularly in the context of the SIRPF2 model, the stochastic nature of ecological processes is adeptly captured using the principles of Brownian motion. This concept, fundamental to the theory of random particle movement in fluids, is extended in our model to encapsulate the inherent randomness in ecological dynamics. In the SIRPF2 model, Brownian motion is manifested through stochastic terms ($\zeta_F$, $\zeta_{S_n}$, $\zeta_{S_i}$, $\zeta_{I_n}$, $\zeta_{I_i}$, $\zeta_{F}$, $\zeta_P$), which follow a normal distribution with zero mean and specific standard deviations. These stochastic elements are crucial for modeling the unpredictable variances in the population dynamics and environmental interactions, especially in the face of unforeseen environmental perturbations or random demographic events.\\

The integration of these stochastic terms within the SIRPF2 framework, as detailed in Table 1, enhances the model's fidelity in simulating ecological interactions. The model, grounded in the Lotka-Volterra competition framework and augmented with stochastic components, provides a nuanced understanding of the interplay between native and invasive hosts, and their interaction with the surrounding resources and parasites. This incorporation of stochasticity, particularly through the lens of Brownian motion, is pivotal in transcending the limitations of deterministic models, thereby offering a more robust and comprehensive tool for ecological analysis.\\

\subsection{SIRPF adaption: SF and SF2 model}

The approach of analyzing Non-parasitized Dynamics through adapted models of the SIRPF framework, such as the Susceptible - Food (SF) and SF2 models, is a strategic method to quantify the impact of environmental resources, like food($F$), on the interactions between native($S_n$) and invasive($S_i$) host species. The SF model focuses on the dynamics between susceptible individuals and food resources, while the SF2 model extends this concept to include Mixed Race Dynamics. This extension is particularly useful in understanding how different species, native and invasive, interact with each other and with their environment in the absence of parasitic factors.\\

The adaptation of the SIRPF model to SF and SF2 as following allows for a more focused examination of the roles that resources play in ecological systems, especially when considering the competition or coexistence between different species. By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.\\

\begin{figure}
\begin{center}
%%%%% SEAIR diagram
\resizebox{8cm}{!}{
\begin{tikzpicture}[
  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
  bendy/.style={bend left=10},
  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
]
\tikzset{>={}}; % this is needed to implement the arrow redefinition

\node (Sn) at (-0.5,0) [square] {$S_n$};
\node (F) at (2.5,0) [report] {$F$};
\node (Si) at (5.5,0) [square] {$S_i$};
\node (R) at (2.5,2) [square] {$R$};

\draw [->] (F) --  (Sn);
\draw [->] (F) --  (Si);
\draw [->] (F) --  (R);
\draw [->] (Sn) --  (R);
\draw [->] (Si) --  (R);
\draw (Sn) edge[loop above] (Sn);
\draw (Si) edge[loop above] (Si);
\draw (F) edge[loop below] (F);

\draw [dotted, thick, red] (3.5, -0.75) rectangle (6.5, 1.5);
\end{tikzpicture}
}
\end{center}
\vspace{-5mm}
\caption{A flow diagram for the SF and SF2 model. In this system, both types of Daphnia and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$. As time progresses, individuals in $F$, $S_n$, and $S_i$ transition to the $R$ state, signifying mortality. While the whole graph show SF2 model, if we remove the node in the dashed rectangle, the graph will represent the SF model.}
\label{fig:SF&SF2}
\end{figure}

\begin{align}
&d S_n=r_n S_n dt- \theta_{S_n}S_n dt -\delta S_ndt + S_n d\zeta_{S_n}, \\
&d S_i=r_i S_i dt- \theta_{S_i} S_i dt -\delta S_idt + S_i d\zeta_{S_i}, \\
&dF=\alpha F\left(1-\frac{F}{K_F}\right)\cdot dt-\gamma_n F S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F \cdot dt+ \mu \cdot dt +F d \zeta_F.\\
&\zeta_F \sim N \left( 0, \sigma_F \right) \\
&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}  \right) \\
&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i} \right) 
\end{align}

As a adaption form of SIRPF model, the upper equations show SF2 model, with the meaning of each parameter is the same as the SIRPF model. The only difference is that the SF and SF2 model do not have the parasitic term. While if we only keep one species in SF2 model, the model will become SF model. Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of Daphnia, particularly under conditions of potential food scarcity. This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia. Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver. Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia. In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of Daphnia.\\

\subsection{Measurement and Initial condition}
In the ecological study of Daphnia populations, we employ the negative binomial distribution to address the inherent overdispersion observed in population counts. This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \cite{ning2021systemic}. In our model, the parameter $F$ represents the number of successful Daphnia reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of Daphnia per liter per sampling time in each trial. This framework aligns with the observed dynamics in Daphnia populations, where variability in reproduction rates, influenced by fluctuating food availability, leads to overdispersed population counts. The negative binomial distribution, therefore, provides a robust statistical tool to assess population dynamics, accommodating the irregularities in reproduction and survival patterns shaped by environmental constraints.\\

The model's effectiveness in capturing the Daphnia population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \cite{breto2011compound}. As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing Daphnia populations, particularly under varying food conditions.

\begin{align}
F \sim \operatorname{NBinomial}(n=S / \delta, \tau = k_F), \delta=0.013 
\end{align}

\section{Results}
\label{sec:res}
When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability. In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality. For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of Daphnia, and 8 for the rest dynamics.\\

Permitting parameters to be either shared or unique leads to varying dimensions for each model. With respective amount of units $d$ in the panel, adding a shared parameter increases the model's dimension by 1, and a specific parameter will add dimensions by $d$. We encountered a balance between fitting accuracy and simplicity when selecting models. Akaike information criterion (AIC) is a tool for model selection that balances model complexity against goodness of fit \cite{bozdogan1987model}. It is rooted in information theory, particularly the concept of entropy, and provides a relative measure of the information lost when a given model is used to represent the process that generated the data. A lower AIC value indicates a model with a better balance between fit and simplicity. In our study, we utilized AIC to compare models, ensuring adjustments were made for the number of parameters, thus facilitating a fair comparison across models with differing complexities.\\

The parameters are estimated via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models, which are presented in the Supplementary Material. Our analysis demonstrates that the 'All shared parameter' model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics. In cases like the Non-parasitized dynamics of single species, where this model's AIC is slightly higher (but within 2 points of the lowest AIC), it doesn't necessarily imply inferiority of model. AIC balances fit and complexity but doesn't consider stability or unseen data prediction. Here, Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \cite{nylund2007deciding}. This highlights the importance of looking beyond AIC for model assessment. As a result, in our case, the 'All shared parameter' model can be the best choice for all dynamics.\\

By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species. The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \cite{ma2021unified}. This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge. The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are 1.3 and 1.3 respectivly. While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to 0.13 for native Daphnia and to 0.13 for invasive species. In such scenario, we observe significant fluctuations in the population density of Daphnia. These oscillations are likely attributable to the constraints imposed by limited food resources within the environment. Such environmental limitations typically lead to variations in population dynamics. The observed variability in the survivability ratio can be understood within this assumption. At the same time, it is important to note that the relative survivability between the two Daphnia species studied remains consistent, despite an overall simultaneous decrease in their population densities. This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them. And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are 1.00 and 0.99 for single species dynamics, and 2.4, 0.35 for native and invasive population in mixed species dynamics respectively.\\

These findings underscore the importance of parameter sharing in ecological modeling when multiple species or units are involved. Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between Daphnia populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones. The results further reinforce the utility of the AIC in guiding the selection of models that not only provide a good fit to the data but also maintain a degree of simplicity, avoiding overfitting. The SIRPF model's performance with shared parameters points to a commonality in the factors affecting both native and invasive species, a result that could have implications for the management of these populations and their associated ecological interactions.\\

\section{Discussion}
\label{sec:dis}

The SIRPF model is a powerful tool for analyzing ecological data, as it can be used to estimate parameters and predict future population dynamics. The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia. The model's ability to estimate parameters and predict future population dynamics is a powerful tool for analyzing ecological data. The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia.\\

The SIRPF model extends the traditional SIR framework by including food and parasites, which is crucial in ecological models where these factors significantly influence host dynamics. This approach reflects the understanding that the dynamics of pathogens, including coinfections and community composition, are significantly influenced by ecological factors such as food availability and parasitic interactions. The complexity of these interactions is highlighted by studies emphasizing the role of community ecology in infectious disease research and the phylogenetic structure of host-parasite relationships. The SIRPF model's inclusion of these aspects offers a more comprehensive understanding of ecological dynamics, particularly in systems where traditional SIR models may fall short.\\

The PanelPOMP framework has shown remarkable efficacy in managing complex ecological systems \cite{Carles2020}, particularly when data points are limited and exhibit high volatility. Utilizing Panel Iterated Filtering, PanelPOMP adeptly fits data and estimates parameters that correspond to maximum log likelihood. This characteristic is crucial in ecological studies where panel data is prevalent, as it allows for more accurate modeling of complex, dynamic systems. However, the efficacy of PanelPOMP could be further enhanced with a larger dataset, as this would allow for a more restricted range of variations in simulations, leading to potentially more precise parameter estimations. This finding underscores the importance of robust data collection in ecological modeling and the value of PanelPOMP in analyzing such data.\\

The PanelPOMP model, while offering significant advantages in analyzing complex ecological data \cite{Carles2020}, does encounter certain limitations, especially when dealing with high-dimensional models like the SIRPF. One of the primary challenges associated with PanelPOMP is the increased computational resources and time required to achieve optimal log likelihood estimations. This is particularly pronounced in high-dimensional models due to the complex and intricate nature of the data and model structure. Thus, the model's effectiveness in these scenarios might necessitate complementary methodologies that are more attuned to the practical situations encountered at higher latitudes. Despite these challenges, the PanelPOMP model still represents a significant advancement in the field of ecological modeling, especially in its ability to handle partially observed Markov process models. The model's framework allows for a more nuanced understanding of complex ecological systems by facilitating likelihood-based inference for intricate and interrelated ecological processes.\\

\bibliography{reference}
\end{document}

### 2023-12-12 ms.Rnw
Update of paper

--- 2023-12-12 Paper.tex
@@
 \begin{document}
-\input{Paper-concordance}
+\SweaveOpts{concordance=TRUE}
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
   \title{\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
-  \author{Author 1\thanks{
-    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
-    Department of YYY, University of XXX\\
-    and \\
-    Author 2 \\
-    Department of ZZZ, University of WWW}
+\author{Bo Yang\\Department of Biostatistics, University of Michigan \And
+        Jesse Wheeler\\Department of Statistics, University of Michigan \And
+        Aaron A. King\\Department of Ecology and Evolutionary Biology \& \\ Center for the Study of Complex Systems, University of Michigan\\\And
+        Edward L. Ionides\\Department of Statistics, University of Michigan}
   \maketitle
@@
 } \fi
+
+<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+ library(plyr)
+ library(reshape2)
+ library(magrittr)
+ library(foreach)
+ library(readxl)
+ library(doParallel)
+ registerDoParallel(cores=36)
+ library(pomp)
+ library(panelPomp)
+ library(ggplot2)
+
+ myround <- function (x, digits = 1) {
+  if (length(digits) > 1) {
+    digits <- digits[1]
+    warning("Using only digits[1]")
+  }
+  if (digits < 1) {
+    as.character(round(x,digits))
+  } else {
+    tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
+    zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
+    tmp[tmp == paste0("-", zero)] <- zero
+    tmp
+  }
+}
+
+  mysignif <- function (x, digits = 1) {
+   myround(x, digits - ceiling(log10(abs(x))))
+  }
+
+ #---------------------------Host Daphnia----------------------------------------------------------------------
+ #----------------load data---------------------------------------------------------------
+load("/Users/ybb/Desktop/Daphnia-article/data/raw_data.RData")
+ #-----------------------------Host with para----------------------------------------------------------------------------
+ host_para <- Mesocosm_data[101:180, ]
+ host_para <- subset(host_para, select = c(rep, day, dent.adult,dent.inf ))
+ host_para <- host_para[80: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ host_para$day = (host_para$day - 1) * 5 + 7
+ host_para_data = list()
+ trails = c("K","L","M","N","O","P","Q","S")
+ for (i in 1: length(trails)){
+   host_para_data[[i]]=subset(host_para, select = c("day", "dent.adult","dent.inf"),
+                              host_para$rep == trails[i])
+ }
+
+ host_para_plot <- ggplot(host_para,aes(x=day,y=dentadult))+
+   geom_line(data=host_para,aes(x=day,y=dent.adult,color = "Susceptible")) +
+   geom_line(data=host_para,aes(x=day,y=dent.inf,color = "Infected")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia dentifera adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Host without para----------------------------------------------------------------------------
+ host_nopara <- Mesocosm_data[1:100, ]
+ host_nopara <- subset(host_nopara, select = c(rep, day, dent.adult))
+ host_nopara <- host_nopara[100: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ host_nopara$day = (host_nopara$day - 1) * 5 + 7
+ host_nopara_data = list()
+ trails = c("A","B","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   host_nopara_data[[i]]=subset(host_nopara, select = c("day", "dent.adult"),
+                                host_nopara$rep == trails[i])
+ }
+
+ host_nopara_plot <- ggplot(host_nopara,aes(x=day,y=dentadult))+
+   geom_line(data=host_nopara,aes(x=day,y=dent.adult)) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia dentifera adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Invasive with para----------------------------------------------------------------------------
+ invasive_para <- Mesocosm_data_invasive[101:190, ]
+ invasive_para <- subset(invasive_para, select = c(rep, day, lum.adult,lum.adult.inf))
+ invasive_para <- invasive_para[90: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ invasive_para$day = (invasive_para$day - 1) * 5 + 7
+ invasive_para_data = list()
+ trails = c("L","M","N","O","P","Q","R","S","T")
+ for (i in 1: length(trails)){
+   invasive_para_data[[i]]=subset(invasive_para, select = c("day", "lum.adult","lum.adult.inf"),
+                                  invasive_para$rep == trails[i])
+ }
+
+ invasive_para_plot <- ggplot(invasive_para,aes(x=day,y=lumadult))+
+   geom_line(data=invasive_para,aes(x=day,y=lum.adult,color = "Susceptible")) +
+   geom_line(data=invasive_para,aes(x=day,y=lum.adult.inf,color = "Infected")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia lumholtzi adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Invasive without para----------------------------------------------------------------------------
+ invasive_nopara <- Mesocosm_data_invasive[1:100, ]
+ invasive_nopara <- subset(invasive_nopara, select = c(rep, day, lum.adult))
+ invasive_nopara <- invasive_nopara[100: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ invasive_nopara$day = (invasive_nopara$day - 1) * 5 + 7
+ invasive_nopara_data = list()
+ trails = c("A","B","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   invasive_nopara_data[[i]]=subset(invasive_nopara, select = c("day", "lum.adult"),
+                                    invasive_nopara$rep == trails[i])
+ }
+
+ invasive_nopara_plot <- ggplot(invasive_nopara,aes(x=day,y=lumadult))+
+   geom_line(data=invasive_nopara,aes(x=day,y=lum.adult)) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia lumholtzi adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Mix with para----------------------------------------------------------------------------
+ mixed_para <- Mesocosm_data_mixed[91:170, ]
+ mixed_para <- subset(mixed_para, select = c(rep, day, dent.adult,dent.inf,lum.adult,lum.adult.inf))
+ mixed_para <- mixed_para[80: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ mixed_para$day = (mixed_para$day - 1) * 5 + 7
+ mixed_para_data = list()
+ trails = c("K","L","M","N","O","P","Q","S")
+ for (i in 1: length(trails)){
+   mixed_para_data[[i]]=subset(mixed_para, select = c("day", "dent.adult","dent.inf","lum.adult","lum.adult.inf"),
+                               mixed_para$rep == trails[i])
+ }
+
+ mixed_para_plot <- ggplot()+
+   geom_line(data=mixed_para,aes(x=day,y=lum.adult,color = "Susceptible lumholtzi"))+
+   geom_line(data=mixed_para,aes(x=day,y=lum.adult.inf,color = "Infected lumholtzi")) +
+   geom_line(data=mixed_para,aes(x=day,y=dent.adult,color = "Susceptible dentifera")) +
+   geom_line(data=mixed_para,aes(x=day,y=dent.inf,color = "Infected dentifera")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Mix without para----------------------------------------------------------------------------
+ mixed_nopara <- Mesocosm_data_mixed[1:90, ]
+ mixed_nopara <- subset(mixed_nopara, select = c(rep, day, dent.adult,lum.adult))
+ mixed_nopara <- mixed_nopara[90: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ mixed_nopara$day = (mixed_nopara$day - 1) * 5 + 7
+ mixed_nopara_data = list()
+ trails = c("A","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   mixed_nopara_data[[i]]=subset(mixed_nopara, select = c("day", "dent.adult","lum.adult"),
+                                 mixed_nopara$rep == trails[i])
+ }
+
+ mixed_nopara_plot <- ggplot()+
+   geom_line(data=mixed_nopara,aes(x=day,y=lum.adult,color = "Lumholtzi"))+
+   geom_line(data=mixed_nopara,aes(x=day,y=dent.adult,color = "Dentifera")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+ #----------Only lum Model-----------------------------------------------------------------------------------------------
+ invasive_nopara_rpro <- Csnippet("
+                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
+                      double noiB, noiG ;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiB = rnorm(0, sigB * sqrt(dt));
+                      noiG = rnorm(0, sigG * sqrt(dt));
+
+                      //------------F-------------
+
+                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
+                      Fpred =  S * F * (Beta * dt);
+                      Fspl = F * (delta * dt);
+                      Ffill = 0.37 * dt;
+
+                      //------------S--------------
+                      Sgrow = S * F * (Theta * dt) + S * noiG;
+                      Sdeath = S * (gamma * dt);
+                      Sspl = S * (delta * dt);
+
+                      //M is the total Daphnia sample out
+                      M += Sspl;
+                      F += Fbirth - Fpred + Ffill - Fspl;
+                      S += Sgrow - Sdeath  - Sspl;
+
+                      if (S <= 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F <= 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      ")
+ invasive_nopara_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     M = 0;
+                     error_count = 0;
+                     ")
+ invasive_nopara_dmeas = Csnippet("
+                 double delta = 0.013;
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = log(tol);
+                 } else {
+                   lik = dpois(lumadult,M,give_log);
+                 }
+                 ")
+ invasive_nopara_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 lumadult = rpois(M);
+                 ")
+ invasive_nopara_pt <- parameter_trans(
+   log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
+ )
+ invasive_nopara_pomplist=list()
+ #Not sure about the consumption rate of alga
+ invasive_nopara_parameters = c(1, 0.0002500899,40,2, 10,0.2496332878,0.3)
+ names(invasive_nopara_parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")
+
+ for (i in 1:10){
+   colnames(invasive_nopara_data[[i]]) <- c('day', 'lumadult')
+   pomp(data = invasive_nopara_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(invasive_nopara_rpro,delta.t=1/4),
+        rinit = invasive_nopara_init,
+        dmeasure = invasive_nopara_dmeas,
+        rmeasure = invasive_nopara_rmeas,
+        partrans = invasive_nopara_pt,
+        obsnames = "lumadult",
+        accumvars = c("error_count","M"),
+        paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"),
+        statenames = c("S", "error_count", "F","M")
+   ) -> invasive_nopara_pomplist[[i]]
+   coef(invasive_nopara_pomplist[[i]])=invasive_nopara_parameters
+ }
+ names(invasive_nopara_pomplist)=paste("u", 1:10,sep = "")
+
+ invasive_nopara_shared_parameter =        c(0.0002500899 ,40   ,10    ,0.2496332878,  1, 2)
+ names(invasive_nopara_shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")
+
+ invasive_nopara_specific_mat = matrix(data = c(rep(0.3, 10)),
+                                       nrow = 1, nc = 10, byrow = T,
+                                       dim = list(rownames = c("sigG"),
+                                                  colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))
+
+ invasive_nopara_panelfood = panelPomp(invasive_nopara_pomplist, shared=invasive_nopara_shared_parameter,specific = invasive_nopara_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------Only dent Model-----------------------------------------------------------------------------------------------
+ host_nopara_rpro <- Csnippet("
+                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
+                      double noiB, noiG ;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiB = rnorm(0, sigB * sqrt(dt));
+                      noiG = rnorm(0, sigG * sqrt(dt));
+
+                      //------------F-------------
+
+                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
+                      Fpred =  S * F * (Beta * dt);
+                      Fspl = F * (delta * dt);
+                      Ffill = 0.37 * dt;
+
+                      //------------S--------------
+                      Sgrow = S * F * (Theta * dt) + S * noiG;
+                      Sdeath = S * (gamma * dt);
+                      Sspl = S * (delta * dt);
+
+                      //M is the total Daphnia sample out
+                      M += Sspl;
+                      F += Fbirth - Fpred + Ffill - Fspl;
+                      S += Sgrow - Sdeath  - Sspl;
+
+                      if (S <= 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F <= 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      ")
+ # Initial state. Assume t0 = day 4
+ host_nopara_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     M = 0;
+                     error_count = 0;
+                     ")
+
+ host_nopara_dmeas = Csnippet("
+                 double delta = 0.013;
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = log(tol);
+                 } else {
+                   lik = dpois(dentadult,M,give_log);
+                 }
+                 ")
+
+ host_nopara_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 dentadult = rpois(M);
+                 ")
+
+ host_nopara_pt <- parameter_trans(
+   log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
+ )
+
+ host_nopara_pomplist=list()
+
+ #Not sure about the consumption rate of alga
+ host_nopara_parameters = c(0.5, 0.0005,0.03,0.1, 8,0.6,0.3)
+ names(host_nopara_parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")
+
+ for (i in 1:10){
+   colnames(host_nopara_data[[i]]) <- c('day', 'dentadult')
+   pomp(data = host_nopara_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(host_nopara_rpro,delta.t=1/4),
+        rinit = host_nopara_init,
+        dmeasure = host_nopara_dmeas,
+        rmeasure = host_nopara_rmeas,
+        partrans = host_nopara_pt,
+        obsnames = "dentadult",
+        accumvars = c("error_count","M"),
+        paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"),
+        statenames = c("S", "error_count", "F","M")
+   ) -> host_nopara_pomplist[[i]]
+   coef(host_nopara_pomplist[[i]])=host_nopara_parameters
+ }
+ names(host_nopara_pomplist)=paste("u", 1:10,sep = "")
+
+ host_nopara_shared_parameter =        c(7.773061e-05  ,2.052644e-01   ,1.790036    ,6.940058e-03  ,1.170719e-01, 0.15)
+ names(host_nopara_shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")
+
+ host_nopara_specific_mat = matrix(data = c(rep(0.3, 10)),
+                                   nrow = 1, nc = 10, byrow = T,
+                                   dim = list(rownames = c("sigG"),
+                                              colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))
+
+ host_nopara_panelfood = panelPomp(host_nopara_pomplist, shared=host_nopara_shared_parameter,specific = host_nopara_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------dent par Model----------------------------------------------
+ dent_par_rpro <- Csnippet("
+                      double S_term, I_term, F_term, P_term;
+                      double noiS, noiI, noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiS = rnorm(0, sigS * sqrt(dt));
+                      noiI = rnorm(0, sigI * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------s-------------
+                      //Model without death part
+                      S_term = r  * F * (S + zeta * I) * dt - theta_S * S * dt - prob * f_S * S * P * dt - delta * S * dt + S * noiS;
+
+                      //------------I--------------
+
+                      I_term = prob * f_S * S * P * dt - theta_I * I *dt - delta * I * dt + I * noiI;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamma * F * (S + zeta * I) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = beta * theta_I * I * dt - f_S * S * P * dt - f_I * I * P * dt - theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      S += S_term;
+                      I += I_term;
+                      P += P_term;
+
+                      T_D += delta * S * dt;
+                      T_I += delta * I * dt;
+
+                      //Trace time
+
+                     if (t - 4.0 < 0.000000001 && t - 4.0 > -0.000000001){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                      if (S < 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (I < 0) {
+                        I = 0;
+                        error_count += 0.001;
+                      }
+                      if (P < 0 && t > 3.9) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      ")
+
+ # Initial state. Assume t0 = day 4
+ dent_par_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     T_D = 0;
+                     T_I = 0;
+                     I = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ dent_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                  if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(dentadult,k_S,T_D,give_log) +  dnbinom_mu(dentinf,k_I,T_I,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(dentadult,k_S,T_D,give_log) *  dnbinom_mu(dentinf,k_I,T_I,give_log);
+                    }
+                      }
+
+                 ")
+
+ dent_par_rmeas = Csnippet("
+                 dentadult = rnbinom_mu(k_S,T_D);
+                 dentinf = rnbinom_mu(k_I,T_I);
+                 ")
+
+ dent_par_pt <- parameter_trans(
+   log = c("alpha", "beta", "gamma", "sigS", "sigI", "sigF","sigP","K_F","f_S","theta_P","r", "f_I","k_S","k_I"),
+   logit = c('prob',"zeta","theta_S","theta_I")
+ )
+
+ dent_par_pomplist=list()
+
+ dent_par_parameters =         c(0.16,   2.8   ,   0.00011,   0.1,    0.1,   0.7,  0.2,     0.01,     0.8,     2.3,  0.001,   0.2,      0.1,  0.06 ,  0.7,   0.003,10,10)
+ names(dent_par_parameters) =  c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I")
+
+ for (i in 1:8){
+   colnames(host_para_data[[i]]) <- c('day', 'dentadult', 'dentinf')
+   pomp(data = host_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(dent_par_rpro,delta.t=1/4),
+        rinit = dent_par_init,
+        dmeasure = dent_par_dmeas,
+        rmeasure = dent_par_rmeas,
+        partrans = dent_par_pt,
+        obsnames = c("dentadult", "dentinf"),
+        accumvars = c("error_count","T_D","T_I"),
+        paramnames = c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I"),
+        statenames = c("S",  "I", "error_count", "F", "T_D","T_I","P")
+   ) -> dent_par_pomplist[[i]]
+   coef(dent_par_pomplist[[i]])=dent_par_parameters
+ }
+ names(dent_par_pomplist)=paste("u", 1:8,sep = "")
+
+ shared_dent_par_parameter =        c(0.0115,   0.39 , 9.628e-06,  0.5447 , 0.1004, 0.04401,    0.3,   0.5798,  0.1489,  0.0008096, 0.004265,    0.1432,  0.003244 ,  0.6551,   0.003584,   8,   8)
+ names(shared_dent_par_parameter) = c("alpha", "sigS",  "gamma",  "sigI",  "sigF", "sigP", "theta_S","theta_I","K_F",      "f_S",  "theta_P",     "r",    "zeta",   "prob",     "f_I","k_S","k_I")
+
+ specific_dent_par_mat = matrix(data = c(rep(2.2,8)),
+                                nrow = 1, nc = 8, byrow = T,
+                                dim = list(rownames = "beta",
+                                           colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8")))
+
+ dent_par_panelfood = panelPomp(dent_par_pomplist, shared=shared_dent_par_parameter,specific = specific_dent_par_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------lum par Model----------------------------------------------
+ lum_par_rpro <- Csnippet("
+                      double S_term, I_term, F_term, P_term;
+                      double noiS, noiI, noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiS = rnorm(0, sigS * sqrt(dt));
+                      noiI = rnorm(0, sigI * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------s-------------
+                      //Model without death part
+                      S_term = r  * F * (S + zeta * I) * dt - theta_S * S * dt - prob * f_S * S * P * dt - delta * S * dt + S * noiS;
+
+                      //------------I--------------
+
+                      I_term = prob * f_S * S * P * dt - theta_I * I *dt - delta * I * dt + I * noiI;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamma * F * (S + zeta * I) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = beta * theta_I * I * dt - f_S * S * P * dt - f_I * I * P * dt - theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      S += S_term;
+                      I += I_term;
+                      P += P_term;
+
+                      if (S < 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (I < 0) {
+                        I = 0;
+                        error_count += 0.001;
+                      }
+                      if (P < 0) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      T_D += delta * S * dt;
+                      T_I += delta * I * dt;
+
+                      //Trace time
+
+                     if (t - 4.00 < 0.1 && t - 4.00 > -0.1){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                     // Rprintf(\"P=%f\\n\",P);
+
+                      ")
+
+ lum_par_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     T_D = 0;
+                     T_I = 0;
+                     I = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ lum_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                  if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(lumadult,k_S,T_D,give_log) +  dnbinom_mu(luminf,k_I,T_I,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(lumadult,k_S,T_D,give_log) *  dnbinom_mu(luminf,k_I,T_I,give_log);
+                    }
+                      }
+
+                 ")
+
+ lum_par_rmeas = Csnippet("
+                 lumadult = rnbinom_mu(k_S,T_D);
+                 luminf = rnbinom_mu(k_I,T_I);
+                 ")
+
+ lum_par_pt <- parameter_trans(
+   log = c("alpha", "beta", "gamma", "sigS", "sigI", "sigF","sigP","K_F","f_S","theta_P","r", "f_I","k_S","k_I"),
+   logit = c('prob',"zeta","theta_S","theta_I")
+ )
+
+ lum_par_pomplist=list()
+
+ lum_par_parameters =         c(0.16,   2.8   ,   0.00011,   0.1,    0.1,   0.7,  0.2,     0.01,     0.8,     2.3,  0.001,   0.2,      0.1,  0.06 ,  0.7,   0.003,10,10)
+ names(lum_par_parameters) =  c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I")
+
+ for (i in 1:9){
+   colnames(invasive_para_data[[i]]) <- c('day', 'lumadult', 'luminf')
+   pomp(data = invasive_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(lum_par_rpro,delta.t=1/4),
+        rinit = lum_par_init,
+        dmeasure = lum_par_dmeas,
+        rmeasure = lum_par_rmeas,
+        partrans = lum_par_pt,
+        obsnames = c("lumadult", "luminf"),
+        accumvars = c("error_count","T_D","T_I"),
+        paramnames = c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I"),
+        statenames = c("S",  "I", "error_count", "F", "T_D","T_I","P")
+   ) -> lum_par_pomplist[[i]]
+   coef(lum_par_pomplist[[i]])=lum_par_parameters
+ }
+ names(lum_par_pomplist)=paste("u", 1:9,sep = "")
+
+ lum_par_shared_parameter =        c(0.0115,   0.39 , 9.628e-06,  0.5447 , 0.1004, 0.04401,    0.3,   0.5798,  0.1489,  0.0008096, 0.004265,    0.1432,  0.003244 ,  0.6551,   0.003584,   8,   8)
+ names(lum_par_shared_parameter) = c("alpha", "sigS",  "gamma",  "sigI",  "sigF", "sigP", "theta_S","theta_I","K_F",      "f_S",  "theta_P",     "r",    "zeta",   "prob",     "f_I","k_S","k_I")
+
+ lum_par_specific_mat = matrix(data = c(rep(2.2,9)),
+                               nrow = 1, nc = 9, byrow = T,
+                               dim = list(rownames = "beta",
+                                          colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9")))
+
+ lum_par_panelfood = panelPomp(lum_par_pomplist, shared=lum_par_shared_parameter,specific = lum_par_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+ #----------mixed par Model-----------------------------------------------------------------------------------------------
+ mix_par_rpro <- Csnippet("
+                      double Sn_term, In_term, F_term, P_term , Si_term, Ii_term;
+                      double noiSn, noiIn, noiSi , noiIi ,noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiSn = rnorm(0, sigSn * sqrt(dt));
+                      noiIn = rnorm(0, sigIn * sqrt(dt));
+                      noiSi = rnorm(0, sigSi * sqrt(dt));
+                      noiIi = rnorm(0, sigIi * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------Sn-------------
+                      Sn_term = rn  * F * (Sn + xi * In) * dt - theta_Sn * Sn * dt -  probn * f_Sn * Sn * P * dt - delta * Sn * dt + Sn * noiSn;
+
+                      //------------In--------------
+
+                      In_term = probn * f_Sn * Sn * P * dt - theta_In * In *dt - delta * In * dt + In * noiIn;
+
+                      //------------Si-------------
+                      Si_term = ri  * F * (Si + xi * Ii) * dt - theta_Si * Si * dt -  probi * f_Si * Si * P * dt - delta * Si * dt + Si * noiSi;
+
+                      //------------Ii--------------
+
+                      Ii_term = probi * f_Si * Si * P * dt - theta_Ii * Ii *dt - delta * Ii * dt + Ii * noiIi;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamman * F * (Sn + xi * In) * dt - gammai * F * (Si + xi * Ii) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = betan * theta_In * In * dt + betai * theta_Ii * Ii * dt - f_Sn * Sn * P * dt - f_In * In * P * dt - f_Si * Si * P * dt - f_Ii * Ii * P * dt- theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      Sn += Sn_term;
+                      In += In_term;
+                      Si += Si_term;
+                      Ii += Ii_term;
+                      P += P_term;
+
+                      //Trace time
+
+                     if (t - 4.0 < 0.001 && t - 4.0 > -0.001){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                      if (Sn < 0.0) {
+                        Sn = 0.0;
+                        error_count += 1;
+                      }
+                      if (Si < 0.0) {
+                        Si = 0.0;
+                        error_count += 1000000;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (In < 0) {
+                        In = 0;
+                        error_count += 0.001;
+                      }
+                      if (Ii < 0) {
+                        Ii = 0;
+                        error_count += 0.000000001;
+                      }
+                      if (P < 0 && t > 3.9) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      T_Sn += delta * Sn * dt;
+                      T_In += delta * In * dt;
+                      T_Si += delta * Si * dt;
+                      T_Ii += delta * Ii * dt;
+
+                      ")
+
+ # Initial state. Assume t0 = day 4
+ mix_par_init = Csnippet("
+                     Sn = 2.333; //2.3333 = 35/15L
+                     Si = 0.667; //0.667 = 10/15L
+                     F = 16.667;
+                     T_Sn = 0;
+                     T_Si = 0;
+                     T_In = 0;
+                     T_Ii = 0;
+                     In = 0;
+                     Ii = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ mix_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(lumadult,k_Si,T_Si,give_log) +  dnbinom_mu(luminf,k_Ii,T_Ii,give_log) + dnbinom_mu(dentadult,k_Sn,T_Sn,give_log) +  dnbinom_mu(dentinf,k_In,T_In,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(lumadult,k_Si,T_Si,give_log) *  dnbinom_mu(luminf,k_Ii,T_Ii,give_log) * dnbinom_mu(dentadult,k_Sn,T_Sn,give_log) *  dnbinom_mu(dentinf,k_In,T_In,give_log);
+                    }
+                      }
+
+                 ")
+
+ mix_par_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 dentadult = rnbinom_mu(k_Sn,T_Sn);
+                 dentinf = rnbinom_mu(k_In,T_In);
+                 lumadult = rnbinom_mu(k_Si,T_Si);
+                 luminf = rnbinom_mu(k_Ii,T_Ii);
+                 ")
+
+ mix_par_pt <- parameter_trans(
+   #Without death part
+   log = c("alpha", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","k_Ii","k_In","k_Sn","k_Si"),
+   logit = c("probn","probi","xi","theta_Sn","theta_In","theta_Si","theta_Ii","theta_P")
+ )
+
+ mix_par_pomplist=list()
+
+ mix_par_parameters =         c(0.16,    0.3,   17,    3  ,    0.0001,  8.34e-06,     0.4,    0.5,    0.8,     0.2,     0.01,  0.2,    0.15,       0.07 ,   0.2,       0.01,    0.096     ,1  ,0.0006,0.001 , 0.0035 ,0.003 , 0.2  , 0.4, 0.6  ,0.8 ,8,8,8,8)
+ names(mix_par_parameters) =  c("alpha","xi", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","theta_Sn","theta_In","theta_Si","theta_P","theta_Ii","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","probn","probi","k_Ii","k_In","k_Sn","k_Si")
+
+ for (i in 1:8){
+   colnames(mixed_para_data[[i]]) <- c('day', 'dentadult', 'dentinf','lumadult','luminf')
+   pomp(data = mixed_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(mix_par_rpro,delta.t=1/4),
+        rinit = mix_par_init,
+        dmeasure = mix_par_dmeas,
+        rmeasure = mix_par_rmeas,
+        partrans = mix_par_pt,
+        obsnames = c("dentadult", "dentinf","lumadult","luminf"),
+        accumvars = c("error_count","T_Sn","T_In","T_Si","T_Ii"),
+        paramnames = c("alpha","xi", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","theta_Sn","theta_In","theta_Si","theta_P","theta_Ii","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","probn","probi","k_Ii","k_In","k_Sn","k_Si"),
+        statenames = c("Sn",  "In", "Si",  "Ii","error_count", "F", "T_Sn","T_In","T_Si","T_Ii","P")
+   ) -> mix_par_pomplist[[i]]
+   coef(mix_par_pomplist[[i]])=mix_par_parameters
+ }
+ names(mix_par_pomplist)=paste("u", 1:8,sep = "")
+
+ mix_par_shared_parameter = c("alpha" = 0.045,"xi" = 0.0005,"gamman" = 9.4e-05,"gammai" =0.00075, "sigSn" = 0.15, "sigIn" = 0.033, "sigSi" =0.28, "sigIi" = 0.03,
+                              "sigF" = 0.08, "theta_Sn" = 0.19,"theta_In" = 0.62,"theta_Si" = 0.22,"theta_P" = 0.08,"theta_Ii" = 0.48,"K_F" = 0.3,"f_Sn" = 0.001,"f_Si" = 0.0011,"f_In" = 0.0012,"f_Ii" = 0.00023,"rn" = 0.25,
+                              "ri" = 0.29,"probn" = 0.93,"probi" = 0.96,"k_Ii" = 0.38,"k_In" = 0.83,"k_Sn" = 11,"k_Si"= 570,"sigP" = 0.02, "betan" = 2, "betai" = 2)
+
+ mix_par_panelfood = panelPomp(mix_par_pomplist, shared=mix_par_shared_parameter)
+@
 
@@
 Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task.
-As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \citep{Levin_1992}.\\
+As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \citep{levin1992problem}.\\
 
 Panel data, sometimes referred to as longitudinal data, is composed of multiple time series.
-Each of these time series, which may be multivariate, contains a series of observations collected from a separate entity  \cite{Carles2020}.
+Each of these time series, which may be multivariate, contains a series of observations collected from a separate entity  \citep{Carles2020}.
 In ecology, panel data is a traditional data form that researchers gather to mitigate the impact of random variations throughout experiments, so that it can more accurately measure changes in experimental objectives and control the variables needed to achieve experimental goals.
 Due to the non-linear characteristics of ecological data, the analysis of panel data entails a high-dimensional structure.
-Such features diminish the explanatory power of traditional time series analysis and certain Monte Carlo inference techniques \cite{johnstone2009statistical}.
+Such features diminish the explanatory power of traditional time series analysis and certain Monte Carlo inference techniques \citep{johnstone2009statistical}.
 The goal of this study is to investigate the practicality of employing mechanistic models on ecological panel data using the panelPOMP structure.\\
 
-The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \cite{King2016}.
+The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \citep{King2016}.
 This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering.
-Stemming from iterated filtering \cite{King2016} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood.
+Stemming from iterated filtering \citep{King2016} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood.
 This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.\\
@@
 In this research, we will show the advancement of the panelPOMP model by analyzing the Daphnia panel data generated in the experiment of Searle et al.
-\cite{Searle2016}.
+\citep{Searle2016}.
 The research group of Searle et al.
-collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \cite{Searle2016}.
+collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \citep{Searle2016}.
 In their lab experiment, they generated standard ecological panel data featuring various independent trials under the same experimental conditions, with each consisting of a brief, nonlinear, and nonstationary time series.
 Conventional mechanistic theories cannot directly interpret these observations.
-Utilizing the panelPOMP approach allows for the combination of theoretical models and empirical data, enabling the extraction of the comprehensive information found in each individual time series \cite{Carles2020}.\\
+Utilizing the panelPOMP approach allows for the combination of theoretical models and empirical data, enabling the extraction of the comprehensive information found in each individual time series \citep{Carles2020}.\\
 
@@
 It is worth noting that the population density of alga is neither observable nor recorded in the experiment.
-That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes \cite{Edwards2012}.
-At the same time, the panelPOMP framework provides an estimation of the parameters in the model, which helps to understand the hidden mechanisms in the ecological system \cite{Carles2020}.
+That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes \citep{Edwards2012}.
+At the same time, the panelPOMP framework provides an estimation of the parameters in the model, which helps to understand the hidden mechanisms in the ecological system \citep{Carles2020}.
 Based on the estimation, a detailed interpretation of relationship between invasive host and native host with presence of parasite is explaned.\\
@@
 
-The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two Daphnia species: Daphnia \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \cite{searle2016salinization}, and Daphnia \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
+The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two Daphnia species: Daphnia \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \citep{searle2016salinization}, and Daphnia \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
 D.
-lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \cite{havel1993daphnia}.
+lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \citep{havel1993daphnia}.
 These Daphnia species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines.
-For their experiments, six clones of each species were selected to eliminate selection bias \cite{Searle2016}.\\
+For their experiments, six clones of each species were selected to eliminate selection bias \citep{Searle2016}.\\
 
 The dynamics also involves \textit{Metschnikowia bicuspidata}, a parasitic fungus predominantly infecting the native host.
-This parasite, ingested as spores by Daphnia, aggressively affects the host's fecundity and longevity \cite{duffy2010temporal}.
+This parasite, ingested as spores by Daphnia, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
 The parasite's presence is identifiable by the opaqueness it causes in normally transparent hosts, with spore release upon host death.
 Searle's study utilized a M.
-bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \cite{Searle2016}.
+bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \citep{Searle2016}.
 The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for Daphnia.
-\textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \cite{arnold1971ingestion}.
-Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of Daphnia species can be closely monitored and manipulated \cite{martinez1994effect}.\\
+\textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \citep{arnold1971ingestion}.
+Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of Daphnia species can be closely monitored and manipulated \citep{martinez1994effect}.\\
 
-Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two Daphnia host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \cite{Searle2016}.
+Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two Daphnia host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \citep{Searle2016}.
 This observation suggests a potential competitive advantage for the invasive species.
@@
 
-In the Non-parasitized level, indoor mesocosms were used \cite{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media.
+In the Non-parasitized level, indoor mesocosms were used \citep{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media.
 The experimental design involved 10 replicates for single-species treatments, and 9 replicates for mixed-species treatment.
@@
 Algal and nutrient levels were supplemented bi-weekly, with COMBO media added as necessary.
-The experiment concluded after 52 days, totaling 10 sampling sessions \cite{Searle2016}.\\
+The experiment concluded after 52 days, totaling 10 sampling sessions \citep{Searle2016}.\\
 
@@
 \begin{subfigure}[b]{\linewidth}
-\includegraphics{Paper-002}
+<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=  
+   mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
+     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
+   }))
+
+ ggplot(mixed_nopara_data_combined, aes(x=day)) + 
+   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
+   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
+   scale_x_continuous(limits = c(0, 52)) +
+   scale_y_continuous(limits = c(0, 175)) +
+   facet_grid(. ~ group, scales = "free", space = "free") +
+   theme(
+     strip.text = element_text(size=10),
+     axis.text.x=element_text(color="black"),
+     axis.ticks.x=element_line(color="black"),
+     axis.text.y=element_text(color="black"),
+     axis.ticks.y=element_line(color="black"),
+     axis.title.x=element_blank(),
+     axis.title.y=element_blank(),
+     legend.position="none",
+     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
+     strip.background = element_blank(), # Optional: Remove facet label background
+     strip.placement = "outside"
+   ) +
+   labs(title="Daphnia density in Non-parasitized Dynamics") +
+   guides(colour=guide_legend(title="Species"))
+@
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
@@
 \begin{subfigure}[b]{\linewidth}
-\includegraphics{Paper-003}
+<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=  
+   mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
+     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
+   }))
+
+ ggplot(mixed_para_data_combined, aes(x=day)) + 
+   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
+   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
+    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
+   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +
+
+   scale_x_continuous(limits = c(0, 52)) +
+   scale_y_continuous(limits = c(0, 175)) +
+   facet_grid(. ~ group, scales = "free", space = "free") +
+   theme(
+     strip.text = element_text(size=10),
+     axis.text.x=element_text(color="black"),
+     axis.ticks.x=element_line(color="black"),
+     axis.text.y=element_text(color="black"),
+     axis.ticks.y=element_line(color="black"),
+     axis.title.x=element_blank(),
+     axis.title.y=element_blank(),
+     legend.position="none",
+     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
+     strip.background = element_blank(), # Optional: Remove facet label background
+     strip.placement = "outside"
+   ) +
+   labs(title="Daphnia density in Parasitized Dynamics") +
+   guides(colour=guide_legend(title="Species"))
+@
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
@@
 To address these issues of missing information, we can utilize partially observed Markov (POMP) models.
-These models combine both latent and observable processes which are reliant on each other \cite{bhadra2011malaria}.\\
+These models combine both latent and observable processes which are reliant on each other \citep{bhadra2011malaria}.\\
 
-The application of POMP models spans various fields, including epidemiology, ecology  \cite{king2022markov}, and finance \cite{welton2005estimation} where they are used to infer the dynamics of complex systems from partial observations.
-For instance, in epidemiological studies, POMP models are instrumental in estimating disease transmission rates and predicting future outbreaks by modeling the spread of infectious diseases when only a fraction of cases are observed \cite{xue2020data}.
-Similarly, in ecology, these models help in understanding animal population dynamics and migration patterns based on limited or sporadic sighting data \cite{wood2001partially}.\\
+The application of POMP models spans various fields, including epidemiology, ecology  \citep{king2022markov}, and finance \citep{welton2005estimation} where they are used to infer the dynamics of complex systems from partial observations.
+For instance, in epidemiological studies, POMP models are instrumental in estimating disease transmission rates and predicting future outbreaks by modeling the spread of infectious diseases when only a fraction of cases are observed \citep{xue2020data}.
+Similarly, in ecology, these models help in understanding animal population dynamics and migration patterns based on limited or sporadic sighting data \citep{wood2001partially}.\\
 
 Methodologically, POMP models are analyzed using a suite of computational techniques.
-Sequential Monte Carlo methods, such as particle filters, are commonly employed for state estimation and parameter inference in these models \cite{bhadra2011malaria}.
-These methods involve simulating a large number of potential realizations of the latent process and updating these simulations as new observations become available, thereby approximating the posterior distribution of the state given the data \cite{King2016}.
+Sequential Monte Carlo methods, such as particle filters, are commonly employed for state estimation and parameter inference in these models \citep{bhadra2011malaria}.
+These methods involve simulating a large number of potential realizations of the latent process and updating these simulations as new observations become available, thereby approximating the posterior distribution of the state given the data \citep{King2016}.
 Additionally, techniques like iterated filtering and particle Markov chain Monte Carlo are used for parameter estimation and model calibration.\\
@@
 
-The PanelPOMP framework demonstrates a strong capability in handling ecological panel data due to its ability to model complex, time-varying processes within and across multiple ecological units \cite{Carles2020}.
+The PanelPOMP framework demonstrates a strong capability in handling ecological panel data due to its ability to model complex, time-varying processes within and across multiple ecological units \citep{Carles2020}.
 It efficiently integrates mechanistic models that capture underlying ecological dynamics with statistical inference methods, accommodating variability and correlations across units.
@@
 Next, we assign weights to the particles based on the measurement density to obtain an ensemble that represents the filtering distribution at time n, where the new weight is calculated via $w_{n,m} = f_{Y_n|X_n}(y_n | x_n^m)$.
-Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \cite{5}.
+Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \citep{5}.
 The particle filter is particularly useful in situations where both the likelihood function and transition density are difficult to calculate analytically.\\
@@
          \end{flalign}
-The particle filter, a form of Sequential Monte Carlo method, is instrumental in processing partially observed Markov processes within complex statistical models \cite{ionides2006inference}.
-This technique hinges on the generation of multiple, simulated "particles" that each represent a potential state of the latent process \cite{ionides2011iterated}.
+The particle filter, a form of Sequential Monte Carlo method, is instrumental in processing partially observed Markov processes within complex statistical models \citep{ionides2006inference}.
+This technique hinges on the generation of multiple, simulated "particles" that each represent a potential state of the latent process \citep{ionides2011iterated}.
 As new observational data becomes available, these particles are updated and assigned weights proportional to their likelihood of representing the true latent state, considering the observed data.\\
@@
 
-Panel Iterated Filtering (PIF) extends the improved iterated filtering algorithm, originally developed for time series data, to panel models \cite{Carles2020}.
+Panel Iterated Filtering (PIF) extends the improved iterated filtering algorithm, originally developed for time series data, to panel models \citep{Carles2020}.
 This method is particularly effective in exploring parameter spaces in complex, nonlinear, non-Gaussian models.
@@
 \label{sec:mecmod}
-Mechanistic models are highly effective in understanding ecological dynamics due to their ability to incorporate detailed biological processes and interactions within ecosystems \cite{young1998data}.
+Mechanistic models are highly effective in understanding ecological dynamics due to their ability to incorporate detailed biological processes and interactions within ecosystems \citep{young1998data}.
 These models simulate the causal mechanisms underlying ecological phenomena, offering insights into complex ecological processes that are often difficult to observe directly.
 By integrating empirical data with theoretical principles, mechanistic models enable the prediction and analysis of ecological outcomes under varying conditions.
-This makes them invaluable tools for addressing challenges in ecology \cite{mouquet2015predictive}.\\
+This makes them invaluable tools for addressing challenges in ecology \citep{mouquet2015predictive}.\\
 
 In mechanistic modeling of dynamic systems, the foundational step involves the formulation of differential or difference equations that encapsulate the temporal evolution of the system.
-This approach is pivotal in time series analysis using mechanistic models, which primarily aims to assess the model's adequacy in capturing the dynamics of the system and, upon validation, estimate unknown parameters that govern these dynamics \cite{breto2009time}.
-The complexity in such models often arises from intrinsic stochasticity, nonlinearity, and the presence of noisy and incomplete observational data, features that are particularly prevalent in biological sciences \cite{wilkinson2018stochastic}.\\
+This approach is pivotal in time series analysis using mechanistic models, which primarily aims to assess the model's adequacy in capturing the dynamics of the system and, upon validation, estimate unknown parameters that govern these dynamics \citep{breto2009time}.
+The complexity in such models often arises from intrinsic stochasticity, nonlinearity, and the presence of noisy and incomplete observational data, features that are particularly prevalent in biological sciences \citep{wilkinson2018stochastic}.\\
 
-To analyze the dynamics of Parasitized and Non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \cite{weiss2013sir}, is utilized.
+To analyze the dynamics of Parasitized and Non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \citep{weiss2013sir}, is utilized.
 This model integrates the aspects of food and parasites, unlike the classical SIR model, to more accurately reflect the interactions between native and invasive hosts in the context of resource availability and parasitic influence.
@@
 In the meantime, the ecological processes of Daphnia can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
-Utilizing panel data on Daphnia, the PanelPOMP framework becomes a valuable analytical tool \cite{Carles2020}.
+Utilizing panel data on Daphnia, the PanelPOMP framework becomes a valuable analytical tool \citep{Carles2020}.
 Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
@@
 \begin{align}
-&d S = r F\left(S+\xi I\right)dt - \theta_{S}S dt - p f_{S} S Pdt - \delta Sdt + S d\zeta_{S},\\
-&d I = p f_{S} S Pdt - \theta_{I} Idt - \delta Idt + I d\zeta_{I},\\
-&dF = \alpha F\left(1-\frac{F}{K_F}\right)dt - \gamma F \left(S+\xi I\right) dt - \delta F dt + \mu dt + F d\zeta_F,\\
-&d P = \beta d IPdt +  - f_{S} S Pdt - f_{I} I Pdt - \theta_{\mathrm{p}} Pdt - \delta P dt + P d\zeta_{P},\\
-&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}\right),\\
-&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}\right),\\
-&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}\right),\\
-&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}\right).
+&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S}(t),\\
+&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}(t),\\
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F(t),\\
+&d P(t) = \beta d I(t)dt +  - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}(t),\\
+&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
+&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}dt\right),\\
+&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
+&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
 \end{align}
@@
 \begin{align}
-&d S_n = r_n F \left(S_n+\xi I_n\right)dt - \theta_{S_n}S_n dt - p_n f_{S_n} S_n Pdt - \delta S_ndt + S_n d\zeta_{S_n},\\
-&d I_n = p_n f_{S_n} S_n Pdt - \theta_{I_n} I_ndt - \delta I_ndt + I_n d\zeta_{I_n},\\
-&d S_i = r_i\left(S_i+\xi I_i\right)dt - \theta_{S_i} S_i dt - p_i f_{S_i} S_i Pdt - \delta S_idt + S_i d\zeta_{S_i},\\
-&d I_i = p_i f_{S_i} S_i Pdt - \theta_{I_i} I_idt - \delta I_idt + I_i d\zeta_{I_i},\\
-&dF = \alpha F\left(1-\frac{F}{K_F}\right)dt - \gamma_n F \left(S_n+\xi I_n\right) dt - \gamma_i F \left(S_i+\xi I_i\right) dt - \delta F dt + \mu dt + F d\zeta_F,\\
-&d P = \beta_n d_n I_nPdt + \beta_i d_i I_idt - f_{S_n} S_n Pdt - f_{I_n} I_n Pdt - f_{S_i} S_i Pdt - f_{\mathrm{I}_i} I_i Pdt - \theta_{\mathrm{p}} Pdt - \delta P dt + P d\zeta_{P},\\
-&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}\right),\\
-&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}\right),\\
-&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}\right),\\
-&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}\right),\\
-&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}\right),\\
-&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}\right).
+\begin{split}
+&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\\
+&\quad \quad \quad\delta S_n(t)dt + S_n(t) d\zeta_{S_n}(t),\\
+\end{split}
+\\
+&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n}(t),\\
+\begin{split}
+&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt - \\
+&\quad \quad \quad \delta S_i(t)dt + S_i(t) d\zeta_{S_i}(t),\\
+\end{split}
+\\
+&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
+\begin{split}
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \\
+&\quad \quad \quad\gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F(t),\\
+\end{split}
+\\
+\begin{split}
+&d P(t) = \beta_n d_n I_n(t)dt + \beta_i d_i I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
+&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}(t),\\
+\end{split}
+\\
+&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}dt\right),\\
+&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}dt\right),\\
+&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}dt\right),\\
+&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}dt\right),\\
+&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
+&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
 \end{align}
@@
 
+\renewcommand{\arraystretch}{0.5}
+\begin{table}[H]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
+ \hline
+Parameter & Definition &Unit\\[0.5ex]
+\hline\hline
+$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
+$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
+$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
+$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
+$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
+$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
+$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
+$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
+$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
+$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
+$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
+$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
+$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
+$\delta$ & Sampling rate&$day ^{-1} $ \\
+$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
+$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
+$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
+$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
+$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
+$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
+$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
+$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
+$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
+$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
+$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
+$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
+$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
+$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
+$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
+$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
+$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
+ \hline
+\end{tabular}}
+\caption{This table shows the units of the parameters of competition SIRPF2 model}
+\label{Table:1}
+\end{table}
+
+\renewcommand{\arraystretch}{1}
 In the realm of ecological modeling, particularly in the context of the SIRPF2 model, the stochastic nature of ecological processes is adeptly captured using the principles of Brownian motion.
@@
 In the ecological study of Daphnia populations, we employ the negative binomial distribution to address the inherent overdispersion observed in population counts.
-This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \cite{ning2021systemic}.
+This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \citep{ning2021systemic}.
 In our model, the parameter $F$ represents the number of successful Daphnia reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of Daphnia per liter per sampling time in each trial.
@@
 
-The model's effectiveness in capturing the Daphnia population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \cite{breto2011compound}.
+The model's effectiveness in capturing the Daphnia population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
 As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing Daphnia populations, particularly under varying food conditions.
@@
 \label{sec:res}
+<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+ load('SIRPF_model_estimate.RData')
+
+ ratio_SF_i = SF_invasive_estimate[length(SF_invasive_estimate)]/SF_invasive_estimate[4]
+ ratio_SF_n = ratio_SF_i
+ ratio_SF2_i = SF2_estimate[11]/SF2_estimate[8]
+ ratio_SF2_n = SF2_estimate[10]/SF2_estimate[7]
+ ratio_SIRPF_is = SIRPF_invasive_estimate[13]/SIRPF_invasive_estimate[8]
+ ratio_SIRPF_ii = SIRPF_invasive_estimate[13] * SIRPF_invasive_estimate[14]/SIRPF_invasive_estimate[9]
+ ratio_SIRPF_ns = SIRPF_native_estimate[13]/SIRPF_native_estimate[8]
+ ratio_SIRPF_ni = SIRPF_native_estimate[13] * SIRPF_native_estimate[14]/SIRPF_native_estimate[9]
+ ratio_SIRPF2_ns = SIRPF2_estimate[20]/SIRPF2_estimate[10]
+ ratio_SIRPF2_ni = SIRPF2_estimate[20] *  SIRPF2_estimate[2]/SIRPF2_estimate[12]
+ ratio_SIRPF2_is = SIRPF2_estimate[21]/SIRPF2_estimate[11]
+ ratio_SIRPF2_ii = SIRPF2_estimate[21] *  SIRPF2_estimate[2]/SIRPF2_estimate[14]
+@
 When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability.
@@
 We encountered a balance between fitting accuracy and simplicity when selecting models.
-Akaike information criterion (AIC) is a tool for model selection that balances model complexity against goodness of fit \cite{bozdogan1987model}.
+Akaike information criterion (AIC) is a tool for model selection that balances model complexity against goodness of fit \citep{bozdogan1987model}.
 It is rooted in information theory, particularly the concept of entropy, and provides a relative measure of the information lost when a given model is used to represent the process that generated the data.
@@
 AIC balances fit and complexity but doesn't consider stability or unseen data prediction.
-Here, Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \cite{nylund2007deciding}.
+Here, Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \citep{nylund2007deciding}.
 This highlights the importance of looking beyond AIC for model assessment.
@@
 By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species.
-The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \cite{ma2021unified}.
+The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}.
 This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.
-The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are 1.3 and 1.3 respectivly.
-While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to 0.13 for native Daphnia and to 0.13 for invasive species.
+The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,2)} and \Sexpr{mysignif(ratio_SF_i,2)} respectivly.
+While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,2)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,2)} for invasive species.
 In such scenario, we observe significant fluctuations in the population density of Daphnia.
@@
 This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them.
-And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are 1.00 and 0.99 for single species dynamics, and 2.4, 0.35 for native and invasive population in mixed species dynamics respectively.\\
+And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,2)} and \Sexpr{mysignif(ratio_SIRPF_is,2)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,2)}, \Sexpr{mysignif(ratio_SIRPF2_is,2)} for native and invasive population in mixed species dynamics respectively.\\
 
@@
 
-The PanelPOMP framework has shown remarkable efficacy in managing complex ecological systems \cite{Carles2020}, particularly when data points are limited and exhibit high volatility.
+The PanelPOMP framework has shown remarkable efficacy in managing complex ecological systems \citep{Carles2020}, particularly when data points are limited and exhibit high volatility.
 Utilizing Panel Iterated Filtering, PanelPOMP adeptly fits data and estimates parameters that correspond to maximum log likelihood.
@@
 
-The PanelPOMP model, while offering significant advantages in analyzing complex ecological data \cite{Carles2020}, does encounter certain limitations, especially when dealing with high-dimensional models like the SIRPF.
+The PanelPOMP model, while offering significant advantages in analyzing complex ecological data \citep{Carles2020}, does encounter certain limitations, especially when dealing with high-dimensional models like the SIRPF.
 One of the primary challenges associated with PanelPOMP is the increased computational resources and time required to achieve optimal log likelihood estimations.

### 2023-12-12 ms.tex
Update of paper

--- 2023-12-12 Paper.tex
@@
 \begin{document}
-\input{Paper-concordance}
+\input{ms-concordance}
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 \begin{subfigure}[b]{\linewidth}
-\includegraphics{Paper-002}
+\includegraphics{ms-002}
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
@@
 \begin{subfigure}[b]{\linewidth}
-\includegraphics{Paper-003}
+\includegraphics{ms-003}
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
@@
 By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species.
-The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \cite{ma2021unified}.
+The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}.
 This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.

### 2023-12-12 si/si.Rnw
Update of paper

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear,numbers]{natbib}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyphenat}
\tolerance=1000
\usepackage{url}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{hyphenat}
\usepackage{Sweave}
\bibliographystyle{apalike}
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\begin{document}
\SweaveOpts{concordance=TRUE}
\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\if1\blind
{
  \title{\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
  \author{Author 1\thanks{
    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
    Department of YYY, University of XXX\\
    and \\
    Author 2 \\
    Department of ZZZ, University of WWW}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
\end{center}
  \medskip
} \fi

\bigskip
<<Setup, include=FALSE,echo=FALSE,results=hide>>=
library(dplyr)
library(ggplot2)
@

\section{Simulation}
\label{sec:sim}
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
       x = "Day",
       y = "Daphnia dentifera adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
       x = "Day",
       y = "Daphnia lumholtzi adult density")
@
\end{subfigure}%
\caption{Plot of Simulation of Non-Parasitize Dyanmics with SF2 Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
       x = "Day",
       y = "Susceptible Daphnia dentifera adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
       x = "Day",
       y = "Susceptible Daphnia lumholtzi adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = iymini, ymax = iymaxi, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
       x = "Day",
       y = "Infected Daphnia lumholtzi adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = iymin, ymax = iymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
       x = "Day",
       y = "Infected Daphnia dentifera adult density")
@
\end{subfigure}%
\caption{Plot of Simulation of Parasitize Dyanmics with SIRPF2 Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
       x = "Day",
       y = "Susceptible Daphnia lumholtzi adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
       x = "Day",
       y = "Infected Daphnia lumholtzi adult density")
@
\end{subfigure}%
\caption{Plot of Simulation of Invasive Parasitize Dyanmics with SIRPF Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
       x = "Day",
       y = "Susceptible Daphnia dentifera adult density")
@
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
ggplot() +
  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
       x = "Day",
       y = "Infected Daphnia dentifera adult density")
@
\end{subfigure}%
\caption{Plot of Simulation of Native Parasitize Dyanmics with SIRPF Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
       x = "Day",
       y = "Daphnia dentifera adult density")
@
\caption{Plot of Simulation of Native Non-Parasitize Dyanmics with SF Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
ggplot() +
  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
  geom_line() +
  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
  facet_wrap(~unit, scales = "free_y") +
  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
  theme_minimal() +
  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
       x = "Day",
       y = "Infected Daphnia dentifera adult density")
@
\caption{Plot of Simulation of Invsaive Non-Parasitize Dyanmics with SF Model. The blue shadow implies the possible range from simulation, while the red line shows the real-world data. The various trail represents different buckets in the experiment}
\end{figure}

\newpage
\section{Tables}
<<Setup2, include=FALSE,echo=FALSE,results=hide>>=
library(dplyr)
library(ggplot2)
 myround <- function (x, digits = 1) {
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  if (digits < 1) {
    as.character(round(x,digits))
  } else {
    tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
    zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
    tmp[tmp == paste0("-", zero)] <- zero
    tmp
  }
}

  mysignif <- function (x, digits = 1) {
   myround(x, digits - ceiling(log10(abs(x))))
  }

# SF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2_ll.RData")
# SFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN_ll.RData")
# SFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_invasive/SFI_ll.RData")
# SIRPF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2_ll.RData")
# SIRPFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN_ll.RData")
# SIRPFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI_ll.RData")
load("/Users/ybb/Desktop/Daphnia-article/SIRPF_model_estimate.RData")
@
\renewcommand{\arraystretch}{0.5}
\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
 \hline
Parameter & Definition &Unit\\[0.5ex]
\hline\hline
$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$&Variable\\
$F$& Alga density& $10^5 cells \cdot L ^{-1}$&Variable\\
$P$ & Spore density & $10^3 spores \cdot L ^{-1}$&Variable\\
$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[15],2)}\\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[20],2)}\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[21],2)}\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[11],2)} \\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[14],2)} \\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[12],2)} \\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[10],2)} \\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[13],2)} \\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[22],2)}\\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[23],2)}\\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[16],2)}\\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[17],2)}\\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[18],2)}\\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[19],2)} \\
$\delta$ & Sampling rate&$day ^{-1} $& constant = 0.013\\
$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[1],2)}\\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[3],2)}\\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[4],2)}\\
$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & constant = 0.37\\
$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(SIRPF2_estimate[2],2)}\\
$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[30],2)}\\
$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[29],2)}\\
$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[5],2)}\\
$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[7],2)}\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[6],2)}\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[8],2)}\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[9],2)}\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[28],2)}\\
$k_{S_n}$ & Overdispersion parameter& $Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[26],2)}\\
$k_{S_i}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[27],2)}\\
$k_{I_n}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[25],2)}\\
$k_{I_i}$ & Overdispersion parameter&$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[24],2)}\\[1ex]
 \hline
\end{tabular}}
\caption{This table shows the units of the parameters of competition SIRPF2 model}
\label{Table:1}
\end{table}
\renewcommand{\arraystretch}{1}

\bibliography{reference}
\end{document}

### 2023-12-12 si/si.tex
Update of paper

--- 2023-12-12 si/si.Rnw
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{si-concordance}
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 \bigskip
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
-library(dplyr)
-library(ggplot2)
-@
 
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
-       x = "Day",
-       y = "Daphnia dentifera adult density")
-@
+\includegraphics{si-002}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
-       x = "Day",
-       y = "Daphnia lumholtzi adult density")
-@
+\includegraphics{si-003}
 \end{subfigure}%
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Susceptible Daphnia dentifera adult density")
-@
+\includegraphics{si-004}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Susceptible Daphnia lumholtzi adult density")
-@
+\includegraphics{si-005}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = iymini, ymax = iymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Infected Daphnia lumholtzi adult density")
-@
+\includegraphics{si-006}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = iymin, ymax = iymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
+\includegraphics{si-007}
 \end{subfigure}%
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Susceptible Daphnia lumholtzi adult density")
-@
+\includegraphics{si-008}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Infected Daphnia lumholtzi adult density")
-@
+\includegraphics{si-009}
 \end{subfigure}%
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Susceptible Daphnia dentifera adult density")
-@
+\includegraphics{si-010}
 \end{subfigure}
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
+\includegraphics{si-011}
 \end{subfigure}%
@@
 \centering
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
-       x = "Day",
-       y = "Daphnia dentifera adult density")
-@
+\includegraphics{si-012}
 \caption{Plot of Simulation of Native Non-Parasitize Dyanmics with SF Model.
@@
 \centering
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
+\includegraphics{si-013}
 \caption{Plot of Simulation of Invsaive Non-Parasitize Dyanmics with SF Model.
@@
 \section{Tables}
-<<Setup2, include=FALSE,echo=FALSE,results=hide>>=
-library(dplyr)
-library(ggplot2)
- myround <- function (x, digits = 1) {
-  if (length(digits) > 1) {
-    digits <- digits[1]
-    warning("Using only digits[1]")
-  }
-  if (digits < 1) {
-    as.character(round(x,digits))
-  } else {
-    tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
-    zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
-    tmp[tmp == paste0("-", zero)] <- zero
-    tmp
-  }
-}
-
-  mysignif <- function (x, digits = 1) {
-   myround(x, digits - ceiling(log10(abs(x))))
-  }
-
-# SF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2_ll.RData")
-# SFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN_ll.RData")
-# SFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_invasive/SFI_ll.RData")
-# SIRPF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2_ll.RData")
-# SIRPFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN_ll.RData")
-# SIRPFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI_ll.RData")
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_model_estimate.RData")
-@
 \renewcommand{\arraystretch}{0.5}
@@
 $P$ & Spore density & $10^3 spores \cdot L ^{-1}$&Variable\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[15],2)}\\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[20],2)}\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[21],2)}\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[11],2)} \\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[14],2)} \\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[12],2)} \\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[10],2)} \\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[13],2)} \\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[22],2)}\\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[23],2)}\\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[16],2)}\\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[17],2)}\\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[18],2)}\\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[19],2)} \\
+$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $& 0.092\\
+$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$& 0.23\\
+$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$&0.23\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$&0.66 \\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$&0.37 \\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$&0.086 \\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$& 0.098 \\
+$\theta_{P}$ & Spore degradation rate&$ day^{-1}$&0.015 \\
+$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $& 0.92\\
+$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $& 0.98\\
+$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.00027\\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.00027\\
+$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.0018\\
+$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.000067 \\
 $\delta$ & Sampling rate&$day ^{-1} $& constant = 0.013\\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[1],2)}\\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[3],2)}\\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[4],2)}\\
+$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ & 0.011\\
+$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.000030\\
+$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.00067\\
 $\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & constant = 0.37\\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(SIRPF2_estimate[2],2)}\\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[30],2)}\\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[29],2)}\\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[5],2)}\\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[7],2)}\\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[6],2)}\\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[8],2)}\\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[9],2)}\\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[28],2)}\\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[26],2)}\\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[27],2)}\\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[25],2)}\\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[24],2)}\\[1ex]
+$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & 0.00041\\
+$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ &3.8\\
+$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ &3.2\\
+$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & 0.061\\
+$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & 0.11\\
+$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & 0.45\\
+$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & 0.0018\\
+$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & 0.29\\
+$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & 0.20\\
+$k_{S_n}$ & Overdispersion parameter& $Unitless$ &4.5\\
+$k_{S_i}$ &Overdispersion parameter &$Unitless$ &5.4\\
+$k_{I_n}$ &Overdispersion parameter &$Unitless$ &1.4\\
+$k_{I_i}$ & Overdispersion parameter&$Unitless$ &1.2\\[1ex]
  \hline

### 2023-12-12 ms.Rnw
Update

@@
 \bibliographystyle{apalike}
-\newcommand{\blind}{0}
+\newcommand{\blind}{1}
 
@@
   \title{\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
-\author{Bo Yang\\Department of Biostatistics, University of Michigan \And
-        Jesse Wheeler\\Department of Statistics, University of Michigan \And
-        Aaron A. King\\Department of Ecology and Evolutionary Biology \& \\ Center for the Study of Complex Systems, University of Michigan\\\And
+\author{Bo Yang\\Department of Biostatistics, University of Michigan \\
+        Jesse Wheeler\\Department of Statistics, University of Michigan \\
+        Aaron A. King\\Department of Ecology and Evolutionary Biology \& \\ Center for the Study of Complex Systems, University of Michigan\\
         Edward L. Ionides\\Department of Statistics, University of Michigan}

### 2023-12-12 ms.Rnw
Update

@@
 \begin{align}
-&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S}(t),\\
-&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}(t),\\
-&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F(t),\\
-&d P(t) = \beta d I(t)dt +  - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}(t),\\
+&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S},\\
+&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I},\\
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
+&d P(t) = \beta d I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
 &d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
@@
 \begin{align}
-\begin{split}
-&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\\
-&\quad \quad \quad\delta S_n(t)dt + S_n(t) d\zeta_{S_n}(t),\\
-\end{split}
-\\
-&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n}(t),\\
-\begin{split}
-&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt - \\
-&\quad \quad \quad \delta S_i(t)dt + S_i(t) d\zeta_{S_i}(t),\\
-\end{split}
-\\
+% \begin{split}
+&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n},\\
+% \end{split}
+% \\
+&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n},\\
+% \begin{split}
+&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i},\\
+% \end{split}
+% \\
 &d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
 \begin{split}
-&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \\
-&\quad \quad \quad\gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F(t),\\
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
+&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
 \end{split}
@@
 &d P(t) = \beta_n d_n I_n(t)dt + \beta_i d_i I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
-&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}(t),\\
+&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
 \end{split}
@@
 \begin{align}
-&d S_n=r_n S_n dt- \theta_{S_n}S_n dt -\delta S_ndt + S_n d\zeta_{S_n}, \\
-&d S_i=r_i S_i dt- \theta_{S_i} S_i dt -\delta S_idt + S_i d\zeta_{S_i}, \\
-&dF=\alpha F\left(1-\frac{F}{K_F}\right)\cdot dt-\gamma_n F S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F \cdot dt+ \mu \cdot dt +F d \zeta_F.\\
-&\zeta_F \sim N \left( 0, \sigma_F \right) \\
-&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}  \right) \\
-&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i} \right) 
+&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
+&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
+&dF=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
+&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
+&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
+&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
 \end{align}

### 2024-01-24 ms.Rnw
Update of the ms file

@@
   square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
-  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1},
+  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1,inner sep=0cm},
   travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
@@
 
+<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+ load('all_shared.RData')
+
+parms = mif.estimate
+parms = as.data.frame(t(parms))
+
+@
+
 \renewcommand{\arraystretch}{0.5}
@@
 \centering
-\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
+\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
  \hline
-Parameter & Definition &Unit\\[0.5ex]
+Parameter & Definition &Unit & Value\\[0.5ex]
 \hline\hline
-$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
-$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
-$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
-$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$\delta$ & Sampling rate&$day ^{-1} $ \\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
+$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
+$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$& Variable\\
+$F$& Alga density& $10^5 cells \cdot L ^{-1}$& Variable\\
+$P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
+$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
+$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
+$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
+$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
+$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
+$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
+$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
+$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
+$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
+$\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
+$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
+$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
+$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
+$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
+$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(parms$xi,2)}\\
+$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betai,2)}\\
+$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betan,2)}\\
+$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
+$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
+$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
+$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
+$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
+$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
+$k_{S_n}$ & Overdispersion parameter& $Unitless$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
+$k_{S_i}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_Si,2)}\\
+$k_{I_n}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_In,2)}\\
+$k_{I_i}$ & Overdispersion parameter&$Unitless$ & \Sexpr{mysignif(parms$k_Ii,2)}\\[1ex]
  \hline
@@
 The adaptation of the SIRPF model to SF and SF2 as following allows for a more focused examination of the roles that resources play in ecological systems, especially when considering the competition or coexistence between different species.
-By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.\\
+By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.
 

### 2024-01-24 ms.tex
Update of the ms file

@@
 \centering
-\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
+\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
  \hline
-Parameter & Definition &Unit\\[0.5ex]
+Parameter & Definition &Unit & Value\\[0.5ex]
 \hline\hline
-$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
-$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
-$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
-$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$\delta$ & Sampling rate&$day ^{-1} $ \\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
+$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
+$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$& Variable\\
+$F$& Alga density& $10^5 cells \cdot L ^{-1}$& Variable\\
+$P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
+$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & 0.092\\
+$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & 0.23\\
+$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & 0.23\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & 0.66\\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & 0.37\\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & 0.086\\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & 0.098\\
+$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & 0.015\\
+$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & 0.92\\
+$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & 0.98\\
+$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & 0.00027\\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.00027 \\
+$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& 0.0018 \\
+$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & 0.000067\\
+$\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
+$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& 0.011 \\
+$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.000030\\
+$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & 0.00067\\
+$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
+$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & 0.00041\\
+$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ & 3.8\\
+$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & 3.2\\
+$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & 0.061\\
+$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & 0.11\\
+$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & 0.45\\
+$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & 0.0018\\
+$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & 0.29\\
+$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & 0.20\\
+$k_{S_n}$ & Overdispersion parameter& $Unitless$ & 4.5\\
+$k_{S_i}$ &Overdispersion parameter &$Unitless$ & 5.4\\
+$k_{I_n}$ &Overdispersion parameter &$Unitless$ & 1.4\\
+$k_{I_i}$ & Overdispersion parameter&$Unitless$ & 1.2\\[1ex]
  \hline

### 2024-01-29 ms.Rnw

@@
 \usepackage{hyphenat}
-\usepackage{Sweave}
+%\usepackage{Sweave}
 \bibliographystyle{apalike}
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Setup,include=FALSE,echo=FALSE,results="hide">>=
  library(plyr)
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=  
+<<echo=FALSE, fig.cap="Plot1", fig.align="center", fig.width=14, fig.height=4>>=  
    mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
@@
 \begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=  
+<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig.width=14, fig.height=4>>=  
    mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
@@
 
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Setup-params,include=FALSE,echo=FALSE,results="hide">>=
  load('all_shared.RData')
@@
 \label{sec:res}
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Setup-results, include=FALSE,echo=FALSE,results="hide">>=
  load('SIRPF_model_estimate.RData')

### 2024-01-31 ms.tex
modify Rnw and tex

--- 2023-12-12 ms.tex
@@
 } \fi
-
-\bigskip
-\begin{abstract}
-In ecological research, panel data, often comprising multi-magnitude time series, presents unique analytical challenges due to its high dimensionality.
-This study employs the PanelPOMP framework and the Panel Iterated Filtering (PIF) method to explore the practicality and advancement of mechanistic model development for such complex datasets.
-Mechanistic models, especially valuable in ecological research, allow for quantitative analysis and deeper understanding of dynamic systems.
-However, they are often constrained by the non-Gaussian nature of ecological data.
-Our study applies these methods to the Daphnia panel data, consisting of independent time series with 10 data points each, to derive parameters that minimize the Akaike Information Criterion (AIC).
-The comparison of AIC values across models reveals the significant impact of alga, the prey of Daphnia, and parasite on their population dynamics which is observed in the SIRPF model highlights this relationship.
-The panelPOMP framework's ability to handle complex panel data analysis is demonstrated, emphasizing its utility in uncovering latent processes not captured by traditional models.
-However, the study also discusses limitations and interpretations of mechanistic models and PanelPOMP framework in ecological data analysis.
-\end{abstract}
-
-\noindent%
-{\it Keywords:}  Panel Iterated Filtering; SIRPF model; PanelPOMP; Daphnia. 
-\vfill
-
-\newpage
-\spacingset{1.9} % DON'T change the spacing!
-\section{Introduction}
-\label{sec:intro}
-
-Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary.
-As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \citep{Stalidzans2020}.
-Utilizing the established process, they elucidate the actions of hidden mechanisms \citep{Duarte2003}.
-The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems.
-This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters.
-Mechanistic modeling has been widely employed in social sciences, industrial engineering, and physics for an extended period.
-Improved computational capabilities now make it more effective.
-Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information.
-Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task.
-As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \citep{levin1992problem}.\\
-
-Panel data, sometimes referred to as longitudinal data, is composed of multiple time series.
-Each of these time series, which may be multivariate, contains a series of observations collected from a separate entity  \citep{Carles2020}.
-In ecology, panel data is a traditional data form that researchers gather to mitigate the impact of random variations throughout experiments, so that it can more accurately measure changes in experimental objectives and control the variables needed to achieve experimental goals.
-Due to the non-linear characteristics of ecological data, the analysis of panel data entails a high-dimensional structure.
-Such features diminish the explanatory power of traditional time series analysis and certain Monte Carlo inference techniques \citep{johnstone2009statistical}.
-The goal of this study is to investigate the practicality of employing mechanistic models on ecological panel data using the panelPOMP structure.\\
-
-The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \citep{King2016}.
-This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering.
-Stemming from iterated filtering \citep{King2016} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood.
-This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.\\
-
-In this research, we will show the advancement of the panelPOMP model by analyzing the Daphnia panel data generated in the experiment of Searle et al.
-\citep{Searle2016}.
-The research group of Searle et al.
-collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \citep{Searle2016}.
-In their lab experiment, they generated standard ecological panel data featuring various independent trials under the same experimental conditions, with each consisting of a brief, nonlinear, and nonstationary time series.
-Conventional mechanistic theories cannot directly interpret these observations.
-Utilizing the panelPOMP approach allows for the combination of theoretical models and empirical data, enabling the extraction of the comprehensive information found in each individual time series \citep{Carles2020}.\\
-
-For a more in-depth examination of the concealed processes in the panel data, this research incorporates the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model - within the panelPOMP framework.
-After applying the model to the Daphnia panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach.
-These findings are subsequently compared to discover a more accurate and appropriate model for the given data.
-The superior performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is Daphnia's prey, impact the dynamics of both species of Daphnia.
-It is worth noting that the population density of alga is neither observable nor recorded in the experiment.
-That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes \citep{Edwards2012}.
-At the same time, the panelPOMP framework provides an estimation of the parameters in the model, which helps to understand the hidden mechanisms in the ecological system \citep{Carles2020}.
-Based on the estimation, a detailed interpretation of relationship between invasive host and native host with presence of parasite is explaned.\\
-
-Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the Daphnia panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system.
-By analyzing Daphnia panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data.
-Additionally, the later section will further discuss how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models.
-
-\section{Study System}
-\label{sec:studysystem}
-
-The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two Daphnia species: Daphnia \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \citep{searle2016salinization}, and Daphnia \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
-D.
-lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \citep{havel1993daphnia}.
-These Daphnia species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines.
-For their experiments, six clones of each species were selected to eliminate selection bias \citep{Searle2016}.\\
-
-The dynamics also involves \textit{Metschnikowia bicuspidata}, a parasitic fungus predominantly infecting the native host.
-This parasite, ingested as spores by Daphnia, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
-The parasite's presence is identifiable by the opaqueness it causes in normally transparent hosts, with spore release upon host death.
-Searle's study utilized a M.
-bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \citep{Searle2016}.
-The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for Daphnia.
-\textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \citep{arnold1971ingestion}.
-Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of Daphnia species can be closely monitored and manipulated \citep{martinez1994effect}.\\
-
-Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two Daphnia host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \citep{Searle2016}.
-This observation suggests a potential competitive advantage for the invasive species.
-Building on these insights, Searle and colleagues structured the ecological system we analyze here, focusing on dynamics at the community level.
-Their approach involved a mesocosm experiment that varied species composition and parasite presence.
-The experimental design was a 3x2x1 factorial arrangement, incorporating three species combinations (native only, invasive only, or both existing species), two parasite exposure scenarios (exposed or unexposed), and a single algal food source.
-This design enabled an in-depth analysis of the invasive host's impact on native populations, manifesting either directly through competition for resources, indirectly via the amplification of a mutual parasite, or through a reduction in parasite effects.\\
-
-In essence, the impact of an invasive Daphnia \textit{lumholtzi}, particularly in terms of resource competition like food, can be significant.
-This scenario would lead to a decreased survival rate of the native host in the presence of the invasive species, irrespective of parasite interference.
-This effect is quantifiable through the ratio of natality to mortality in the native host population.
-Conversely, should the invasive species primarily influence the native host indirectly via interactions with parasites, a different pattern emerges.
-Here, the survivability of the native host is expected to be lower in scenarios involving both the parasite and the invasive host, as opposed to situations where the native host is only affected by the parasite.
-This effect is quantifiable through the ratio of natality to mortality in the native host population as well.\\
-
-\subsection{Non-parasitized Dynamics}
-This research stratifies the ecological dynamics above into two analytical categories delineated by parasitic presence.
-Under the 'Non-parasitized Dynamics' classification, three experimental groups of Daphnia species—solely native, solely invasive, or a combined population—are statistically analyzed.
-This categorization facilitates a quantitative assessment of the interaction dynamics between native and invasive Daphnia species, particularly in environments not influenced by parasitic interactions.\\
-
-In the Non-parasitized level, indoor mesocosms were used \citep{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media.
-The experimental design involved 10 replicates for single-species treatments, and 9 replicates for mixed-species treatment.
-A nitrogen-to-phosphorus ratio of 25:1 was established, and each mesocosm was seeded with 2.5 $\times$ $10^8$ cells of \textit{Ankistrodesmus falcatus}.
-The laboratory conditions included an average temperature of 23.27$^\circ C$ ($\pm$0.2) and a 16-hour light, 8-hour dark cycle.
-For single-species treatments, either 45 native or invasive Daphnia hosts were added, representing six genotypes per species to eliminate the selection bias.
-In mixed-species treatments, 35 native and 10 invasive hosts were introduced, simulating early invasion stages.\\
-
-\subsection{Parasitized Dynamics}
-Parasitized dynamics is employed with a key difference, which is the presence of the parasite \textit{Metschnikowia bicuspidata} .
-Based on the same setup of Non-parasitized Dynamics, after a 4-day acclimation period in the mesocosms, hosts are exposed to parasites, with spore counts set at 25 per mL in parasite-exposed treatments.
-This category allows for the investigation of the impact of parasites on the interaction dynamics between native and invasive Daphnia species, particularly how parasitism influences competition and survival rates.
-This approach complements the "Non-parasitized Dynamics" by providing a contrasting scenario where parasitic factors are at play, thereby enriching the understanding of ecological interactions under the presence of parasite.\\
-
-\subsection{Data Collection and Visualization}
-Population densities and infection prevalence(if exists) were quantified every five days, beginning a week after the experiment started.
-On each sampling day, mesocosms were thoroughly mixed, and a 1-liter sample was removed and replaced with fresh COMBO media.
-The samples were concentrated using a 153-$\mu m$ mesh and the Daphnia were counted under a microscope.
-For each Daphnia, species (native or invasive), infection status, age, and sex were determined.
-Regular stirring of mesocosms ensured resuspension of algae and parasites.
-Algal and nutrient levels were supplemented bi-weekly, with COMBO media added as necessary.
-The experiment concluded after 52 days, totaling 10 sampling sessions \citep{Searle2016}.\\
-
-As juveniles rarely exhibit clear signs of infection, only adult Daphnia were used for quantifying effects in both Non-parasitized and Parasitized Dynamics to minimize information bias.
-Figures below illustrates these dynamics, with the figure 1 depicting Non-parasitized Dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
-Figure 2 shows Parasitized Dynamics, where the interplay of susceptible and infected Daphnia of different species is further influenced by parasites, complicating the variation of susceptible and infected Daphnia of different races, which makes it difficult to directly analyze the sources of influence on their variation.
-Non-parasitized Dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species.
-And based on this result, the correlation between native host and invasive host in Parasitized Dynamics will be further explored.
-
-\begin{figure}[htbp]
-\centering
-\begin{subfigure}[b]{\linewidth}
-\includegraphics{ms-002}
-\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-\includegraphics{ms-003}
-\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
-\end{subfigure}%
-\end{figure}
-
-\section{Methodology}
-\label{sec:meth}
-\subsection{Partially Observed Markov Processes}
-Frequently, we lack a comprehensive understanding of the underlying mechanisms that drive the evolution of a natural system.
-For example, the precise number of individuals who are exposed to a disease or the density of alga at any given time is typically unknown.
-To address these issues of missing information, we can utilize partially observed Markov (POMP) models.
-These models combine both latent and observable processes which are reliant on each other \citep{bhadra2011malaria}.\\
-
-The application of POMP models spans various fields, including epidemiology, ecology  \citep{king2022markov}, and finance \citep{welton2005estimation} where they are used to infer the dynamics of complex systems from partial observations.
-For instance, in epidemiological studies, POMP models are instrumental in estimating disease transmission rates and predicting future outbreaks by modeling the spread of infectious diseases when only a fraction of cases are observed \citep{xue2020data}.
-Similarly, in ecology, these models help in understanding animal population dynamics and migration patterns based on limited or sporadic sighting data \citep{wood2001partially}.\\
-
-Methodologically, POMP models are analyzed using a suite of computational techniques.
-Sequential Monte Carlo methods, such as particle filters, are commonly employed for state estimation and parameter inference in these models \citep{bhadra2011malaria}.
-These methods involve simulating a large number of potential realizations of the latent process and updating these simulations as new observations become available, thereby approximating the posterior distribution of the state given the data \citep{King2016}.
-Additionally, techniques like iterated filtering and particle Markov chain Monte Carlo are used for parameter estimation and model calibration.\\
-
-With the random variables $Y_{0:N}$ which are the observable states process, the measurement model is $f_{Y_n|X_n}(y_n|x_n)$, and the collected data $y_{0:N}$ are observations of this process.
-We suppose random variables $X_{0:N}$ represent the latent states, where $X_0$ represents the initial state of the latent process, $f_{X|X_{n-1}}(x_n|x_{n-1})$.
-We assume that each $Y_n$ depends only upon the latent process at time $n$, $X_n$, and is conditionally independent of the other variables representing the measurement and latent processes, $Y_{n-1}$ and $X_{n-1}$, where $1 \leq n$.
-Together, $X_{0:N}$ and $Y_{0:N}$ form our POMP model.
-POMP models are vital in situations where we encounter incomplete or noisy measurements of a latent, unobserved Markov process.
-These models, also known as hidden Markov models or state space models, are particularly adept at handling time series data where observations are only partial or indirect representations of the underlying state of the system.
-Their strength lies in the combination of a latent process, which evolves according to Markov dynamics, and an observation process, which relates the latent states to the actual data observed.\\
-
-\subsection{PanelPomp Framework}
-In a panel data set containing K independent time series, K entities are represented by the symbols $1,2,...,K$, with N periodic observations collected for each unit $(t_1,t_2,...,t_{N})$.
-A stochastic observable process, denoted as $y_{k,1},...,y_{k,N}$, generates these measurements.
-PanelPOMP is an adaptation of the partially observable Markov process (POMP) tailored for panel data analysis.
-It entails developing a POMP model for each entity and employing the Panel Iterated Filtering (PIF) method to achieve the highest log-likelihood.
-Based on the POMP framework, a process and a measurement model were devised for every time series unit.
-The measurement model elucidates how the potential real population density generates observable time series data with the measurement density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$, while the process model portrays the actual population's dynamics through the one-step transition density $f_{Xn|Xn-1}(x_n|x_{n-1};\theta)$.
-Ultimately, the joint density equals the initial density multiplied by the product of the process and measurement models for each state.\\
-
-The PanelPOMP framework demonstrates a strong capability in handling ecological panel data due to its ability to model complex, time-varying processes within and across multiple ecological units \citep{Carles2020}.
-It efficiently integrates mechanistic models that capture underlying ecological dynamics with statistical inference methods, accommodating variability and correlations across units.
-This framework is adept at dealing with incomplete or noisy observational data, typical in ecological studies, providing robust estimates and predictions that are crucial for understanding and managing ecological systems.\\
-
-\subsection{Likelihood of models and PIF}
-
-In statistical inference challenges, likelihood frequently serves as the basis for estimating parameters and selecting models.
-Considering a model parameterized by vector $\theta$ within the $m$-dimensional parameter space $\Theta_m$, the likelihood function represents the combined probability density of the dataset, $y_{1:N}$, at $\theta$:
-        \begin{equation}
-        \mathcal{L}(\theta) = f_{Y_{1:N}}(y_{1:N}; \theta)
-        \end{equation}
-The goal is to get the estimation of $\theta$, $\hat{\theta}$ that maximizes the likelihood function $\mathcal{L}(\hat{\theta})$, or log-likelihood function $\mathbf{\ell}(\hat{\theta})$.
-
-An ecological model's effectiveness in tracking disease spread hinges on its capability to predict cases or incidence rates.
-This capability is inherently tied to our trust in the model's precision in forecasting and our grasp of the temporal changes in latent states.
-As a result, we encounter two interrelated issues: ascertaining the distribution of $X_{n}$ at time $n$ based on $y_{1:n}$, and pinpointing parameter values, $\hat{\theta}$, that maximize the data likelihood.
-These processes are commonly respectively concluded as the filtering process and the inference process.\\
-
-Working with POMP models can prove difficult due to the complexity of formulating both the likelihood function and transition density in intricate settings.
-To address this issue, several techniques have been devised, including the particle filter.
-Utilizing the particle filter method, it is only necessary to supply data, initial density simulators, single time-step transition density simulators for the latent process, and an evaluator for the observation process density reliant on the hidden process in order to achieve maximum likelihood estimates.\\
-
-To implement the particle filter, we start by initializing $M$ particles at time 1, where $\{ X_1^m; m \in 1:M \}$.
-All with the required latent state detail and a vector of parameter values, $\theta$.
-As a result, for every time step n, where $n \in $ [$N$], we advance the each particle by doing the one-step random walk based on transition density, resulting in a set of particles that represents the prediction distribution at time n, which can be represented as $f_{X_n|X_{n-1}}(\cdot | X_{n-1}^m; \theta)$.
-Next, we assign weights to the particles based on the measurement density to obtain an ensemble that represents the filtering distribution at time n, where the new weight is calculated via $w_{n,m} = f_{Y_n|X_n}(y_n | x_n^m)$.
-Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \citep{5}.
-The particle filter is particularly useful in situations where both the likelihood function and transition density are difficult to calculate analytically.\\
-
-As we know, the measurement process variables are mutually independent.
-In the mean time, the latent process states are dependent to one other.
-We also notice that the weight we assign to the particles will vary among different particles, which makes the calculation hard.
-As M to be large enough, we may approximate the weight of each particle to be the mean value of the to calculate the log-likelihood.
-Based on such fact, we can summarize as following:
-         \begin{flalign}
-          \mathcal{L}(\theta) &= f_{Y_{1:n}}(y_{1:N}; \theta) \\
-         & = \prod_{n=1}^N f_{Y_{n} | Y_{1:n-1}}(y_n | y_{n-1}; \theta) \\
-         & = \prod_{n=1}^N f_{Y_n | X_n}(y_n | x_n ; \theta)\\
-         & \approx \prod_{n=1}^N \frac{1}{M} \sum_{m = 1}^M f_{Y_n|X_n}(y_n | x_n^m) 
-         \end{flalign}
-The particle filter, a form of Sequential Monte Carlo method, is instrumental in processing partially observed Markov processes within complex statistical models \citep{ionides2006inference}.
-This technique hinges on the generation of multiple, simulated "particles" that each represent a potential state of the latent process \citep{ionides2011iterated}.
-As new observational data becomes available, these particles are updated and assigned weights proportional to their likelihood of representing the true latent state, considering the observed data.\\
-
-This weighting system is key to the particle filter's efficacy.
-It allows for the approximation of the distribution of the latent process at any given point in time, as well as the conditional likelihood of the observed data given the model.
-Essentially, the particle filter distills a complex, high-dimensional problem into a manageable computation by focusing on the most probable states of the latent process, as indicated by the observed data.\\
-
-Panel Iterated Filtering (PIF) extends the improved iterated filtering algorithm, originally developed for time series data, to panel models \citep{Carles2020}.
-This method is particularly effective in exploring parameter spaces in complex, nonlinear, non-Gaussian models.
-It does this by introducing random perturbations to the model parameters and utilizing a particle filter to explore the parameter space at each iteration.
-The convergence to a maximum likelihood estimate is achieved through procedures that iterate this search over the parameter space while diminishing the intensity of perturbations.\\
-
-PIF's functionality is especially valuable in analyzing complex dynamical systems where the standard modeling approaches may fall short due to the intricacies of the data and the underlying processes.
-The ability of PIF to handle high-dimensional parameter spaces and its flexibility in managing different types of data structures make it a robust tool for statistical inference in complex models, particularly those encountered in ecological and epidemiological studies.\\
-
-\section{Mechanistic models}
-\label{sec:mecmod}
-Mechanistic models are highly effective in understanding ecological dynamics due to their ability to incorporate detailed biological processes and interactions within ecosystems \citep{young1998data}.
-These models simulate the causal mechanisms underlying ecological phenomena, offering insights into complex ecological processes that are often difficult to observe directly.
-By integrating empirical data with theoretical principles, mechanistic models enable the prediction and analysis of ecological outcomes under varying conditions.
-This makes them invaluable tools for addressing challenges in ecology \citep{mouquet2015predictive}.\\
-
-In mechanistic modeling of dynamic systems, the foundational step involves the formulation of differential or difference equations that encapsulate the temporal evolution of the system.
-This approach is pivotal in time series analysis using mechanistic models, which primarily aims to assess the model's adequacy in capturing the dynamics of the system and, upon validation, estimate unknown parameters that govern these dynamics \citep{breto2009time}.
-The complexity in such models often arises from intrinsic stochasticity, nonlinearity, and the presence of noisy and incomplete observational data, features that are particularly prevalent in biological sciences \citep{wilkinson2018stochastic}.\\
-
-To analyze the dynamics of Parasitized and Non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \citep{weiss2013sir}, is utilized.
-This model integrates the aspects of food and parasites, unlike the classical SIR model, to more accurately reflect the interactions between native and invasive hosts in the context of resource availability and parasitic influence.
-The SIRPF model incorporates these factors as quantitative latent state variables, enabling a more comprehensive understanding of the system's dynamics.
-In the meantime, the ecological processes of Daphnia can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
-Utilizing panel data on Daphnia, the PanelPOMP framework becomes a valuable analytical tool \citep{Carles2020}.
-Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
-
-The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the following graph and model show:
-
-\usetikzlibrary{positioning}
-\usetikzlibrary {arrows.meta}
-\usetikzlibrary{shapes.geometric}
-
-\begin{figure}
-\begin{center}
-%%%%% SEAIR diagram
-\resizebox{8cm}{!}{
-\begin{tikzpicture}[
-  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
-  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1},
-  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
-  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
-  bendy/.style={bend left=10},
-  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
-  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
-]
-\tikzset{>={}}; % this is needed to implement the arrow redefinition
-
-\node (S) at (-0.5,-0.25) [square] {$S$};
-\node (I) at (4.5,-0.25) [square] {$I$};
-\node (P) at (2.5,0.75)  [rhombus] {$P$};
-\node (R) at (3.5,2)[travel] {$R$};
-
-\node (F) at (4.5,4.5) [report] {$F$};
-
-\draw [->] (F) --  (R);
-\draw [->] (P) --  (R);
-\draw [->] (S) -- (R);
-\draw [->] (I) -- (R);
-
-\draw [->] (F) -- (S);
-\draw [->] (F) -- (I);
-
-\draw (S) edge[loop left] (S);
-\draw (F) edge[loop above] (F);
-
-\draw [->, bendy] (S) to  (I);
-\draw [->, bendy] (I) to  (S);
-\draw [->, bendy] (P) to (I);
-% \path [->] (S) edge (I) 
-%            (P) edge (I);
-\draw [->, bendy] (I) to (P);
-\end{tikzpicture}
-}
-\end{center}
-\vspace{-5mm}
-\caption{A flow diagram for the SIRPF model.
-Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
-For Daphnia either dentifera or lumholtzi, susceptible, infected Daphnia and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$.
-Some susceptible Daphnia get infected by parasites, transitioning from $S$ to $I$.
-Both susceptible and infected Daphnia contribute to the $R$ state upon death, releasing surviving parasites back into the environment.
-And the parasites continue to impact other individuals and get into another cycle.
-Food is consumed by both $S$ and $I$ individuals.
-Over time, elements in $F$ and $P$ also transit to the $R$ state.}
-\label{fig:flow_SIRPF}
-\end{figure}
-
-{
-\small
-\begin{align}
-&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S},\\
-&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I},\\
-&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
-&d P(t) = \beta d I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
-&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
-&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}dt\right),\\
-&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
-&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
-\end{align}
-}
-
-Each equation intricately ties the state variables to depict the complex dynamics of the single species Parasitized system.
-Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race Parasitized system by introducing another Daphnia species.
-This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$)hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows.
-In the SIRPF2 model, the subscript n is used to denote parameters and variables related to the native host, i for those pertaining to the invasive host, and p for the parasite spores.
-The SIRPF2 model thus offers a more detailed representation of the ecological dynamics by accounting for the complex interplay between different species and environmental factors in a mechanistic framework.
-\begin{figure}
-\begin{center}
-%%%%% SEAIR diagram
-\resizebox{8cm}{!}{
-\begin{tikzpicture}[
-  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
-  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
-  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
-  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
-  bendy/.style={bend left=10},
-  bendy2/.style={bend left=100},
-  bendy3/.style={bend left=-100},
-  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
-  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
-]
-\tikzset{>={}}; % this is needed to implement the arrow redefinition
-
-\node (Sn) at (1,2) [square] {$S_n$};
-\node (In) at (4,2) [square] {$I_n$};
-\node (Si) at (1,-2) [square] {$S_i$};
-\node (Ii) at (4,-2) [square] {$I_i$};
-\node (R) at (2.5,0)  [rhombus] {$R$};
-\node (P) at (5.5,0)[travel] {$P$};
-\node (F) at (-0.5,0) [report] {$F$};
-
-\draw [->, bendy] (Sn) to  (In);
-\draw [->, bendy] (Si) to  (Ii);
-\draw [->, bendy] (In) to  (Sn);
-\draw [->, bendy] (Ii) to  (Si);
-
-\draw [->] (F) --  (R);
-\draw [->] (P) --  (R);
-\draw [->] (Sn) -- (R);
-\draw [->] (In) -- (R);
-\draw [->] (Si) -- (R);
-\draw [->] (Ii) -- (R);
-\draw [->] (F) -- (Sn);
-\draw [->] (F) -- (Si);
-\draw (Sn) edge[loop below] (Sn);
-\draw (Si) edge[loop above] (Si);
-\draw (F) edge[loop left] (F);
-
-\draw [->, bendy] (P) to (In);
-\draw [->, bendy] (P) to (Ii);
-\draw [->, bendy] (In) to (P);
-\draw [->, bendy] (Ii) to (P);
-\draw [->, bendy2] (F) to (In);
-\draw [->, bendy3] (F) to (Ii);
-
-\end{tikzpicture}
-}
-\end{center}
-\vspace{-5mm}
-\caption{A flow diagram for the SIRPF2 model.
-Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
-The flow diagram shows the how two species of Daphnia interact with each other indirectly through $F$ and $P$.
-The marginal processes have been shown in SIRPF mode, and SIRPF2 model is the adaption form in mixed-species dynamcis}
-\label{fig:flow_SIRPF2}
-\end{figure}
-
-{
-\small
-\begin{align}
-% \begin{split}
-&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n},\\
-% \end{split}
-% \\
-&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n},\\
-% \begin{split}
-&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i},\\
-% \end{split}
-% \\
-&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
-\begin{split}
-&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
-&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
-\end{split}
-\\
-\begin{split}
-&d P(t) = \beta_n d_n I_n(t)dt + \beta_i d_i I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
-&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
-\end{split}
-\\
-&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}dt\right),\\
-&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}dt\right),\\
-&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}dt\right),\\
-&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}dt\right),\\
-&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right),\\
-&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right).
-\end{align}
-}
-
-In the SIRPF2 model, detailed in Table 1, parameters and variables are distinctly annotated for native hosts (subscript $n$), invasive hosts (subscript $i$), and parasite spores (subscript $p$).
-Intraspecific competition among hosts is conceptualized via the Lotka-Volterra framework, with $r_n$ and $r_i$ denoting respective reproductive rates.
-Host density growth, influenced by algal density, is posited to be linear, indicating a direct relationship between food availability and Daphnia reproduction.
-Mortality rates of susceptible hosts are denoted as $\theta_{S_n}$ and $\theta_{S_i}$.
-The model encapsulates the dynamics of infection through \textit{Metschnikowia bicuspidata} intake, affecting susceptible hosts.
-This interaction consequently modulates reproductive rates, represented as $\xi r_n$.
-Infection rates correlate with spore density ($P$) and host ingestion rates ($f_{S_n}$, $f_{S_i}$).
-Upon death, infected hosts ($\theta_{I_n}$, $\theta_{I_i}$) release spores, which are subsequently subject to ingestion and environmental degradation.
-The consumption dynamics of food by susceptible and infected hosts are captured by $\gamma_n$ and $\gamma_i$, and food growth is inversely proportional to its carrying capacity ($K_F$), reflective of resource limitation.
-The model also incorporates a constant removal rate $\delta$, representing the loss of hosts, spores, and food due to external sampling.
-The constraint of food supply, indicated by a constant rate $\mu = 0.37 \times 10^5$ cells per day, is based on observed declines in Daphnia populations in the panel data, suggesting a pivotal role of food availability in driving population dynamics.\\
-
-\renewcommand{\arraystretch}{0.5}
-\begin{table}[H]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c||}
- \hline
-Parameter & Definition &Unit\\[0.5ex]
-\hline\hline
-$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$\\
-$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$\\
-$F$& Alga density& $10^5 cells \cdot L ^{-1}$\\
-$P$ & Spore density & $10^3 spores \cdot L ^{-1}$\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ \\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$ \\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$ \\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ \\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ \\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ \\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ \\
-$\delta$ & Sampling rate&$day ^{-1} $ \\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ \\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ \\
-$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ \\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ \\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ \\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ \\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ \\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ \\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ \\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ \\[1ex]
- \hline
-\end{tabular}}
-\caption{This table shows the units of the parameters of competition SIRPF2 model}
-\label{Table:1}
-\end{table}
-
-\renewcommand{\arraystretch}{1}
-In the realm of ecological modeling, particularly in the context of the SIRPF2 model, the stochastic nature of ecological processes is adeptly captured using the principles of Brownian motion.
-This concept, fundamental to the theory of random particle movement in fluids, is extended in our model to encapsulate the inherent randomness in ecological dynamics.
-In the SIRPF2 model, Brownian motion is manifested through stochastic terms ($\zeta_F$, $\zeta_{S_n}$, $\zeta_{S_i}$, $\zeta_{I_n}$, $\zeta_{I_i}$, $\zeta_{F}$, $\zeta_P$), which follow a normal distribution with zero mean and specific standard deviations.
-These stochastic elements are crucial for modeling the unpredictable variances in the population dynamics and environmental interactions, especially in the face of unforeseen environmental perturbations or random demographic events.\\
-
-The integration of these stochastic terms within the SIRPF2 framework, as detailed in Table 1, enhances the model's fidelity in simulating ecological interactions.
-The model, grounded in the Lotka-Volterra competition framework and augmented with stochastic components, provides a nuanced understanding of the interplay between native and invasive hosts, and their interaction with the surrounding resources and parasites.
-This incorporation of stochasticity, particularly through the lens of Brownian motion, is pivotal in transcending the limitations of deterministic models, thereby offering a more robust and comprehensive tool for ecological analysis.\\
-
-\subsection{SIRPF adaption: SF and SF2 model}
-
-The approach of analyzing Non-parasitized Dynamics through adapted models of the SIRPF framework, such as the Susceptible - Food (SF) and SF2 models, is a strategic method to quantify the impact of environmental resources, like food($F$), on the interactions between native($S_n$) and invasive($S_i$) host species.
-The SF model focuses on the dynamics between susceptible individuals and food resources, while the SF2 model extends this concept to include Mixed Race Dynamics.
-This extension is particularly useful in understanding how different species, native and invasive, interact with each other and with their environment in the absence of parasitic factors.\\
-
-The adaptation of the SIRPF model to SF and SF2 as following allows for a more focused examination of the roles that resources play in ecological systems, especially when considering the competition or coexistence between different species.
-By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.\\
-
-\begin{figure}
-\begin{center}
-%%%%% SEAIR diagram
-\resizebox{8cm}{!}{
-\begin{tikzpicture}[
-  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
-  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
-  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
-  bendy/.style={bend left=10},
-  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
-  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
-]
-\tikzset{>={}}; % this is needed to implement the arrow redefinition
-
-\node (Sn) at (-0.5,0) [square] {$S_n$};
-\node (F) at (2.5,0) [report] {$F$};
-\node (Si) at (5.5,0) [square] {$S_i$};
-\node (R) at (2.5,2) [square] {$R$};
-
-\draw [->] (F) --  (Sn);
-\draw [->] (F) --  (Si);
-\draw [->] (F) --  (R);
-\draw [->] (Sn) --  (R);
-\draw [->] (Si) --  (R);
-\draw (Sn) edge[loop above] (Sn);
-\draw (Si) edge[loop above] (Si);
-\draw (F) edge[loop below] (F);
-
-\draw [dotted, thick, red] (3.5, -0.75) rectangle (6.5, 1.5);
-\end{tikzpicture}
-}
-\end{center}
-\vspace{-5mm}
-\caption{A flow diagram for the SF and SF2 model.
-In this system, both types of Daphnia and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$.
-As time progresses, individuals in $F$, $S_n$, and $S_i$ transition to the $R$ state, signifying mortality.
-While the whole graph show SF2 model, if we remove the node in the dashed rectangle, the graph will represent the SF model.}
-\label{fig:SF&SF2}
-\end{figure}
-
-\begin{align}
-&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
-&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
-&dF=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
-&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
-&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
-&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
-\end{align}
-
-As a adaption form of SIRPF model, the upper equations show SF2 model, with the meaning of each parameter is the same as the SIRPF model.
-The only difference is that the SF and SF2 model do not have the parasitic term.
-While if we only keep one species in SF2 model, the model will become SF model.
-Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of Daphnia, particularly under conditions of potential food scarcity.
-This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia.
-Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver.
-Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia.
-In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of Daphnia.\\
-
-\subsection{Measurement and Initial condition}
-In the ecological study of Daphnia populations, we employ the negative binomial distribution to address the inherent overdispersion observed in population counts.
-This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \citep{ning2021systemic}.
-In our model, the parameter $F$ represents the number of successful Daphnia reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of Daphnia per liter per sampling time in each trial.
-This framework aligns with the observed dynamics in Daphnia populations, where variability in reproduction rates, influenced by fluctuating food availability, leads to overdispersed population counts.
-The negative binomial distribution, therefore, provides a robust statistical tool to assess population dynamics, accommodating the irregularities in reproduction and survival patterns shaped by environmental constraints.\\
-
-The model's effectiveness in capturing the Daphnia population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
-As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing Daphnia populations, particularly under varying food conditions.
-
-\begin{align}
-F \sim \operatorname{NBinomial}(n=S / \delta, \tau = k_F), \delta=0.013 
-\end{align}
-
-\section{Results}
-\label{sec:res}
-When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability.
-In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality.
-For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of Daphnia, and 8 for the rest dynamics.\\
-
-Permitting parameters to be either shared or unique leads to varying dimensions for each model.
-With respective amount of units $d$ in the panel, adding a shared parameter increases the model's dimension by 1, and a specific parameter will add dimensions by $d$.
-We encountered a balance between fitting accuracy and simplicity when selecting models.
-Akaike information criterion (AIC) is a tool for model selection that balances model complexity against goodness of fit \citep{bozdogan1987model}.
-It is rooted in information theory, particularly the concept of entropy, and provides a relative measure of the information lost when a given model is used to represent the process that generated the data.
-A lower AIC value indicates a model with a better balance between fit and simplicity.
-In our study, we utilized AIC to compare models, ensuring adjustments were made for the number of parameters, thus facilitating a fair comparison across models with differing complexities.\\
-
-The parameters are estimated via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models, which are presented in the Supplementary Material.
-Our analysis demonstrates that the 'All shared parameter' model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics.
-In cases like the Non-parasitized dynamics of single species, where this model's AIC is slightly higher (but within 2 points of the lowest AIC), it doesn't necessarily imply inferiority of model.
-AIC balances fit and complexity but doesn't consider stability or unseen data prediction.
-Here, Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \citep{nylund2007deciding}.
-This highlights the importance of looking beyond AIC for model assessment.
-As a result, in our case, the 'All shared parameter' model can be the best choice for all dynamics.\\
-
-By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species.
-The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}.
-This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.
-The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are 1.3 and 1.3 respectivly.
-While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to 0.13 for native Daphnia and to 0.13 for invasive species.
-In such scenario, we observe significant fluctuations in the population density of Daphnia.
-These oscillations are likely attributable to the constraints imposed by limited food resources within the environment.
-Such environmental limitations typically lead to variations in population dynamics.
-The observed variability in the survivability ratio can be understood within this assumption.
-At the same time, it is important to note that the relative survivability between the two Daphnia species studied remains consistent, despite an overall simultaneous decrease in their population densities.
-This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them.
-And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are 1.00 and 0.99 for single species dynamics, and 2.4, 0.35 for native and invasive population in mixed species dynamics respectively.\\
-
-These findings underscore the importance of parameter sharing in ecological modeling when multiple species or units are involved.
-Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between Daphnia populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones.
-The results further reinforce the utility of the AIC in guiding the selection of models that not only provide a good fit to the data but also maintain a degree of simplicity, avoiding overfitting.
-The SIRPF model's performance with shared parameters points to a commonality in the factors affecting both native and invasive species, a result that could have implications for the management of these populations and their associated ecological interactions.\\
-
-\section{Discussion}
-\label{sec:dis}
-
-The SIRPF model is a powerful tool for analyzing ecological data, as it can be used to estimate parameters and predict future population dynamics.
-The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia.
-The model's ability to estimate parameters and predict future population dynamics is a powerful tool for analyzing ecological data.
-The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia.\\
-
-The SIRPF model extends the traditional SIR framework by including food and parasites, which is crucial in ecological models where these factors significantly influence host dynamics.
-This approach reflects the understanding that the dynamics of pathogens, including coinfections and community composition, are significantly influenced by ecological factors such as food availability and parasitic interactions.
-The complexity of these interactions is highlighted by studies emphasizing the role of community ecology in infectious disease research and the phylogenetic structure of host-parasite relationships.
-The SIRPF model's inclusion of these aspects offers a more comprehensive understanding of ecological dynamics, particularly in systems where traditional SIR models may fall short.\\
-
-The PanelPOMP framework has shown remarkable efficacy in managing complex ecological systems \citep{Carles2020}, particularly when data points are limited and exhibit high volatility.
-Utilizing Panel Iterated Filtering, PanelPOMP adeptly fits data and estimates parameters that correspond to maximum log likelihood.
-This characteristic is crucial in ecological studies where panel data is prevalent, as it allows for more accurate modeling of complex, dynamic systems.
-However, the efficacy of PanelPOMP could be further enhanced with a larger dataset, as this would allow for a more restricted range of variations in simulations, leading to potentially more precise parameter estimations.
-This finding underscores the importance of robust data collection in ecological modeling and the value of PanelPOMP in analyzing such data.\\
-
-The PanelPOMP model, while offering significant advantages in analyzing complex ecological data \citep{Carles2020}, does encounter certain limitations, especially when dealing with high-dimensional models like the SIRPF.
-One of the primary challenges associated with PanelPOMP is the increased computational resources and time required to achieve optimal log likelihood estimations.
-This is particularly pronounced in high-dimensional models due to the complex and intricate nature of the data and model structure.
-Thus, the model's effectiveness in these scenarios might necessitate complementary methodologies that are more attuned to the practical situations encountered at higher latitudes.
-Despite these challenges, the PanelPOMP model still represents a significant advancement in the field of ecological modeling, especially in its ability to handle partially observed Markov process models.
-The model's framework allows for a more nuanced understanding of complex ecological systems by facilitating likelihood-based inference for intricate and interrelated ecological processes.\\
-
-\bibliography{reference}
-\end{document}

### 2024-03-06 ms.Rnw
modify Rnw and tex
Update

--- 2024-01-24 ms.Rnw
@@
 
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Results, include=FALSE,echo=FALSE,results=hide>>=
 load('all_shared.RData')
@@
 parms = mif.estimate
-parms = as.data.frame(t(parms))
-
+# parms = as.data.frame(t(parms))
+
+parms$K_F = parms$K_F * 1000
+parms$ri = parms$ri * 10
+parms$rn = parms$rn * 10
+parms$theta_Ii = parms$theta_Ii * 10
+parms$theta_In = parms$theta_In * 10
+parms$theta_Si = parms$theta_Si * 10
+parms$theta_Sn = parms$theta_Sn * 10
+parms$theta_P = parms$theta_P * 1000
+parms$probi = parms$probi * 10
+parms$probn = parms$probi * 10
+parms$f_Sn = parms$f_Sn * 10000
+parms$f_Si = parms$f_Si * 10000
+parms$f_In = parms$f_In * 1000
+parms$f_Ii = parms$f_Ii * 1000000
+parms$alpha = parms$alpha * 10000
+parms$gamman = parms$gamman * 1e8
+parms$gammai = parms$gammai * 1e4
+
+parms$sigSi = parms$sigSi * 1e4
+parms$sigIn = parms$sigIn * 1e5
+parms$sigIi = parms$sigIi * 1e5
+parms$sigF = parms$sigF * 10
+parms$sigP = parms$sigP * 100
 @
@@
 $P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
+$K_F$ & Alga carrying capacity&$10^2\cdot cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
+$r_n$ & Native growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
+$r_i$ & Invasive growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ 10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
+$\theta_{P}$ & Spore degradation rate&$10^{-1} \cdot  day^{-3}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
+$p_n$ &Native probability of infection per spore &$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
+$p_i$ & Invasive probability of infection per spore&$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
+$f_{S_n}$ & Native susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
+$f_{I_n}$ & Native infected host filtering rate&$10^{-3} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
+$f_{I_i}$ & Invasive infected host filtering rate&$10^{-6} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
 $\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
+$\alpha$ & Alga maximum exponential growth rate&$10^{-4} \cdot day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
+$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
+$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
 $\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
@@
 $\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
+$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
+$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
+$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
+$\sigma_{F}$ & Standard deviation of brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
+$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
 $k_{S_n}$ & Overdispersion parameter& $Unitless$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
@@
 &d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
-&dF=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
+&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
 &\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\

### 2024-03-13 ms.Rnw
Update

@@
  library(reshape2)
+library(patchwork)
  library(magrittr)
@@
  #----------------load data---------------------------------------------------------------
-load("/Users/ybb/Desktop/Daphnia-article/data/raw_data.RData")
+load("data/raw_data.RData")
  #-----------------------------Host with para----------------------------------------------------------------------------
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
 \centering
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=  
+<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=16, height=10>>=  
    mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
@@
    }))
-
- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
+    mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
+     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
+   }))
+
+p1<- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
    geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
@@
    scale_y_continuous(limits = c(0, 175)) +
-   facet_grid(. ~ group, scales = "free", space = "free") +
+   facet_grid(. ~ group, scales = "fixed", space = "free") +
    theme(
@@
    guides(colour=guide_legend(title="Species"))
-@
-\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=  
-   mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
-     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
-   }))
-
- ggplot(mixed_para_data_combined, aes(x=day)) + 
+
+p2<- ggplot(mixed_para_data_combined, aes(x=day)) + 
    geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
@@
    guides(colour=guide_legend(title="Species"))
+combined_plot <- p1 / p2 
+combined_plot
 @
-\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia. The trail A to J represents different buckets in the experiment}
-\end{subfigure}%
+\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia.
+The trail A to J represents different buckets in different experiment.
+The black lines represent the Daphnia dentifera density, the red lines show the Daphnia lumholtzi density, the green lines are the infected Daphnia lumholtzi density, and the blue lines indicate the infected Daphnia dentifera density.}
 \end{figure}
@@
 &dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
-&d P(t) = \beta d I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
+&d P(t) = \beta \theta_I I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
 &d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
@@
 \begin{split}
-&d P(t) = \beta_n d_n I_n(t)dt + \beta_i d_i I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
+&d P(t) = \beta_n \theta_{I_n} I_n(t)dt + \beta_i \theta_{I_i} I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
 &\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
@@
 $\alpha$ & Alga maximum exponential growth rate&$10^{-4} \cdot day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
-$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
-$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
+$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
+$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
 $\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
@@
 $\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betan,2)}\\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
+$\sigma_{S_n}$ & Standard deviation of Brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
+$\sigma_{S_i}$ & Standard deviation of Brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
+$\sigma_{I_n}$ & Standard deviation of Brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
+$\sigma_{I_i}$ & Standard deviation of Brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
+$\sigma_{F}$ & Standard deviation of Brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
+$\sigma_{P}$ &Standard deviation of Brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
 $k_{S_n}$ & Overdispersion parameter& $Unitless$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
@@
 
-\begin{figure}
+\begin{figure}[H]
 \begin{center}
@@
 &d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
-&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
+&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n(t) \cdot dt - \gamma_i F(t) S_i(t) \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
 &\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
@@
 This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia.
-Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver.
+Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver \citep{xiaotong2018}.
 Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia.
@@
  ratio_SF_i = SF_invasive_estimate[length(SF_invasive_estimate)]/SF_invasive_estimate[4]
- ratio_SF_n = ratio_SF_i
+ ratio_SF_n = SF_native_estimate[length(SF_native_estimate)]/SF_native_estimate[4]
  ratio_SF2_i = SF2_estimate[11]/SF2_estimate[8]
@@
  ratio_SIRPF_ni = SIRPF_native_estimate[13] * SIRPF_native_estimate[14]/SIRPF_native_estimate[9]
- ratio_SIRPF2_ns = SIRPF2_estimate[20]/SIRPF2_estimate[10]
+ ratio_SIRPF2_ns = mif.estimate$rn/mif.estimate$theta_Sn
  ratio_SIRPF2_ni = SIRPF2_estimate[20] *  SIRPF2_estimate[2]/SIRPF2_estimate[12]
@@
  ratio_SIRPF2_ii = SIRPF2_estimate[21] *  SIRPF2_estimate[2]/SIRPF2_estimate[14]
+
+ save(SF_invasive_estimate=SF_invasive_estimate,SF_native_estimate=SF_native_estimate,SF2_estimate=SF2_estimate,SIRPF_invasive_estimate=SIRPF_invasive_estimate,SIRPF_native_estimate=SIRPF_native_estimate,SIRPF2_estimate=SIRPF2_estimate,file = "SIRPF_model_estimate.RData")
 @
@@
 This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.
-The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,2)} and \Sexpr{mysignif(ratio_SF_i,2)} respectivly.
-While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,2)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,2)} for invasive species.
+The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectivly.
+While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species.
 In such scenario, we observe significant fluctuations in the population density of Daphnia.
@@
 This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them.
-And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,2)} and \Sexpr{mysignif(ratio_SIRPF_is,2)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,2)}, \Sexpr{mysignif(ratio_SIRPF2_is,2)} for native and invasive population in mixed species dynamics respectively.\\
+And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.\\
 

### 2024-03-20 ms.Rnw
Update

@@
  library(ggplot2)
+ library(gridExtra)
 
@@
 However, they are often constrained by the non-Gaussian nature of ecological data.
-Our study applies these methods to the Daphnia panel data, consisting of independent time series with 10 data points each, to derive parameters that minimize the Akaike Information Criterion (AIC).
-The comparison of AIC values across models reveals the significant impact of alga, the prey of Daphnia, and parasite on their population dynamics which is observed in the SIRPF model highlights this relationship.
+Our study applies these methods to the \textit{Daphnia} panel data, consisting of independent time series with 10 data points each, to derive parameters that minimize the Akaike Information Criterion (AIC).
+The comparison of AIC values across models reveals the significant impact of alga, the prey of \textit{Daphnia}, and parasite on their population dynamics which is observed in the SIRPF model highlights this relationship.
 The panelPOMP framework's ability to handle complex panel data analysis is demonstrated, emphasizing its utility in uncovering latent processes not captured by traditional models.
@@
 
-In this research, we will show the advancement of the panelPOMP model by analyzing the Daphnia panel data generated in the experiment of Searle et al.
+In this research, we will show the advancement of the panelPOMP model by analyzing the \textit{Daphnia} panel data generated in the experiment of Searle et al.
 \citep{Searle2016}.
 The research group of Searle et al.
-collected panel data on the population densities of different species of Daphnia under different conditions, and interpret how invasive Daphnia species (\textit{lumholtzi}) affect the response of native Daphnia species (\textit{dentifera}) to parasites with panel data \citep{Searle2016}.
+collected panel data on the population densities of different species of \textit{Daphnia} under different conditions, and interpret how invasive \textit{Daphnia} species (\textit{lumholtzi}) affect the response of native \textit{Daphnia} species (\textit{dentifera}) to parasites with panel data \citep{Searle2016}.
 In their lab experiment, they generated standard ecological panel data featuring various independent trials under the same experimental conditions, with each consisting of a brief, nonlinear, and nonstationary time series.
@@
 For a more in-depth examination of the concealed processes in the panel data, this research incorporates the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model - within the panelPOMP framework.
-After applying the model to the Daphnia panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach.
+After applying the model to the \textit{Daphnia} panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach.
 These findings are subsequently compared to discover a more accurate and appropriate model for the given data.
-The superior performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is Daphnia's prey, impact the dynamics of both species of Daphnia.
+The superior performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is \textit{Daphnia}'s prey, impact the dynamics of both species of \textit{Daphnia}.
 It is worth noting that the population density of alga is neither observable nor recorded in the experiment.
@@
 
-Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the Daphnia panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system.
-By analyzing Daphnia panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data.
+Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the \textit{Daphnia} panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system.
+By analyzing \textit{Daphnia} panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data.
 Additionally, the later section will further discuss how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models.
@@
 
-The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two Daphnia species: Daphnia \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \citep{searle2016salinization}, and Daphnia \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
+The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two \textit{Daphnia} species: \textit{Daphnia} \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \citep{searle2016salinization}, and \textit{Daphnia} \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
 D.
 lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \citep{havel1993daphnia}.
-These Daphnia species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines.
+These \textit{Daphnia} species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines.
 For their experiments, six clones of each species were selected to eliminate selection bias \citep{Searle2016}.\\
@@
 The dynamics also involves \textit{Metschnikowia bicuspidata}, a parasitic fungus predominantly infecting the native host.
-This parasite, ingested as spores by Daphnia, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
+This parasite, ingested as spores by \textit{Daphnia}, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
 The parasite's presence is identifiable by the opaqueness it causes in normally transparent hosts, with spore release upon host death.
@@
 bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \citep{Searle2016}.
-The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for Daphnia.
+The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for \textit{Daphnia}.
 \textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \citep{arnold1971ingestion}.
-Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of Daphnia species can be closely monitored and manipulated \citep{martinez1994effect}.\\
-
-Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two Daphnia host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \citep{Searle2016}.
+Its significance is particularly evident in controlled experimental settings where the growth conditions and feeding rates of \textit{Daphnia} species can be closely monitored and manipulated \citep{martinez1994effect}.\\
+
+Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two \textit{Daphnia} host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \citep{Searle2016}.
 This observation suggests a potential competitive advantage for the invasive species.
@@
 
-In essence, the impact of an invasive Daphnia \textit{lumholtzi}, particularly in terms of resource competition like food, can be significant.
+In essence, the impact of an invasive \textit{D.
+lumholtzi}, particularly in terms of resource competition like food, can be significant.
 This scenario would lead to a decreased survival rate of the native host in the presence of the invasive species, irrespective of parasite interference.
@@
 This research stratifies the ecological dynamics above into two analytical categories delineated by parasitic presence.
-Under the 'Non-parasitized Dynamics' classification, three experimental groups of Daphnia species—solely native, solely invasive, or a combined population—are statistically analyzed.
-This categorization facilitates a quantitative assessment of the interaction dynamics between native and invasive Daphnia species, particularly in environments not influenced by parasitic interactions.\\
+Under the 'Non-parasitized Dynamics' classification, three experimental groups of \textit{Daphnia} species—solely native, solely invasive, or a combined population—are statistically analyzed.
+This categorization facilitates a quantitative assessment of the interaction dynamics between native and invasive \textit{Daphnia} species, particularly in environments not influenced by parasitic interactions.\\
 
@@
 The laboratory conditions included an average temperature of 23.27$^\circ C$ ($\pm$0.2) and a 16-hour light, 8-hour dark cycle.
-For single-species treatments, either 45 native or invasive Daphnia hosts were added, representing six genotypes per species to eliminate the selection bias.
+For single-species treatments, either 45 native or invasive \textit{Daphnia} hosts were added, representing six genotypes per species to eliminate the selection bias.
 In mixed-species treatments, 35 native and 10 invasive hosts were introduced, simulating early invasion stages.\\
@@
 Based on the same setup of Non-parasitized Dynamics, after a 4-day acclimation period in the mesocosms, hosts are exposed to parasites, with spore counts set at 25 per mL in parasite-exposed treatments.
-This category allows for the investigation of the impact of parasites on the interaction dynamics between native and invasive Daphnia species, particularly how parasitism influences competition and survival rates.
+This category allows for the investigation of the impact of parasites on the interaction dynamics between native and invasive \textit{Daphnia} species, particularly how parasitism influences competition and survival rates.
 This approach complements the "Non-parasitized Dynamics" by providing a contrasting scenario where parasitic factors are at play, thereby enriching the understanding of ecological interactions under the presence of parasite.\\
@@
 On each sampling day, mesocosms were thoroughly mixed, and a 1-liter sample was removed and replaced with fresh COMBO media.
-The samples were concentrated using a 153-$\mu m$ mesh and the Daphnia were counted under a microscope.
-For each Daphnia, species (native or invasive), infection status, age, and sex were determined.
+The samples were concentrated using a 153-$\mu m$ mesh and the \textit{Daphnia} were counted under a microscope.
+For each \textit{Daphnia}, species (native or invasive), infection status, age, and sex were determined.
 Regular stirring of mesocosms ensured resuspension of algae and parasites.
@@
 
-As juveniles rarely exhibit clear signs of infection, only adult Daphnia were used for quantifying effects in both Non-parasitized and Parasitized Dynamics to minimize information bias.
+As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both Non-parasitized and Parasitized Dynamics to minimize information bias.
 Figures below illustrates these dynamics, with the figure 1 depicting Non-parasitized Dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
-Figure 2 shows Parasitized Dynamics, where the interplay of susceptible and infected Daphnia of different species is further influenced by parasites, complicating the variation of susceptible and infected Daphnia of different races, which makes it difficult to directly analyze the sources of influence on their variation.
+Figure 2 shows Parasitized Dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation.
 Non-parasitized Dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species.
@@
    guides(colour=guide_legend(title="Species"))
-combined_plot <- p1 / p2 
-combined_plot
+grid.arrange(p1, p2, ncol = 1)
 @
-\caption{Plot of density of hosts in dynamics including both native and invasive host species of Daphnia.
+\caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}.
 The trail A to J represents different buckets in different experiment.
-The black lines represent the Daphnia dentifera density, the red lines show the Daphnia lumholtzi density, the green lines are the infected Daphnia lumholtzi density, and the blue lines indicate the infected Daphnia dentifera density.}
+The black lines represent the \textit{D.
+dentifera} density, the red lines show the \textit{D.
+lumholtzi} density, the green lines are the infected \textit{D.
+dentifera} density, and the blue lines indicate the infected \textit{D.
+lumholtzi} density.}
 \end{figure}
@@
 The SIRPF model incorporates these factors as quantitative latent state variables, enabling a more comprehensive understanding of the system's dynamics.
-In the meantime, the ecological processes of Daphnia can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
-Utilizing panel data on Daphnia, the PanelPOMP framework becomes a valuable analytical tool \citep{Carles2020}.
+In the meantime, the ecological processes of \textit{Daphnia} can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
+Utilizing panel data on \textit{Daphnia}, the PanelPOMP framework becomes a valuable analytical tool \citep{Carles2020}.
 Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
@@
 Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
-For Daphnia either dentifera or lumholtzi, susceptible, infected Daphnia and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$.
-Some susceptible Daphnia get infected by parasites, transitioning from $S$ to $I$.
-Both susceptible and infected Daphnia contribute to the $R$ state upon death, releasing surviving parasites back into the environment.
+For \textit{Daphnia} either dentifera or lumholtzi, susceptible, infected \textit{Daphnia} and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$.
+Some susceptible \textit{Daphnia} get infected by parasites, transitioning from $S$ to $I$.
+Both susceptible and infected \textit{Daphnia} contribute to the $R$ state upon death, releasing surviving parasites back into the environment.
 And the parasites continue to impact other individuals and get into another cycle.
@@
 Each equation intricately ties the state variables to depict the complex dynamics of the single species Parasitized system.
-Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race Parasitized system by introducing another Daphnia species.
+Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race Parasitized system by introducing another \textit{Daphnia} species.
 This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$)hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows.
@@
 Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
-The flow diagram shows the how two species of Daphnia interact with each other indirectly through $F$ and $P$.
+The flow diagram shows the how two species of \textit{Daphnia} interact with each other indirectly through $F$ and $P$.
 The marginal processes have been shown in SIRPF mode, and SIRPF2 model is the adaption form in mixed-species dynamcis}
@@
 Intraspecific competition among hosts is conceptualized via the Lotka-Volterra framework, with $r_n$ and $r_i$ denoting respective reproductive rates.
-Host density growth, influenced by algal density, is posited to be linear, indicating a direct relationship between food availability and Daphnia reproduction.
+Host density growth, influenced by algal density, is posited to be linear, indicating a direct relationship between food availability and \textit{Daphnia} reproduction.
 Mortality rates of susceptible hosts are denoted as $\theta_{S_n}$ and $\theta_{S_i}$.
@@
 The model also incorporates a constant removal rate $\delta$, representing the loss of hosts, spores, and food due to external sampling.
-The constraint of food supply, indicated by a constant rate $\mu = 0.37 \times 10^5$ cells per day, is based on observed declines in Daphnia populations in the panel data, suggesting a pivotal role of food availability in driving population dynamics.\\
+The constraint of food supply, indicated by a constant rate $\mu = 0.37 \times 10^5$ cells per day, is based on observed declines in \textit{Daphnia} populations in the panel data, suggesting a pivotal role of food availability in driving population dynamics.\\
 
@@
 \caption{A flow diagram for the SF and SF2 model.
-In this system, both types of Daphnia and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$.
+In this system, both types of \textit{Daphnia} and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$.
 As time progresses, individuals in $F$, $S_n$, and $S_i$ transition to the $R$ state, signifying mortality.
@@
 While if we only keep one species in SF2 model, the model will become SF model.
-Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of Daphnia, particularly under conditions of potential food scarcity.
-This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of Daphnia.
-Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the Daphnia population, thereby suggesting food limitation as a primary driver \citep{xiaotong2018}.
-Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of Daphnia.
-In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of Daphnia.\\
+Utilizing the proposed model facilitates a rigorous way into the influence of algal density on the population dynamics of \textit{Daphnia}, particularly under conditions of potential food scarcity.
+This analytical approach allows for the testing of a specific hypothesis: the critical role of algal density in determining the population trajectory of \textit{Daphnia}.
+Should the model demonstrate a substantial fit to the observed data, it would substantiate the hypothesis that algal density is a pivotal factor in the observed decline of the \textit{Daphnia} population, thereby suggesting food limitation as a primary driver \citep{xiaotong2018}.
+Conversely, a poor fit of the model to the data may indicate that factors other than algal density predominantly govern the population dynamics of \textit{Daphnia}.
+In such a scenario, it becomes imperative to explore alternative variables that could significantly influence the population density of \textit{Daphnia}.\\
 
 \subsection{Measurement and Initial condition}
-In the ecological study of Daphnia populations, we employ the negative binomial distribution to address the inherent overdispersion observed in population counts.
+In the ecological study of \textit{Daphnia} populations, we employ the negative binomial distribution to address the inherent overdispersion observed in population counts.
 This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \citep{ning2021systemic}.
-In our model, the parameter $F$ represents the number of successful Daphnia reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of Daphnia per liter per sampling time in each trial.
-This framework aligns with the observed dynamics in Daphnia populations, where variability in reproduction rates, influenced by fluctuating food availability, leads to overdispersed population counts.
+In our model, the parameter $F$ represents the number of successful \textit{Daphnia} reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of \textit{Daphnia} per liter per sampling time in each trial.
+This framework aligns with the observed dynamics in \textit{Daphnia} populations, where variability in reproduction rates, influenced by fluctuating food availability, leads to overdispersed population counts.
 The negative binomial distribution, therefore, provides a robust statistical tool to assess population dynamics, accommodating the irregularities in reproduction and survival patterns shaped by environmental constraints.\\
 
-The model's effectiveness in capturing the Daphnia population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
-As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing Daphnia populations, particularly under varying food conditions.
+The model's effectiveness in capturing the \textit{Daphnia} population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
+As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing \textit{Daphnia} populations, particularly under varying food conditions.
 
@@
 In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality.
-For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of Daphnia, and 8 for the rest dynamics.\\
+For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of \textit{Daphnia}, and 8 for the rest dynamics.\\
 
@@
 This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.
-The survivability ratio of native and invasive Daphnia in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectivly.
-While if we mix two species of Daphnia into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native Daphnia and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species.
-In such scenario, we observe significant fluctuations in the population density of Daphnia.
+The survivability ratio of native and invasive \textit{Daphnia} in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectivly.
+While if we mix two species of \textit{Daphnia} into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native \textit{Daphnia} and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species.
+In such scenario, we observe significant fluctuations in the population density of \textit{Daphnia}.
 These oscillations are likely attributable to the constraints imposed by limited food resources within the environment.
@@
 The observed variability in the survivability ratio can be understood within this assumption.
-At the same time, it is important to note that the relative survivability between the two Daphnia species studied remains consistent, despite an overall simultaneous decrease in their population densities.
+At the same time, it is important to note that the relative survivability between the two \textit{Daphnia} species studied remains consistent, despite an overall simultaneous decrease in their population densities.
 This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them.
-And similar results can be found in the Parasitized dynamics of invasive species of Daphnia, where the survivability ratio of susceptible native and invasive Daphnia are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.\\
+And similar results can be found in the Parasitized dynamics of invasive species of \textit{Daphnia}, where the survivability ratio of susceptible native and invasive \textit{Daphnia} are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.\\
 
 These findings underscore the importance of parameter sharing in ecological modeling when multiple species or units are involved.
-Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between Daphnia populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones.
+Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between \textit{Daphnia} populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones.
 The results further reinforce the utility of the AIC in guiding the selection of models that not only provide a good fit to the data but also maintain a degree of simplicity, avoiding overfitting.
@@
 The SIRPF model is a powerful tool for analyzing ecological data, as it can be used to estimate parameters and predict future population dynamics.
-The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia.
+The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of \textit{Daphnia}.
 The model's ability to estimate parameters and predict future population dynamics is a powerful tool for analyzing ecological data.
-The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of Daphnia.\\
+The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of \textit{Daphnia}.\\
 

### 2024-04-03 si/si.Rnw
Update

@@
 \usepackage{Sweave}
+\usepackage{tikz}
+\usepackage{float} % For [H] figure placement option
+\usetikzlibrary{arrows.meta, shapes.geometric, positioning}
 \bibliographystyle{apalike}
-\newcommand{\blind}{0}
+\newcommand{\blind}{1}
 
@@
 {
-  \title{\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
-  \author{Author 1\thanks{
-    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
-    Department of YYY, University of XXX\\
-    and \\
-    Author 2 \\
-    Department of ZZZ, University of WWW}
+  \title{\bf Supplement to "Mechanistic models for panel data: analysis of ecological experiments with four interacting species"}
+\author{Bo Yang\\Department of Biostatistics, University of Michigan \\
+        Jesse Wheeler\\Department of Statistics, University of Michigan \\
+        Aaron A. King\\Department of Ecology and Evolutionary Biology \& \\ Center for the Study of Complex Systems, University of Michigan\\
+        Edward L. Ionides\\Department of Statistics, University of Michigan}
   \maketitle
@@
   \begin{center}
-    {\LARGE\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species}
+    {\LARGE\bf Supplement to "Mechanistic models for panel data: analysis of ecological experiments with four interacting species"}
 \end{center}
@@
 
-\bigskip
 <<Setup, include=FALSE,echo=FALSE,results=hide>>=
-library(dplyr)
+ library(plyr)
+ library(reshape2)
+library(patchwork)
+ library(magrittr)
+ library(foreach)
+ library(readxl)
+ library(doParallel)
+ registerDoParallel(cores=36)
+ library(pomp)
+ library(panelPomp)
  library(ggplot2)
-@
+ library(gridExtra)
 
-\section{Simulation}
-\label{sec:sim}
-\begin{figure}[H]
-\centering
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
-       x = "Day",
-       y = "Daphnia dentifera adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF2 Model",
-       x = "Day",
-       y = "Daphnia lumholtzi adult density")
-@
-\end{subfigure}%
-\caption{Plot of Simulation of Non-Parasitize Dyanmics with SF2 Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Susceptible Daphnia dentifera adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Susceptible Daphnia lumholtzi adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = iymini, ymax = iymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Infected Daphnia lumholtzi adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = iymin, ymax = iymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF2 Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
-\end{subfigure}%
-\caption{Plot of Simulation of Parasitize Dyanmics with SIRPF2 Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = lumadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = lumadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Susceptible Daphnia lumholtzi adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = luminf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = luminf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Infected Daphnia lumholtzi adult density")
-@
-\end{subfigure}%
-\caption{Plot of Simulation of Invasive Parasitize Dyanmics with SIRPF Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Susceptible Daphnia dentifera adult density")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymini, ymax = ymaxi, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentinf), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SIRPF Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
-\end{subfigure}%
-\caption{Plot of Simulation of Native Parasitize Dyanmics with SIRPF Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentadult, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
-       x = "Day",
-       y = "Daphnia dentifera adult density")
-@
-\caption{Plot of Simulation of Native Non-Parasitize Dyanmics with SF Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-<<echo=FALSE, fig.cap="Plot2", fig.align='center', fig=TRUE, width=14, height=4>>=
-load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN.RData")
-ggplot() +
-  # geom_line(data = all_sims, aes(y = dentinf, x = day, group = .id)) +
-  geom_line() +
-  geom_ribbon(data = max_min_sims, aes(ymin = ymin, ymax = ymax, x = day), fill = "blue", alpha = 0.2) +
-  facet_wrap(~unit, scales = "free_y") +
-  geom_line(data = data_df, aes(x = day, y = dentadult), col = 'red')+
-  theme_minimal() +
-  labs(title = "Simulation of Non-Parasitize Dyanmics with SF Model",
-       x = "Day",
-       y = "Infected Daphnia dentifera adult density")
-@
-\caption{Plot of Simulation of Invsaive Non-Parasitize Dyanmics with SF Model.
-The blue shadow implies the possible range from simulation, while the red line shows the real-world data.
-The various trail represents different buckets in the experiment}
-\end{figure}
-
-\newpage
-\section{Tables}
-<<Setup2, include=FALSE,echo=FALSE,results=hide>>=
-library(dplyr)
-library(ggplot2)
  myround <- function (x, digits = 1) {
@@
 
-# SF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SF2/SF2_ll.RData")
-# SFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_native/SFN_ll.RData")
-# SFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SF_invasive/SFI_ll.RData")
-# SIRPF2_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF2/SIRPF2_ll.RData")
-# SIRPFN_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_native/SIRPFN_ll.RData")
-# SIRPFI_ll = load("/Users/ybb/Desktop/Daphnia-article/SIRPF_invaive/SIRPFI_ll.RData")
-load("/Users/ybb/Desktop/Daphnia-article/SIRPF_model_estimate.RData")
+ #---------------------------Host Daphnia----------------------------------------------------------------------
+ #----------------load data---------------------------------------------------------------
+load("data/raw_data.RData")
+ #-----------------------------Host with para----------------------------------------------------------------------------
+ host_para <- Mesocosm_data[101:180, ]
+ host_para <- subset(host_para, select = c(rep, day, dent.adult,dent.inf ))
+ host_para <- host_para[80: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ host_para$day = (host_para$day - 1) * 5 + 7
+ host_para_data = list()
+ trails = c("K","L","M","N","O","P","Q","S")
+ for (i in 1: length(trails)){
+   host_para_data[[i]]=subset(host_para, select = c("day", "dent.adult","dent.inf"),
+                              host_para$rep == trails[i])
+ }
+
+ host_para_plot <- ggplot(host_para,aes(x=day,y=dentadult))+
+   geom_line(data=host_para,aes(x=day,y=dent.adult,color = "Susceptible")) +
+   geom_line(data=host_para,aes(x=day,y=dent.inf,color = "Infected")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia dentifera adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Host without para----------------------------------------------------------------------------
+ host_nopara <- Mesocosm_data[1:100, ]
+ host_nopara <- subset(host_nopara, select = c(rep, day, dent.adult))
+ host_nopara <- host_nopara[100: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ host_nopara$day = (host_nopara$day - 1) * 5 + 7
+ host_nopara_data = list()
+ trails = c("A","B","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   host_nopara_data[[i]]=subset(host_nopara, select = c("day", "dent.adult"),
+                                host_nopara$rep == trails[i])
+ }
+
+ host_nopara_plot <- ggplot(host_nopara,aes(x=day,y=dentadult))+
+   geom_line(data=host_nopara,aes(x=day,y=dent.adult)) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia dentifera adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Invasive with para----------------------------------------------------------------------------
+ invasive_para <- Mesocosm_data_invasive[101:190, ]
+ invasive_para <- subset(invasive_para, select = c(rep, day, lum.adult,lum.adult.inf))
+ invasive_para <- invasive_para[90: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ invasive_para$day = (invasive_para$day - 1) * 5 + 7
+ invasive_para_data = list()
+ trails = c("L","M","N","O","P","Q","R","S","T")
+ for (i in 1: length(trails)){
+   invasive_para_data[[i]]=subset(invasive_para, select = c("day", "lum.adult","lum.adult.inf"),
+                                  invasive_para$rep == trails[i])
+ }
+
+ invasive_para_plot <- ggplot(invasive_para,aes(x=day,y=lumadult))+
+   geom_line(data=invasive_para,aes(x=day,y=lum.adult,color = "Susceptible")) +
+   geom_line(data=invasive_para,aes(x=day,y=lum.adult.inf,color = "Infected")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia lumholtzi adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Invasive without para----------------------------------------------------------------------------
+ invasive_nopara <- Mesocosm_data_invasive[1:100, ]
+ invasive_nopara <- subset(invasive_nopara, select = c(rep, day, lum.adult))
+ invasive_nopara <- invasive_nopara[100: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ invasive_nopara$day = (invasive_nopara$day - 1) * 5 + 7
+ invasive_nopara_data = list()
+ trails = c("A","B","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   invasive_nopara_data[[i]]=subset(invasive_nopara, select = c("day", "lum.adult"),
+                                    invasive_nopara$rep == trails[i])
+ }
+
+ invasive_nopara_plot <- ggplot(invasive_nopara,aes(x=day,y=lumadult))+
+   geom_line(data=invasive_nopara,aes(x=day,y=lum.adult)) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia lumholtzi adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Mix with para----------------------------------------------------------------------------
+ mixed_para <- Mesocosm_data_mixed[91:170, ]
+ mixed_para <- subset(mixed_para, select = c(rep, day, dent.adult,dent.inf,lum.adult,lum.adult.inf))
+ mixed_para <- mixed_para[80: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ mixed_para$day = (mixed_para$day - 1) * 5 + 7
+ mixed_para_data = list()
+ trails = c("K","L","M","N","O","P","Q","S")
+ for (i in 1: length(trails)){
+   mixed_para_data[[i]]=subset(mixed_para, select = c("day", "dent.adult","dent.inf","lum.adult","lum.adult.inf"),
+                               mixed_para$rep == trails[i])
+ }
+
+ mixed_para_plot <- ggplot()+
+   geom_line(data=mixed_para,aes(x=day,y=lum.adult,color = "Susceptible lumholtzi"))+
+   geom_line(data=mixed_para,aes(x=day,y=lum.adult.inf,color = "Infected lumholtzi")) +
+   geom_line(data=mixed_para,aes(x=day,y=dent.adult,color = "Susceptible dentifera")) +
+   geom_line(data=mixed_para,aes(x=day,y=dent.inf,color = "Infected dentifera")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #-----------------------------Mix without para----------------------------------------------------------------------------
+ mixed_nopara <- Mesocosm_data_mixed[1:90, ]
+ mixed_nopara <- subset(mixed_nopara, select = c(rep, day, dent.adult,lum.adult))
+ mixed_nopara <- mixed_nopara[90: 1, ]
+ #Convert sampling date to natural date. Samples are collected every 5 days after the #first 7 days.
+ mixed_nopara$day = (mixed_nopara$day - 1) * 5 + 7
+ mixed_nopara_data = list()
+ trails = c("A","C","D","E","F","G","H","I","J")
+ for (i in 1: length(trails)){
+   mixed_nopara_data[[i]]=subset(mixed_nopara, select = c("day", "dent.adult","lum.adult"),
+                                 mixed_nopara$rep == trails[i])
+ }
+
+ mixed_nopara_plot <- ggplot()+
+   geom_line(data=mixed_nopara,aes(x=day,y=lum.adult,color = "Lumholtzi"))+
+   geom_line(data=mixed_nopara,aes(x=day,y=dent.adult,color = "Dentifera")) +
+   facet_wrap(~rep,scales="free")+labs(y = "Daphnia adult density")
+ #----------------------------------------------------------------------------------------------------------------------
+ #----------Only lum Model-----------------------------------------------------------------------------------------------
+ invasive_nopara_rpro <- Csnippet("
+                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
+                      double noiB, noiG ;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiB = rnorm(0, sigB * sqrt(dt));
+                      noiG = rnorm(0, sigG * sqrt(dt));
+
+                      //------------F-------------
+
+                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
+                      Fpred =  S * F * (Beta * dt);
+                      Fspl = F * (delta * dt);
+                      Ffill = 0.37 * dt;
+
+                      //------------S--------------
+                      Sgrow = S * F * (Theta * dt) + S * noiG;
+                      Sdeath = S * (gamma * dt);
+                      Sspl = S * (delta * dt);
+
+                      //M is the total Daphnia sample out
+                      M += Sspl;
+                      F += Fbirth - Fpred + Ffill - Fspl;
+                      S += Sgrow - Sdeath  - Sspl;
+
+                      if (S <= 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F <= 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      ")
+ invasive_nopara_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     M = 0;
+                     error_count = 0;
+                     ")
+ invasive_nopara_dmeas = Csnippet("
+                 double delta = 0.013;
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = log(tol);
+                 } else {
+                   lik = dpois(lumadult,M,give_log);
+                 }
+                 ")
+ invasive_nopara_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 lumadult = rpois(M);
+                 ")
+ invasive_nopara_pt <- parameter_trans(
+   log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
+ )
+ invasive_nopara_pomplist=list()
+ #Not sure about the consumption rate of alga
+ invasive_nopara_parameters = c(1, 0.0002500899,40,2, 10,0.2496332878,0.3)
+ names(invasive_nopara_parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")
+
+ for (i in 1:10){
+   colnames(invasive_nopara_data[[i]]) <- c('day', 'lumadult')
+   pomp(data = invasive_nopara_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(invasive_nopara_rpro,delta.t=1/4),
+        rinit = invasive_nopara_init,
+        dmeasure = invasive_nopara_dmeas,
+        rmeasure = invasive_nopara_rmeas,
+        partrans = invasive_nopara_pt,
+        obsnames = "lumadult",
+        accumvars = c("error_count","M"),
+        paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"),
+        statenames = c("S", "error_count", "F","M")
+   ) -> invasive_nopara_pomplist[[i]]
+   coef(invasive_nopara_pomplist[[i]])=invasive_nopara_parameters
+ }
+ names(invasive_nopara_pomplist)=paste("u", 1:10,sep = "")
+
+ invasive_nopara_shared_parameter =        c(0.0002500899 ,40   ,10    ,0.2496332878,  1, 2)
+ names(invasive_nopara_shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")
+
+ invasive_nopara_specific_mat = matrix(data = c(rep(0.3, 10)),
+                                       nrow = 1, nc = 10, byrow = T,
+                                       dim = list(rownames = c("sigG"),
+                                                  colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))
+
+ invasive_nopara_panelfood = panelPomp(invasive_nopara_pomplist, shared=invasive_nopara_shared_parameter,specific = invasive_nopara_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------Only dent Model-----------------------------------------------------------------------------------------------
+ host_nopara_rpro <- Csnippet("
+                      double Fbirth, Fpred, Fspl, Ffill, Sgrow, Sdeath, Sspl;
+                      double noiB, noiG ;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiB = rnorm(0, sigB * sqrt(dt));
+                      noiG = rnorm(0, sigG * sqrt(dt));
+
+                      //------------F-------------
+
+                      Fbirth = F * (alpha) * (1 - F/kf) * dt + F * noiB;
+                      Fpred =  S * F * (Beta * dt);
+                      Fspl = F * (delta * dt);
+                      Ffill = 0.37 * dt;
+
+                      //------------S--------------
+                      Sgrow = S * F * (Theta * dt) + S * noiG;
+                      Sdeath = S * (gamma * dt);
+                      Sspl = S * (delta * dt);
+
+                      //M is the total Daphnia sample out
+                      M += Sspl;
+                      F += Fbirth - Fpred + Ffill - Fspl;
+                      S += Sgrow - Sdeath  - Sspl;
+
+                      if (S <= 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F <= 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      ")
+ # Initial state. Assume t0 = day 4
+ host_nopara_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     M = 0;
+                     error_count = 0;
+                     ")
+
+ host_nopara_dmeas = Csnippet("
+                 double delta = 0.013;
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = log(tol);
+                 } else {
+                   lik = dpois(dentadult,M,give_log);
+                 }
+                 ")
+
+ host_nopara_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 dentadult = rpois(M);
+                 ")
+
+ host_nopara_pt <- parameter_trans(
+   log = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf")
+ )
+
+ host_nopara_pomplist=list()
+
+ #Not sure about the consumption rate of alga
+ host_nopara_parameters = c(0.5, 0.0005,0.03,0.1, 8,0.6,0.3)
+ names(host_nopara_parameters) =  c("alpha","Beta", "gamma", "Theta","kf","sigB","sigG")
+
+ for (i in 1:10){
+   colnames(host_nopara_data[[i]]) <- c('day', 'dentadult')
+   pomp(data = host_nopara_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(host_nopara_rpro,delta.t=1/4),
+        rinit = host_nopara_init,
+        dmeasure = host_nopara_dmeas,
+        rmeasure = host_nopara_rmeas,
+        partrans = host_nopara_pt,
+        obsnames = "dentadult",
+        accumvars = c("error_count","M"),
+        paramnames = c("alpha", "Beta", "Theta", "gamma", "sigB", "sigG", "kf"),
+        statenames = c("S", "error_count", "F","M")
+   ) -> host_nopara_pomplist[[i]]
+   coef(host_nopara_pomplist[[i]])=host_nopara_parameters
+ }
+ names(host_nopara_pomplist)=paste("u", 1:10,sep = "")
+
+ host_nopara_shared_parameter =        c(7.773061e-05  ,2.052644e-01   ,1.790036    ,6.940058e-03  ,1.170719e-01, 0.15)
+ names(host_nopara_shared_parameter) = c("Beta","gamma","kf","sigB","alpha", "Theta")
+
+ host_nopara_specific_mat = matrix(data = c(rep(0.3, 10)),
+                                   nrow = 1, nc = 10, byrow = T,
+                                   dim = list(rownames = c("sigG"),
+                                              colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9","u10")))
+
+ host_nopara_panelfood = panelPomp(host_nopara_pomplist, shared=host_nopara_shared_parameter,specific = host_nopara_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------dent par Model----------------------------------------------
+ dent_par_rpro <- Csnippet("
+                      double S_term, I_term, F_term, P_term;
+                      double noiS, noiI, noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiS = rnorm(0, sigS * sqrt(dt));
+                      noiI = rnorm(0, sigI * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------s-------------
+                      //Model without death part
+                      S_term = r  * F * (S + zeta * I) * dt - theta_S * S * dt - prob * f_S * S * P * dt - delta * S * dt + S * noiS;
+
+                      //------------I--------------
+
+                      I_term = prob * f_S * S * P * dt - theta_I * I *dt - delta * I * dt + I * noiI;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamma * F * (S + zeta * I) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = beta * theta_I * I * dt - f_S * S * P * dt - f_I * I * P * dt - theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      S += S_term;
+                      I += I_term;
+                      P += P_term;
+
+                      T_D += delta * S * dt;
+                      T_I += delta * I * dt;
+
+                      //Trace time
+
+                     if (t - 4.0 < 0.000000001 && t - 4.0 > -0.000000001){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                      if (S < 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (I < 0) {
+                        I = 0;
+                        error_count += 0.001;
+                      }
+                      if (P < 0 && t > 3.9) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      ")
+
+ # Initial state. Assume t0 = day 4
+ dent_par_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     T_D = 0;
+                     T_I = 0;
+                     I = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ dent_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                  if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(dentadult,k_S,T_D,give_log) +  dnbinom_mu(dentinf,k_I,T_I,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(dentadult,k_S,T_D,give_log) *  dnbinom_mu(dentinf,k_I,T_I,give_log);
+                    }
+                      }
+
+                 ")
+
+ dent_par_rmeas = Csnippet("
+                 dentadult = rnbinom_mu(k_S,T_D);
+                 dentinf = rnbinom_mu(k_I,T_I);
+                 ")
+
+ dent_par_pt <- parameter_trans(
+   log = c("alpha", "beta", "gamma", "sigS", "sigI", "sigF","sigP","K_F","f_S","theta_P","r", "f_I","k_S","k_I"),
+   logit = c('prob',"zeta","theta_S","theta_I")
+ )
+
+ dent_par_pomplist=list()
+
+ dent_par_parameters =         c(0.16,   2.8   ,   0.00011,   0.1,    0.1,   0.7,  0.2,     0.01,     0.8,     2.3,  0.001,   0.2,      0.1,  0.06 ,  0.7,   0.003,10,10)
+ names(dent_par_parameters) =  c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I")
+
+ for (i in 1:8){
+   colnames(host_para_data[[i]]) <- c('day', 'dentadult', 'dentinf')
+   pomp(data = host_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(dent_par_rpro,delta.t=1/4),
+        rinit = dent_par_init,
+        dmeasure = dent_par_dmeas,
+        rmeasure = dent_par_rmeas,
+        partrans = dent_par_pt,
+        obsnames = c("dentadult", "dentinf"),
+        accumvars = c("error_count","T_D","T_I"),
+        paramnames = c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I"),
+        statenames = c("S",  "I", "error_count", "F", "T_D","T_I","P")
+   ) -> dent_par_pomplist[[i]]
+   coef(dent_par_pomplist[[i]])=dent_par_parameters
+ }
+ names(dent_par_pomplist)=paste("u", 1:8,sep = "")
+
+ shared_dent_par_parameter =        c(0.0115,   0.39 , 9.628e-06,  0.5447 , 0.1004, 0.04401,    0.3,   0.5798,  0.1489,  0.0008096, 0.004265,    0.1432,  0.003244 ,  0.6551,   0.003584,   8,   8)
+ names(shared_dent_par_parameter) = c("alpha", "sigS",  "gamma",  "sigI",  "sigF", "sigP", "theta_S","theta_I","K_F",      "f_S",  "theta_P",     "r",    "zeta",   "prob",     "f_I","k_S","k_I")
+
+ specific_dent_par_mat = matrix(data = c(rep(2.2,8)),
+                                nrow = 1, nc = 8, byrow = T,
+                                dim = list(rownames = "beta",
+                                           colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8")))
+
+ dent_par_panelfood = panelPomp(dent_par_pomplist, shared=shared_dent_par_parameter,specific = specific_dent_par_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+
+ #----------lum par Model----------------------------------------------
+ lum_par_rpro <- Csnippet("
+                      double S_term, I_term, F_term, P_term;
+                      double noiS, noiI, noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiS = rnorm(0, sigS * sqrt(dt));
+                      noiI = rnorm(0, sigI * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------s-------------
+                      //Model without death part
+                      S_term = r  * F * (S + zeta * I) * dt - theta_S * S * dt - prob * f_S * S * P * dt - delta * S * dt + S * noiS;
+
+                      //------------I--------------
+
+                      I_term = prob * f_S * S * P * dt - theta_I * I *dt - delta * I * dt + I * noiI;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamma * F * (S + zeta * I) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = beta * theta_I * I * dt - f_S * S * P * dt - f_I * I * P * dt - theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      S += S_term;
+                      I += I_term;
+                      P += P_term;
+
+                      if (S < 0.0) {
+                        S = 0.0;
+                        error_count += 1;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (I < 0) {
+                        I = 0;
+                        error_count += 0.001;
+                      }
+                      if (P < 0) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      T_D += delta * S * dt;
+                      T_I += delta * I * dt;
+
+                      //Trace time
+
+                     if (t - 4.00 < 0.1 && t - 4.00 > -0.1){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                     // Rprintf(\"P=%f\\n\",P);
+
+                      ")
+
+ lum_par_init = Csnippet("
+                     S = 3; //3= 45/15L
+                     F = 16.667;
+                     T_D = 0;
+                     T_I = 0;
+                     I = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ lum_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                  if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(lumadult,k_S,T_D,give_log) +  dnbinom_mu(luminf,k_I,T_I,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(lumadult,k_S,T_D,give_log) *  dnbinom_mu(luminf,k_I,T_I,give_log);
+                    }
+                      }
+
+                 ")
+
+ lum_par_rmeas = Csnippet("
+                 lumadult = rnbinom_mu(k_S,T_D);
+                 luminf = rnbinom_mu(k_I,T_I);
+                 ")
+
+ lum_par_pt <- parameter_trans(
+   log = c("alpha", "beta", "gamma", "sigS", "sigI", "sigF","sigP","K_F","f_S","theta_P","r", "f_I","k_S","k_I"),
+   logit = c('prob',"zeta","theta_S","theta_I")
+ )
+
+ lum_par_pomplist=list()
+
+ lum_par_parameters =         c(0.16,   2.8   ,   0.00011,   0.1,    0.1,   0.7,  0.2,     0.01,     0.8,     2.3,  0.001,   0.2,      0.1,  0.06 ,  0.7,   0.003,10,10)
+ names(lum_par_parameters) =  c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I")
+
+ for (i in 1:9){
+   colnames(invasive_para_data[[i]]) <- c('day', 'lumadult', 'luminf')
+   pomp(data = invasive_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(lum_par_rpro,delta.t=1/4),
+        rinit = lum_par_init,
+        dmeasure = lum_par_dmeas,
+        rmeasure = lum_par_rmeas,
+        partrans = lum_par_pt,
+        obsnames = c("lumadult", "luminf"),
+        accumvars = c("error_count","T_D","T_I"),
+        paramnames = c("alpha", "beta",  "gamma", "sigS", "sigI", "sigF","sigP","theta_S","theta_I","K_F", "f_S",  "theta_P", "r", "zeta", "prob", "f_I","k_S","k_I"),
+        statenames = c("S",  "I", "error_count", "F", "T_D","T_I","P")
+   ) -> lum_par_pomplist[[i]]
+   coef(lum_par_pomplist[[i]])=lum_par_parameters
+ }
+ names(lum_par_pomplist)=paste("u", 1:9,sep = "")
+
+ lum_par_shared_parameter =        c(0.0115,   0.39 , 9.628e-06,  0.5447 , 0.1004, 0.04401,    0.3,   0.5798,  0.1489,  0.0008096, 0.004265,    0.1432,  0.003244 ,  0.6551,   0.003584,   8,   8)
+ names(lum_par_shared_parameter) = c("alpha", "sigS",  "gamma",  "sigI",  "sigF", "sigP", "theta_S","theta_I","K_F",      "f_S",  "theta_P",     "r",    "zeta",   "prob",     "f_I","k_S","k_I")
+
+ lum_par_specific_mat = matrix(data = c(rep(2.2,9)),
+                               nrow = 1, nc = 9, byrow = T,
+                               dim = list(rownames = "beta",
+                                          colnames=c("u1","u2","u3","u4","u5", "u6","u7","u8","u9")))
+
+ lum_par_panelfood = panelPomp(lum_par_pomplist, shared=lum_par_shared_parameter,specific = lum_par_specific_mat)
+
+ #----------------------------------------------------------------------------------------------------------------------
+ #----------mixed par Model-----------------------------------------------------------------------------------------------
+ mix_par_rpro <- Csnippet("
+                      double Sn_term, In_term, F_term, P_term , Si_term, Ii_term;
+                      double noiSn, noiIn, noiSi , noiIi ,noiF, noiP;
+                      double delta = 0.013; //fraction of volume replaced day-1
+
+                      noiSn = rnorm(0, sigSn * sqrt(dt));
+                      noiIn = rnorm(0, sigIn * sqrt(dt));
+                      noiSi = rnorm(0, sigSi * sqrt(dt));
+                      noiIi = rnorm(0, sigIi * sqrt(dt));
+                      noiF = rnorm(0, sigF * sqrt(dt));
+                      noiP = rnorm(0, sigP * sqrt(dt));
+
+                      //------------Sn-------------
+                      Sn_term = rn  * F * (Sn + xi * In) * dt - theta_Sn * Sn * dt -  probn * f_Sn * Sn * P * dt - delta * Sn * dt + Sn * noiSn;
+
+                      //------------In--------------
+
+                      In_term = probn * f_Sn * Sn * P * dt - theta_In * In *dt - delta * In * dt + In * noiIn;
+
+                      //------------Si-------------
+                      Si_term = ri  * F * (Si + xi * Ii) * dt - theta_Si * Si * dt -  probi * f_Si * Si * P * dt - delta * Si * dt + Si * noiSi;
+
+                      //------------Ii--------------
+
+                      Ii_term = probi * f_Si * Si * P * dt - theta_Ii * Ii *dt - delta * Ii * dt + Ii * noiIi;
+
+                      //-----------F---------------
+
+                      F_term = alpha * F * (1 - F/K_F) * dt + F * noiF - gamman * F * (Sn + xi * In) * dt - gammai * F * (Si + xi * Ii) * dt - delta * F * dt + 0.37 * dt;
+
+                      //----------P---------------
+
+                      P_term = betan * theta_In * In * dt + betai * theta_Ii * Ii * dt - f_Sn * Sn * P * dt - f_In * In * P * dt - f_Si * Si * P * dt - f_Ii * Ii * P * dt- theta_P * P * dt - delta * P * dt + P * noiP;
+
+                      //T_D and T_I are the S and I Daphnia sample out
+                      F += F_term;
+                      Sn += Sn_term;
+                      In += In_term;
+                      Si += Si_term;
+                      Ii += Ii_term;
+                      P += P_term;
+
+                      //Trace time
+
+                     if (t - 4.0 < 0.001 && t - 4.0 > -0.001){
+                      //Initial statement
+                        P += 25;
+                      }
+
+                      if (Sn < 0.0) {
+                        Sn = 0.0;
+                        error_count += 1;
+                      }
+                      if (Si < 0.0) {
+                        Si = 0.0;
+                        error_count += 1000000;
+                      }
+                      if (F < 0) {
+                        F = 0;
+                        error_count += 1000;
+                      }
+                      if (In < 0) {
+                        In = 0;
+                        error_count += 0.001;
+                      }
+                      if (Ii < 0) {
+                        Ii = 0;
+                        error_count += 0.000000001;
+                      }
+                      if (P < 0 && t > 3.9) {
+                        P = 0;
+                        error_count += 0.000001;
+                      }
+
+                      T_Sn += delta * Sn * dt;
+                      T_In += delta * In * dt;
+                      T_Si += delta * Si * dt;
+                      T_Ii += delta * Ii * dt;
+
+                      ")
+
+ # Initial state. Assume t0 = day 4
+ mix_par_init = Csnippet("
+                     Sn = 2.333; //2.3333 = 35/15L
+                     Si = 0.667; //0.667 = 10/15L
+                     F = 16.667;
+                     T_Sn = 0;
+                     T_Si = 0;
+                     T_In = 0;
+                     T_Ii = 0;
+                     In = 0;
+                     Ii = 0;
+
+                     error_count = 0.0;
+
+                     //add 25 P at day 4
+                     P = 0;
+                     ")
+
+ mix_par_dmeas = Csnippet("
+                 double tol = 1e-18;
+
+                 if (error_count > 0.0) {
+                   lik = -100;
+                  }
+                 else{
+                    if(give_log){
+                    lik  = dnbinom_mu(lumadult,k_Si,T_Si,give_log) +  dnbinom_mu(luminf,k_Ii,T_Ii,give_log) + dnbinom_mu(dentadult,k_Sn,T_Sn,give_log) +  dnbinom_mu(dentinf,k_In,T_In,give_log);
+                    }
+                    else{
+                    lik  = dnbinom_mu(lumadult,k_Si,T_Si,give_log) *  dnbinom_mu(luminf,k_Ii,T_Ii,give_log) * dnbinom_mu(dentadult,k_Sn,T_Sn,give_log) *  dnbinom_mu(dentinf,k_In,T_In,give_log);
+                    }
+                      }
+
+                 ")
+
+ mix_par_rmeas = Csnippet("
+                 //double delta = 0.013;
+                 dentadult = rnbinom_mu(k_Sn,T_Sn);
+                 dentinf = rnbinom_mu(k_In,T_In);
+                 lumadult = rnbinom_mu(k_Si,T_Si);
+                 luminf = rnbinom_mu(k_Ii,T_Ii);
+                 ")
+
+ mix_par_pt <- parameter_trans(
+   #Without death part
+   log = c("alpha", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","k_Ii","k_In","k_Sn","k_Si"),
+   logit = c("probn","probi","xi","theta_Sn","theta_In","theta_Si","theta_Ii","theta_P")
+ )
+
+ mix_par_pomplist=list()
+
+ mix_par_parameters =         c(0.16,    0.3,   17,    3  ,    0.0001,  8.34e-06,     0.4,    0.5,    0.8,     0.2,     0.01,  0.2,    0.15,       0.07 ,   0.2,       0.01,    0.096     ,1  ,0.0006,0.001 , 0.0035 ,0.003 , 0.2  , 0.4, 0.6  ,0.8 ,8,8,8,8)
+ names(mix_par_parameters) =  c("alpha","xi", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","theta_Sn","theta_In","theta_Si","theta_P","theta_Ii","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","probn","probi","k_Ii","k_In","k_Sn","k_Si")
+
+ for (i in 1:8){
+   colnames(mixed_para_data[[i]]) <- c('day', 'dentadult', 'dentinf','lumadult','luminf')
+   pomp(data = mixed_para_data[[i]],
+        times = "day",
+        t0=1,
+        rprocess=euler(mix_par_rpro,delta.t=1/4),
+        rinit = mix_par_init,
+        dmeasure = mix_par_dmeas,
+        rmeasure = mix_par_rmeas,
+        partrans = mix_par_pt,
+        obsnames = c("dentadult", "dentinf","lumadult","luminf"),
+        accumvars = c("error_count","T_Sn","T_In","T_Si","T_Ii"),
+        paramnames = c("alpha","xi", "betan", "betai", "gamman", "gammai", "sigSn", "sigIn", "sigSi", "sigIi", "sigF","sigP","theta_Sn","theta_In","theta_Si","theta_P","theta_Ii","K_F","f_Sn","f_Si", "f_In","f_Ii","rn","ri","probn","probi","k_Ii","k_In","k_Sn","k_Si"),
+        statenames = c("Sn",  "In", "Si",  "Ii","error_count", "F", "T_Sn","T_In","T_Si","T_Ii","P")
+   ) -> mix_par_pomplist[[i]]
+   coef(mix_par_pomplist[[i]])=mix_par_parameters
+ }
+ names(mix_par_pomplist)=paste("u", 1:8,sep = "")
+
+ mix_par_shared_parameter = c("alpha" = 0.045,"xi" = 0.0005,"gamman" = 9.4e-05,"gammai" =0.00075, "sigSn" = 0.15, "sigIn" = 0.033, "sigSi" =0.28, "sigIi" = 0.03,
+                              "sigF" = 0.08, "theta_Sn" = 0.19,"theta_In" = 0.62,"theta_Si" = 0.22,"theta_P" = 0.08,"theta_Ii" = 0.48,"K_F" = 0.3,"f_Sn" = 0.001,"f_Si" = 0.0011,"f_In" = 0.0012,"f_Ii" = 0.00023,"rn" = 0.25,
+                              "ri" = 0.29,"probn" = 0.93,"probi" = 0.96,"k_Ii" = 0.38,"k_In" = 0.83,"k_Sn" = 11,"k_Si"= 570,"sigP" = 0.02, "betan" = 2, "betai" = 2)
+
+ mix_par_panelfood = panelPomp(mix_par_pomplist, shared=mix_par_shared_parameter)
 @
-\renewcommand{\arraystretch}{0.5}
-\begin{table}[H]
+
+\tableofcontents
+\newpage
+
+\section{Model Diagram}
+
+Dynamic models addressed in this manuscript are comprehensively delineated through the descriptions provided within the main text, augmented by the supplementary details furnished in this section.
+Notwithstanding their complete mathematical specification, visual diagrams of dynamic systems frequently facilitate comprehension of the underlying equations.
+Hence, this section presents a model diagram, which corresponds to the SF model.
+It is imperative to recognize that, given the models are articulated through their mathematical formulations and numerical implementations, these diagrams do not constitute unique visual depictions.
+There exist alternative visualizations that might further aid in elucidating the models discussed herein.
+
+\begin{figure}[H]
+\begin{center}
+\resizebox{5cm}{!}{%
+\begin{tikzpicture}[
+  square/.style={
+    rectangle,
+    draw=black,
+    minimum width=1cm,
+    minimum height=1cm,
+    rounded corners=.1cm,
+    fill=blue!8
+  },
+  report/.style={
+    shape=regular polygon,
+    regular polygon sides=8,
+    draw,
+    fill=red!8,
+    minimum size=1cm,
+    inner sep=0cm
+  },
+  rhombus/.style={
+    diamond,
+    draw=black,
+    minimum width=0.1cm,
+    minimum height=0.1cm,
+    fill=purple!8,
+    aspect = 1
+  },
+  >={Stealth[length=1.5mm,width=1.5mm]} % Corrected arrow tip definition
+]
+
+\node (S) at (-0.5,0) [square] {$S$};
+\node (F) at (2.5,0) [report] {$F$};
+\node (R) at (2.5,2) [square] {$R$};
+
+\draw [->] (F) --  (S);
+\draw [->] (F) --  (R);
+\draw [->] (S) --  (R);
+\draw (S) edge[loop above] node {} (S);
+\draw (F) edge[loop below] node {} (F);
+
+\end{tikzpicture}%
+}
+\end{center}
+\vspace{-5mm}
+\caption{A flow diagram for the SF model.
+In this system, either species of \textit{Daphnia} and the food resource contribute to reproduction, depicted as self-recycling loops for $S$, and $F$.
+As time progresses, individuals in $F$, $S$ transition to the $R$ state, signifying mortality.}
+\label{fig:SF}
+\end{figure}
+
+\newpage
+
+\section{Markov chain and differential equation interpretations in the dynamics}
+\subsection{SF Model}
+The dynamics of the system are governed by a set of differential equations that describe the interactions between different species and their environment.
+The model consists of equations that represent the growth and interactions of \textit{algae}  (denoted as $F$) and two types of \textit{Daphnia} populations, $S_n$ and $S_i$, under certain conditions.
+
+\begin{align}
+&d S(t)=r S(t) dt- \theta_{S}S(t) dt -\delta S(t)dt + S(t) d\zeta_{S} \\
+&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S(t) \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F\\
+&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
+&\zeta_{S} \sim N \left( 0, \sigma_{S}^2dt  \right)
+\end{align}
+
+This model includes parameters such as $\gamma$, $\alpha$, $\sigma_F$, $\sigma_S$, $\theta$, $\beta$, $\delta$, and $k_f$, which are integral to understanding the dynamics between \textit{algae}  and \textit{Daphnia} populations.
+The parameters $\alpha$ and $K_F$ represent the growth rate and carrying capacity of \textit{algae} , respectively, while $\beta$ and $\theta$ describe the consumption rate of \textit{algae}  by Daphnia and the growth rate of \textit{Daphnia}, respectively.
+The death rate of \textit{Daphnia} is denoted by $\gamma$.
+The terms $\zeta_F$ and $\zeta_S$ introduce randomness into the growth processes of \textit{algae}  and \textit{Daphnia}, modeling the stochastic nature of the ecosystem.
+In this model, we set the initial condition to be $M = 0, S = 3$ and $F = 16.667$ by Searles et al.
+For ease of understanding the model equations, we separate the differential equations (12)-(16)  into distinct parts that each correspond to meaningful ecological phenomena:
+\paragraph{Alga}
+
+The change of the density of alga can be viewed as
+        \begin{flalign}
+        & dF_{birth}(t) = \alpha F(t)\left(1-\frac{F(t)}{k_f}\right) dt + F(t) d\zeta_F\\
+        &dF_{cons}(t) = \beta F(t) S(t) dt\\
+        &dF_{spl}(t) = \delta F(t) dt\\
+        &dF_{refill}(t) = \mu dt\\
+        &dF(t) = dF_{birth}(t) - dF_{cons}(t) - dF_{spl}(t) + dF_{refill}(t)
+        \end{flalign}
+In these equations, $\alpha$ represents the growth rate; $k_f$ is the carrying capacity; $\zeta_F$ is the noise term; and $\beta$ represents the rate at which alga are consumed.
+All of these are treated as unknown parameters.
+$\delta$ and $\mu$ are fixed parameters with $\delta = 0.013$ and $\mu = 0.37$, which model the food supply rate.
+We split the change of alga density (dF) into four parts: equation (17) represents the change of population due to the birth of alga and Brownian noise term; equation (18) is the amount of alga that is consumed by \textit{Daphnia}; equations (19) and (20) indicate that change of alga density due to the process of alga refill and sampling.
+Together, these four equations explain the processes and causes of the changes in alga population density.
+
+\paragraph{\textit{Daphnia}}
+
+The change of the density of \textit{Daphnia} can be viewed as
+        \begin{flalign}
+        &dS_{grow}(t) = \theta S(t) F(t) dt+S(t) d\zeta_S\\
+        &dS_{death}(t) = \gamma S(t)dt\\
+        &dS_{spl}(t) = \delta S(t)dt\\
+        &dS(t) =  S_{grow}(t) - S_{death}(t) - S_{spl}(t)
+        \end{flalign}
+In these equations, $\theta$ is the growth rate, $\zeta_S$ is the noise term, and $\gamma$ is the death rate, all of which are unknown parameters.
+$\delta$ is a fixed parameter with $\delta = 0.013$.
+We split the change of \textit{Daphnia} density (dS) into three parts.
+The equation (22) represents the change of population due to the birth of \textit{Daphnia} and Brownian noise terms.
+While the equation (23) shows the amount of alga that is consumed by \textit{Daphnia}, which will make the density decrease.
+The equation (24) indicates that change of \textit{Daphnia} density due to the process of  sampling.
+These three equations explain the processes and causes of the changes in \textit{Daphnia} population density.\\
+
+This model includes 9 parameters: $\gamma,\alpha,\sigma_F,\sigma_S,\theta,\beta,\delta, \mu$ and $k_f$.
+Among those parameters, $\alpha$ represents the rate of growth of algae.
+$k_f$ shows the carrying capacity of algae in each bucket.
+$\beta$ is the consumption rate of algae by certain species of \textit{Daphnia}, $\theta$ shows the rate of growth of \textit{Daphnia} with certain types of food and $\gamma$ is the rate of death of certain species of \textit{Daphnia}.
+Similar to the previous Lotka-Volterra model, we treat $\delta$ as a constant equal to 0.013 and $\mu = 0.37$ as a constant that implies the food refilling rate according to the experimental settings.
+$\zeta_F$ and $\zeta_S$ are Brownian noise terms, which show the randomness of the dynamic system.
+They are induced by a normal distribution with 0 mean,and $\sigma_F^2 \cdot dt, \sigma_S^2 \cdot dt$ to be variance respectively.
+Unlike the previous model, this model restricts the total food supply, more closely mimicking the experimental setting.
+One explanation for the observed precipitous decline in \textit{Daphnia} population density is that there was a meaningful decline to the food supply throughout the experiment.
+Therefore the model equations permit both a natural decline in food levels and an increase of \textit{Daphnia} food corresponding to the food supply rate of the experiment.
+Based on the experimental manipulations, we treated the food supply rate as a constant $\mu = 0.37 \cdot 10^5$ cells per day.\\
+
+Using this model, we can investigate how the population dynamics of \textit{Daphnia} are influenced by the density of algae with the condition that food levels without any food management may not be sufficient to describe \textit{Daphnia} population dynamics.
+If the data are better explained using this model than the alternative hypotheses, it can be concluded that the density of algae plays a significant role in \textit{Daphnia} population dynamics and the lack of food is potentially the reason of the precipitous decline in \textit{Daphnia} population, as other factors may not obviously affect this treatment.
+Alternatively, if the model does not fit the data well, it's possible that the alga may not be the main reason that led to the change of population of \textit{Daphnia}, which means other possible factors that may affect the population density of \textit{Daphnia} should be considered.\\
+
+Incorporating stochastic elements ($\zeta$) into the model adheres to stochastic differential equation (SDE) frameworks, offering a refined perspective on population dynamics that includes the effect of random environmental variations.
+These components, depicted as Gaussian white noise, add complexity to the model by portraying unpredictable external and internal influences on the populations.
+The variances $\sigma_{S}^{2}$ and $\sigma_{F}^{2}$ serve to measure the intensity of stochastic disturbances on both susceptible and infected population groups, alongside their alimentary resources.
+This methodology surpasses traditional deterministic models, providing a comprehensive and authentic analysis of ecological system dynamics.
+
+\subsection{SF2 Model}
+The following SDEs further explores the dynamics between two types of \textit{Daphnia} populations ($S_n$ and $S_i$) and their food source, \textit{algae} ($F$), incorporating stochastic elements to reflect the randomness in natural processes:
+
+\begin{align}
+&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n} \\
+&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i} \\
+&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n(t) \cdot dt - \gamma_i F(t) S_i(t) \cdot dt \nonumber \\
+&\quad \quad -\delta F(t) \cdot dt + \mu \cdot dt +F(t) d \zeta_F\\
+&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
+&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
+&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right)
+\end{align}
+
+These equations delineate the dynamics of a mixed \textit{Daphnia} population, incorporating growth rates ($r_n$, $r_i$), the impact of \textit{algae} consumption ($\theta_{S_n}$, $\theta_{S_i}$), and mortality rates ($\delta$).
+The stochastic components ($\zeta_{S_n}$, $\zeta_{S_i}$, and $\zeta_F$) introduce environmental variability into the model, rendering a more nuanced depiction of ecosystem dynamics.
+Such analysis illuminates the intricate relationships between species and their habitats, enhancing our comprehension of ecological complexities.
+
+The incorporation of stochastic terms ($\zeta$) aligns with the principles of stochastic differential equations (SDEs), presenting a nuanced view of population dynamics that accommodates for random environmental fluctuations.
+These elements, represented as Gaussian white noise, enrich the model by simulating unpredictable environmental and internal processes that impact the populations.
+The specified variances—$\sigma_{S_n}^{2}, \sigma_{S_n}^{2}$ and $\sigma_{F}^{2}$—quantify the magnitude of noise affecting both the susceptible and infected populations, as well as their food sources.
+This approach transcends the limitations of deterministic models, offering a layered and realistic exploration of the forces shaping ecological systems.
+
+\subsection{SIRPF Model}
+The SIRPF model presented encapsulates the dynamics between susceptible ($S$) and infected ($I$) populations, their food source ($F$), and a parasite population ($P$).
+Each differential equation accounts for various biological and ecological processes, augmented with stochastic terms to capture the inherent randomness of these systems.
+
+{
+\small
+\begin{align}
+&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S}\\
+&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}\\
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F\\
+&d P(t) = \beta \theta_I I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}\\
+&d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right)\\
+&d\zeta_{I} \sim N\left(0, \sigma_{I}^{2}dt\right)\\
+&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right)\\
+&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right)
+\end{align}
+}
+
+This model intricately details the interactions and feedback mechanisms between the populations under study.
+The equations describe the growth or decline of each population, modulated by both deterministic factors, such as reproduction and predation rates, and stochastic factors, captured through the $d\zeta$ terms representing environmental variability and other unpredictable influences.
+The inclusion of stochastic terms ($\zeta$) adheres to the principles of stochastic differential equations (SDEs), offering a realistic portrayal of population dynamics by accounting for random fluctuations.
+These terms, modeled as Gaussian white noise, add depth to the analysis by allowing for the simulation of random processes affecting the populations, thereby extending beyond the deterministic skeleton of classical models.
+The variances $\sigma_{S}^{2}, \sigma_{I}^{2}, \sigma_{F}^{2},$ and $\sigma_{P}^{2}$ encapsulate the intensity of environmental and internal process noise affecting susceptible individuals, infected individuals, food resources, and parasite, respectively.
+
+This model's capacity to intertwine biological realism with statistical rigor makes it a potent tool for understanding complex ecological interactions.
+By examining the system's sensitivity to parameter variations and the stochastic components' impact.
+Furthermore, statistical analysis of model outcomes, such as likelihood-based methods for parameter estimation or Bayesian inference for uncertainty quantification, can provide deeper understanding and predictive power, essential for conservation efforts and the management of biological resources.
+
+\subsection{SIRPF2 Model}
+The SIRPF2 model outlined below incorporates a detailed analysis of the interactions between two \textit{Daphnia} populations (denoted as $S_n$ and $S_i$), their infected counterparts ($I_n$ and $I_i$), algae ($F$), and parasites ($P$).
+This model aims to capture the nuanced dynamics of ecological and epidemiological processes within a shared environment.
+{
+\small
+\begin{align}
+% \begin{split}
+&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}\\
+% \end{split}
+% \\
+&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n}\\
+% \begin{split}
+&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i}\\
+% \end{split}
+% \\
+&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t)\\
+\begin{split}
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
+&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F\\
+\end{split}
+\\
+\begin{split}
+&d P(t) = \beta_n \theta_{I_n} I_n(t)dt + \beta_i \theta_{I_i} I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
+&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}\\
+\end{split}
+\\
+&d\zeta_{S_n} \sim N\left(0, \sigma_{S_n}^{2}dt\right)\\
+&d\zeta_{I_n} \sim N\left(0, \sigma_{I_n}^{2}dt\right)\\
+&d\zeta_{S_i} \sim N\left(0, \sigma_{S_i}^{2}dt\right)\\
+&d\zeta_{I_i} \sim N\left(0, \sigma_{I_i}^{2}dt\right)\\
+&d\zeta_{F} \sim N\left(0, \sigma_{F}^{2}dt\right)\\
+&d\zeta_{P} \sim N\left(0, \sigma_{P}^{2}dt\right)
+\end{align}
+}
+This model is characterized by a series of stochastic differential equations that represent the temporal dynamics of each population and their interactions.
+The terms $r_n$ and $r_i$ denote the intrinsic growth rates of the non-infected \textit{Daphnia} populations, while $\theta_{S_n}$, $\theta_{S_i}$, $\theta_{I_n}$, and $\theta_{I_i}$ represent the respective mortality rates due to natural causes other than disease.
+The interaction between \textit{Daphnia} populations and the algae is modeled by the consumption rates $\gamma_n$ and $\gamma_i$, which also affect the \textit{algae}'s growth dynamics along with its intrinsic growth rate $\alpha$ and carrying capacity $K_F$.\\
+
+In the presented SDE model, the dynamics of susceptible and infected \textit{Daphnia} populations, alongside their \textit{algae} food source and parasites, are intricately modeled to incorporate both deterministic biological interactions and stochastic environmental variability.
+The inclusion of stochastic terms, denoted by $d\zeta$, adheres to a normal distribution with zero mean and variance proportional to $dt$, effectively embedding white noise into the system.
+This statistical approach allows the model to capture the inherent unpredictability of ecological processes, offering a nuanced representation that transcends traditional deterministic models.
+The SDE framework facilitates the exploration of population dynamics under random environmental influences, emphasizing the significance of stochasticity in ecological interactions.
+
+\section{Simulations}
+For the model and parameters swarm that result in the best AIC, we did 20 simulations on each unit of the panel data\.
+Viewing the figure, almost all simulations capture the data well.
+And the estimation of parameters swarm used for this simulation is shown in section Parameter.
+\subsection{SF Models}
+\FloatBarrier
+\begin{figure}[h!]
     \centering
-\resizebox{\textwidth}{!}{\begin{tabular}{||c|c |c|c||}
+    \includegraphics[width=10cm]{data/dent_nopara_sim.png}
+    \caption{Simulation of Daphnia dentifera of SF dynamics}
+    \label{fig:dent_nopara_sim}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/lum_nopara_sim.png}
+    \caption{Simulation of Daphnia lumholtzi of SF dynamics}
+    \label{fig:lum_nopara_sim}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\subsection{SF2 Models}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_nopar.png}
+    \caption{Simulation of Daphnia dentifera of SF2 dynamics}
+    \label{fig:sim_mixed_nopar}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_nopar_si.png}
+    \caption{Simulation of Daphnia lumholtzi of SF2 dynamics}
+    \label{fig:sim_mixed_nopar_si}
+\end{figure}
+\FloatBarrier
+
+\subsection{SIRPF Models}
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/dent_para_s_sim.png}
+    \caption{Simulation of Susceptible Daphnia dentifera of SIRPF dynamics}
+    \label{fig:dent_para_s_sim}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/dent_para_i_sim.png}
+    \caption{Simulation of Infected Daphnia dentifera of SIRPF dynamics}
+    \label{fig:dent_para_i_sim}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/lum_para_s_sim.png}
+    \caption{Simulation of Susceptible Daphnia lumholtzi of SIRPF dynamics}
+    \label{fig:lum_para_s_sim}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/lum_par_i_sim.png}
+    \caption{Simulation of Infected Daphnia lumholtzi of SIRPF dynamics}
+    \label{fig:lum_par_i_sim}
+\end{figure}
+\FloatBarrier
+
+\subsection{SIRPF2 Models}
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_par_sn.png}
+    \caption{Simulation of Susceptible Daphnia dentifera of SIRPF2 dynamics}
+    \label{fig:sim_mixed_par}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_par_si.png}
+    \caption{Simulation of Susceptible Daphnia lumholtzi of SIRPF2 dynamics}
+    \label{fig:sim_mixed_nopar_si}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_par_in.png}
+    \caption{Simulation of Infected Daphnia dentifera of SIRPF2 dynamics}
+    \label{fig:sim_mixed_par_in}
+\end{figure}
+\FloatBarrier
+\FloatBarrier
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm]{data/sim_mixed_par_ii.png}
+    \caption{Simulation of Infected Daphnia lumholtzi of SIRPF2 dynamics}
+    \label{fig:sim_mixed_par_ii}
+\end{figure}
+\FloatBarrier
+\section{Results and Tables}
+\begin{table}[h!]
+\centering
+\begin{tabular}{||c |c |c |c||}
  \hline
-Parameter & Definition &Unit\\[0.5ex]
+Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex]
 \hline\hline
-$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
-$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$&Variable\\
-$F$& Alga density& $10^5 cells \cdot L ^{-1}$&Variable\\
-$P$ & Spore density & $10^3 spores \cdot L ^{-1}$&Variable\\
-$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[15],2)}\\
-$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[20],2)}\\
-$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[21],2)}\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[11],2)} \\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[14],2)} \\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[12],2)} \\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$& \Sexpr{mysignif(SIRPF2_estimate[10],2)} \\
-$\theta_{P}$ & Spore degradation rate&$ day^{-1}$&\Sexpr{mysignif(SIRPF2_estimate[13],2)} \\
-$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[22],2)}\\
-$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[23],2)}\\
-$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[16],2)}\\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[17],2)}\\
-$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[18],2)}\\
-$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(SIRPF2_estimate[19],2)} \\
-$\delta$ & Sampling rate&$day ^{-1} $& constant = 0.013\\
-$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[1],2)}\\
-$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[3],2)}\\
-$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(SIRPF2_estimate[4],2)}\\
-$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & constant = 0.37\\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(SIRPF2_estimate[2],2)}\\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[30],2)}\\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ &\Sexpr{mysignif(SIRPF2_estimate[29],2)}\\
-$\sigma_{S_n}$ & Standard deviation of brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[5],2)}\\
-$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[7],2)}\\
-$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[6],2)}\\
-$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[8],2)}\\
-$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[9],2)}\\
-$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(SIRPF2_estimate[28],2)}\\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[26],2)}\\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[27],2)}\\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[25],2)}\\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ &\Sexpr{mysignif(SIRPF2_estimate[24],2)}\\[1ex]
+All shared    & Sexpr{all_dim}         & \Sexpr{d_all_ll}        & \Sexpr{d_all_ll}   \\
+  $\sigma_S$   &   \Sexpr{one_dim}      & \Sexpr{d_sG_ll}          & \Sexpr{d_sG_ll}  \\
+$\sigma_S,\theta$   & \Sexpr{two_dim}        & \Sexpr{d_sGT_ll}          & \Sexpr{d_sGT_ll}    \\
+  $\sigma_S,\beta$    & \Sexpr{two_dim}        & \Sexpr{d_sGB_ll}         & \Sexpr{d_sGB_ll}    \\
+  $\sigma_S,\alpha$   & \Sexpr{two_dim}        & \Sexpr{d_sGa_ll}         & \Sexpr{d_sGa_ll}  \\
+  $\sigma_S,\gamma$   & \Sexpr{two_dim}        & \Sexpr{d_sGg_ll}          & \Sexpr{d_sGg_ll}  \\
+  $\sigma_S,\sigma_F$ & \Sexpr{two_dim}        & \Sexpr{d_sBsG_ll}          & \Sexpr{d_sBsG_ll}  \\
+$\sigma_S,k_f$      & \Sexpr{two_dim}        & \Sexpr{d_sGk_ll}          & \Sexpr{d_sGk_ll} \\
+  $\alpha, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBa_ll}          & \Sexpr{d_sBa_ll} \\
+  $\alpha, \beta$     & \Sexpr{two_dim}        & \Sexpr{d_aB_ll}            & \Sexpr{d_aB_ll}   \\
+     $\alpha$   & \Sexpr{one_dim}        & \Sexpr{d_a_ll}          & \Sexpr{d_a_ll}  \\
+  $\beta,\gamma$      & \Sexpr{two_dim}        & \Sexpr{d_Bg_ll}         & \Sexpr{d_Bg_ll} \\
+ $\gamma, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBg_ll}          & \Sexpr{d_sBg_ll}  \\
+  $\beta, \sigma_F$   & \Sexpr{two_dim}        & \Sexpr{d_sBB_ll}          & \Sexpr{d_sBB_ll}  \\
+  $\sigma_F$   & \Sexpr{one_dim}        & \Sexpr{d_sB_ll}          & \Sexpr{d_sB_ll}  \\
+  $\alpha,\theta$     & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}         & \Sexpr{d_aT_ll}     \\
+  $\theta, \gamma$    & \Sexpr{two_dim}        & \Sexpr{d_Tg_ll}          & \Sexpr{d_Tg_ll} \\
+  $\alpha,\gamma$     & \Sexpr{two_dim}        & \Sexpr{d_ag_ll}          & \Sexpr{d_ag_ll} \\
+ $\beta$   & \Sexpr{one_dim}        & \Sexpr{d_B_ll}         & \Sexpr{d_B_ll}  \\
+ $k_f, \sigma_F$     & \Sexpr{two_dim}        & \Sexpr{d_sBk_ll}          & \Sexpr{d_sBk_ll}  \\
+  $\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          & \Sexpr{d_TB_ll}  \\
+  $\theta$   & \Sexpr{one_dim}        & \Sexpr{d_T_ll}          & \Sexpr{d_T_ll}    \\
+$\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          &  \Sexpr{d_TB_ll}\\
+  $\alpha, \theta$    & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}          & \Sexpr{d_aT_ll}  \\
+     $k_f$   & \Sexpr{one_dim}        & \Sexpr{d_k_ll}          & \Sexpr{d_k_ll}   \\
+  $\alpha , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_ak_ll}            & \Sexpr{d_ak_ll}   \\
+  $\gamma , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_gk_ll}          & \Sexpr{d_gk_ll} \\
+  $\theta , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_Tk_ll}          & \Sexpr{d_Tk_ll} \\
+  $\beta , k_f$       & \Sexpr{two_dim}        & \Sexpr{d_Bk_ll}         & \Sexpr{d_Bk_ll} \\[1ex]
 \hline
-\end{tabular}}
-\caption{This table shows the units of the parameters of competition SIRPF2 model}
+\end{tabular}
+\caption{This table includes the parameter comparison for SIR model with food management, which was newly introduced fitted to \textit{Daphnia dentifera} with different choices of parameters to be either unit specific or shared}
 \label{Table:1}
 \end{table}
-\renewcommand{\arraystretch}{1}
 
-\bibliography{reference}
+\begin{table}[h!]
+\centering
+\begin{tabular}{||c |c |c |c||}
+ \hline
+Specific parameters & dimension & max log-likelihood & AIC \\ [0.5ex]
+\hline\hline
+All shared    & Sexpr{all_dim}         & \Sexpr{d_all_ll}        & \Sexpr{d_all_ll}   \\
+  $\sigma_S$   &   \Sexpr{one_dim}      & \Sexpr{d_sG_ll}          & \Sexpr{d_sG_ll}  \\
+$\sigma_S,\theta$   & \Sexpr{two_dim}        & \Sexpr{d_sGT_ll}          & \Sexpr{d_sGT_ll}    \\
+  $\sigma_S,\beta$    & \Sexpr{two_dim}        & \Sexpr{d_sGB_ll}         & \Sexpr{d_sGB_ll}    \\
+  $\sigma_S,\alpha$   & \Sexpr{two_dim}        & \Sexpr{d_sGa_ll}         & \Sexpr{d_sGa_ll}  \\
+  $\sigma_S,\gamma$   & \Sexpr{two_dim}        & \Sexpr{d_sGg_ll}          & \Sexpr{d_sGg_ll}  \\
+  $\sigma_S,\sigma_F$ & \Sexpr{two_dim}        & \Sexpr{d_sBsG_ll}          & \Sexpr{d_sBsG_ll}  \\
+$\sigma_S,k_f$      & \Sexpr{two_dim}        & \Sexpr{d_sGk_ll}          & \Sexpr{d_sGk_ll} \\
+  $\alpha, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBa_ll}          & \Sexpr{d_sBa_ll} \\
+  $\alpha, \beta$     & \Sexpr{two_dim}        & \Sexpr{d_aB_ll}            & \Sexpr{d_aB_ll}   \\
+     $\alpha$   & \Sexpr{one_dim}        & \Sexpr{d_a_ll}          & \Sexpr{d_a_ll}  \\
+  $\beta,\gamma$      & \Sexpr{two_dim}        & \Sexpr{d_Bg_ll}         & \Sexpr{d_Bg_ll} \\
+ $\gamma, \sigma_F$  & \Sexpr{two_dim}        & \Sexpr{d_sBg_ll}          & \Sexpr{d_sBg_ll}  \\
+  $\beta, \sigma_F$   & \Sexpr{two_dim}        & \Sexpr{d_sBB_ll}          & \Sexpr{d_sBB_ll}  \\
+  $\sigma_F$   & \Sexpr{one_dim}        & \Sexpr{d_sB_ll}          & \Sexpr{d_sB_ll}  \\
+  $\alpha,\theta$     & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}         & \Sexpr{d_aT_ll}     \\
+  $\theta, \gamma$    & \Sexpr{two_dim}        & \Sexpr{d_Tg_ll}          & \Sexpr{d_Tg_ll} \\
+  $\alpha,\gamma$     & \Sexpr{two_dim}        & \Sexpr{d_ag_ll}          & \Sexpr{d_ag_ll} \\
+ $\beta$   & \Sexpr{one_dim}        & \Sexpr{d_B_ll}         & \Sexpr{d_B_ll}  \\
+ $k_f, \sigma_F$     & \Sexpr{two_dim}        & \Sexpr{d_sBk_ll}          & \Sexpr{d_sBk_ll}  \\
+  $\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          & \Sexpr{d_TB_ll}  \\
+  $\theta$   & \Sexpr{one_dim}        & \Sexpr{d_T_ll}          & \Sexpr{d_T_ll}    \\
+$\theta , \beta$    & \Sexpr{two_dim}        & \Sexpr{d_TB_ll}          &  \Sexpr{d_TB_ll}\\
+  $\alpha, \theta$    & \Sexpr{two_dim}        & \Sexpr{d_aT_ll}          & \Sexpr{d_aT_ll}  \\
+     $k_f$   & \Sexpr{one_dim}        & \Sexpr{d_k_ll}          & \Sexpr{d_k_ll}   \\
+  $\alpha , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_ak_ll}            & \Sexpr{d_ak_ll}   \\
+  $\gamma , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_gk_ll}          & \Sexpr{d_gk_ll} \\
+  $\theta , k_f$      & \Sexpr{two_dim}        & \Sexpr{d_Tk_ll}          & \Sexpr{d_Tk_ll} \\
+  $\beta , k_f$       & \Sexpr{two_dim}        & \Sexpr{d_Bk_ll}         & \Sexpr{d_Bk_ll} \\[1ex]
+\hline
+\end{tabular}
+\caption{This table includes the parameter comparison for SIR model with food management, which was newly introduced fitted to \textit{Daphnia lumholtzi} with different choices of parameters to be either unit specific or shared}
+\label{Table:2}
+\end{table}
+
 \end{document}

### 2024-04-03 ms.Rnw
Update

@@
  mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
-     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
+  cbind(mixed_nopara_data[[i]], Trial = paste("Trial", trails[i]))
 }))
+
+# Combine parasitized data with trial information
 mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
-     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
+  cbind(mixed_para_data[[i]], Trial = paste("Trial", trails[i]))
 }))
 
-p1<- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
-   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
-   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
+# Assuming 'mixed_nopara_data_combined' is already your data frame as shown
+# Create a mapping from Trial letters to bucket numbers
+trial_to_bucket <- function(trial) {
+  # Extract the letter from the trial name
+  letter <- substr(trial, 7, 7)
+  # Convert the letter to a bucket number (A=1, B=2, ..., J=10)
+  bucket_number <- match(letter, LETTERS[1:26])
+  # Return the bucket number prefixed with "Bucket"
+  return(paste("Bucket", bucket_number))
+}
+
+# Apply the function to create a new 'Bucket' column
+mixed_nopara_data_combined$Bucket <- sapply(mixed_nopara_data_combined$Trial, trial_to_bucket)
+mixed_para_data_combined$Bucket <- sapply(mixed_para_data_combined$Trial, trial_to_bucket)
+p1 <- ggplot(mixed_nopara_data_combined, aes(x = day)) +
+  geom_line(aes(y = dent.adult, group = interaction(Bucket, "dent.adult"), color = "Dent Adult"), size = 1) +  # Map color to a factor for legend
+  geom_line(aes(y = lum.adult, group = interaction(Bucket, "lum.adult"), color = "Lum Adult"), size = 1) +  # Map color to a factor for legend
   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
-   facet_grid(. ~ group, scales = "fixed", space = "free") +
-   theme(
-     strip.text = element_text(size=10),
-     axis.text.x=element_text(color="black"),
-     axis.ticks.x=element_line(color="black"),
-     axis.text.y=element_text(color="black"),
-     axis.ticks.y=element_line(color="black"),
-     axis.title.x=element_blank(),
-     axis.title.y=element_blank(),
-     legend.position="none",
-     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
-     strip.background = element_blank(), # Optional: Remove facet label background
-     strip.placement = "outside"
-   ) +
-   labs(title="Daphnia density in Non-parasitized Dynamics") +
-   guides(colour=guide_legend(title="Species"))
-
-p2<- ggplot(mixed_para_data_combined, aes(x=day)) + 
-   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
-   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
-    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
-   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +
-
+  facet_grid(. ~ Bucket, scales = "fixed", space = "free") +
+  theme_minimal() +
+  labs(title = "Daphnia Density in Non-parasitized Dynamics",
+       color = "Species") +
+  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue"),
+                     labels = c("Dent Adult" = "Dent adult", "Lum Adult" = "Lum adult")) +
+  theme(legend.position = "right")  # Adjust the legend position as needed
+
+p2 <- ggplot(mixed_para_data_combined, aes(x = day)) +
+  geom_line(aes(y = dentadult, color = "Dent Adult", group = interaction(Bucket, "dentadult"))) +  # Added color mapping for dentadult
+  geom_line(aes(y = lumadult, color = "Lum Adult", group = interaction(Bucket, "lumadult"))) +  # Ensure consistent color mapping
+  geom_line(aes(y = luminf, color = "Lum Inf", group = interaction(Bucket, "luminf"))) +  # Ensure consistent color mapping
+  geom_line(aes(y = dentinf, color = "Dent Inf", group = interaction(Bucket, "dentinf"))) +  # Ensure consistent color mapping
   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
-   facet_grid(. ~ group, scales = "fixed", space = "free") +
-   theme(
-     strip.text = element_text(size=10),
-     axis.text.x=element_text(color="black"),
-     axis.ticks.x=element_line(color="black"),
-     axis.text.y=element_text(color="black"),
-     axis.ticks.y=element_line(color="black"),
-     axis.title.x=element_blank(),
-     axis.title.y=element_blank(),
-     legend.position="none",
-     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
-     strip.background = element_blank(), # Optional: Remove facet label background
-     strip.placement = "outside"
-   ) +
-   labs(title="Daphnia density in Parasitized Dynamics") +
-   guides(colour=guide_legend(title="Species"))
+  facet_grid(. ~ Bucket, scales = "fixed", space = "free") +
+  theme_minimal() +
+  labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
+  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red"),
+                     labels = c("Dent Adult" = "Dent adult", "Lum Adult" = "Lum adult", "Lum Inf" = "Lum inf", "Dent Inf" = "Dent inf")) +
+  guides(color = guide_legend(title = "Species"))  # Ensure the legend is properly configured
+
+# Arrange plots vertically
 grid.arrange(p1, p2, ncol = 1)
+
 @

### 2024-04-11 ms.Rnw
Update of paper

--- 2024-03-20 ms.Rnw
@@
  library(gridExtra)
+ library(scales)  # For formatting scales
+ library(patchwork)
 
@@
 
-As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both Non-parasitized and Parasitized Dynamics to minimize information bias.
-Figures below illustrates these dynamics, with the figure 1 depicting Non-parasitized Dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
-Figure 2 shows Parasitized Dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation.
-Non-parasitized Dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species.
-And based on this result, the correlation between native host and invasive host in Parasitized Dynamics will be further explored.
+As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both non-parasitized and prarasitized dynamics to minimize information bias.
+Figures below illustrates these dynamics, with the first row of Figure 1 depicting non-parasitized dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
+The second row of Figure 1 shows parasitized dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation.
+Non-parasitized dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species.
+And based on this result, the correlation between native host and invasive host in parasitized dynamics will be further explored.
 
@@
 \centering
-<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=16, height=10>>=  
+<<echo=FALSE, fig.cap="Plot1", fig=TRUE, width=25, height=10,out.width='\\textwidth'>>=
 mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
-     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
+  cbind(mixed_nopara_data[[i]], Trial = paste("Trial", trails[i]))
 }))
+mixed_nopara_data_combined$Bucket = paste0(rep("Bucket ",80),rep(1:9, each = 10))
+
 mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
-     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
+  cbind(mixed_para_data[[i]], Trial = paste("Trial", trails[i]))
 }))
 
-p1<- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
-   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
-   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
-   scale_x_continuous(limits = c(0, 52)) +
-   scale_y_continuous(limits = c(0, 175)) +
-   facet_grid(. ~ group, scales = "fixed", space = "free") +
-   theme(
-     strip.text = element_text(size=10),
-     axis.text.x=element_text(color="black"),
-     axis.ticks.x=element_line(color="black"),
-     axis.text.y=element_text(color="black"),
-     axis.ticks.y=element_line(color="black"),
-     axis.title.x=element_blank(),
-     axis.title.y=element_blank(),
-     legend.position="none",
-     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
-     strip.background = element_blank(), # Optional: Remove facet label background
-     strip.placement = "outside"
-   ) +
-   labs(title="Daphnia density in Non-parasitized Dynamics") +
-   guides(colour=guide_legend(title="Species"))
-
-p2<- ggplot(mixed_para_data_combined, aes(x=day)) + 
-   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
-   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
-    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
-   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +
-
-   scale_x_continuous(limits = c(0, 52)) +
-   scale_y_continuous(limits = c(0, 175)) +
-   facet_grid(. ~ group, scales = "fixed", space = "free") +
-   theme(
-     strip.text = element_text(size=10),
-     axis.text.x=element_text(color="black"),
-     axis.ticks.x=element_line(color="black"),
-     axis.text.y=element_text(color="black"),
-     axis.ticks.y=element_line(color="black"),
-     axis.title.x=element_blank(),
-     axis.title.y=element_blank(),
-     legend.position="none",
-     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
-     strip.background = element_blank(), # Optional: Remove facet label background
-     strip.placement = "outside"
-   ) +
-   labs(title="Daphnia density in Parasitized Dynamics") +
-   guides(colour=guide_legend(title="Species"))
+mixed_para_data_combined$Bucket = paste0(rep("Bucket ",80),rep(1:8, each = 10))
+
+bucket_labeller <- as_labeller(function(bucket) gsub("Bucket ", "Bucket-", bucket))
+# Adjusting p1 with specified modifications
+library(ggplot2)
+
+# Adjusting p1 with specified font size modifications
+p1 <- ggplot(mixed_nopara_data_combined, aes(x = day, y = sqrt(dent.adult))) +
+  geom_line(aes(color = "Dent Adult"), size = 1) +
+  geom_line(aes(y = sqrt(lum.adult), color = "Lum Adult"), size = 1) +
+  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_text(size = 10),  # Y-axis title
+        axis.text.x = element_text(size = 10),   # X-axis text
+        axis.text.y = element_text(size = 10),   # Y-axis text
+        legend.title = element_text(size = 10),  # Legend title
+        legend.text = element_text(size = 10),   # Legend text
+        strip.text = element_text(size = 10),    # Facet label text
+        strip.background = element_blank(),
+        strip.placement = "outside") +
+  ylab('Square root of Daphnia density') +
+  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
+  labs(title = "Daphnia Density in Non-parasitized Dynamics", color = "Species") +
+  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue")) +
+  theme_minimal() +
+  guides(color = guide_legend(title = "Species"))
+
+# Adjusting p2 with specified font size modifications
+p2 <- ggplot(mixed_para_data_combined, aes(x = day, y = sqrt(dentadult))) +
+  geom_line(aes(color = "Dent Adult"), size = 1) +
+  geom_line(aes(y = sqrt(lumadult), color = "Lum Adult"), size = 1) +
+  geom_line(aes(y = sqrt(luminf), color = "Lum Inf"), size = 1) +
+  geom_line(aes(y = sqrt(dentinf), color = "Dent Inf"), size = 1) +
+  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
+  theme(axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 10),
+        strip.text = element_text(size = 10),
+        strip.background = element_blank(),
+        strip.placement = "outside") +
+  ylab('Square root of Daphnia density') +
+  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
+  labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
+  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red")) +
+  theme_minimal() +
+  guides(color = guide_legend(title = "Species"))
+
 grid.arrange(p1, p2, ncol = 1)
 @
+% \includegraphics[width=1\textwidth]{"Density_plot.png"} 
+% \label{fig:plot1}
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}.
-The trail A to J represents different buckets in different experiment.
+The Buckets 1 to 9 represents different buckets in different experiment.
 The black lines represent the \textit{D.
@@
 
-To analyze the dynamics of Parasitized and Non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \citep{weiss2013sir}, is utilized.
+To analyze the dynamics of parasitized and non-parasitized systems, the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model, an adaptation of the SIR model \citep{weiss2013sir}, is utilized.
 This model integrates the aspects of food and parasites, unlike the classical SIR model, to more accurately reflect the interactions between native and invasive hosts in the context of resource availability and parasitic influence.
@@
 In the meantime, the ecological processes of \textit{Daphnia} can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
-Utilizing panel data on \textit{Daphnia}, the PanelPOMP framework becomes a valuable analytical tool \citep{Carles2020}.
+Utilizing panel data on \textit{Daphnia}, the PanelPOMP framework becomes a valuable analytic tool \citep{Carles2020}.
 Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
 
-The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the following graph and model show:
+The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the Figure 2 and model show:
 
@@
 \begin{align}
-&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S},\\
-&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I},\\
-&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
-&d P(t) = \beta \theta_I I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
+&d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S}\\
+&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}\\
+&dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F\\
+&d P(t) = \beta \theta_I I(t)dt - f_{S} S(t) P(t)dt - f_{I} I(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}\\
 &d\zeta_{S} \sim N\left(0, \sigma_{S}^{2}dt\right),\\
@@
 
-Each equation intricately ties the state variables to depict the complex dynamics of the single species Parasitized system.
-Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race Parasitized system by introducing another \textit{Daphnia} species.
+Each equation intricately ties the state variables to depict the complex dynamics of the single species parasitized system.
+Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race parasitized system by introducing another \textit{Daphnia} species.
 This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$)hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows.
@@
 \caption{A flow diagram for the SIRPF2 model.
-Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
+Apart from the quantified states above, the graph includes $R$ state, which represents death status.
 The flow diagram shows the how two species of \textit{Daphnia} interact with each other indirectly through $F$ and $P$.
-The marginal processes have been shown in SIRPF mode, and SIRPF2 model is the adaption form in mixed-species dynamcis}
+The marginal processes have been shown in SIRPF mode, and SIRPF2 model is the adaption form in mixed-species dynamics}
 \label{fig:flow_SIRPF2}
@@
 \begin{align}
+\begin{split}
+&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt \\
+& \quad \quad \quad + S_n(t) d\zeta_{S_n}\\
+\end{split}\\
+% \\
+&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n}\\
 % \begin{split}
-&d S_n(t) = r_n F(t) \left(S_n(t)+\xi I_n(t)\right)dt - \theta_{S_n}S_n(t) dt - p_n f_{S_n} S_n(t) P(t)dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n},\\
+&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i}\\
 % \end{split}
 % \\
-&d I_n(t) = p_n f_{S_n} S_n(t) Pdt - \theta_{I_n} I_n(t)dt - \delta I_n(t)dt + I_n d\zeta_{I_n},\\
-% \begin{split}
-&d S_i(t) = r_i F(t)\left(S_i(t)+\xi I_i(t)\right)dt - \theta_{S_i} S_i(t) dt - p_i f_{S_i} S_i(t) P(t)dt -  \delta S_i(t)dt + S_i(t) d\zeta_{S_i},\\
-% \end{split}
-% \\
-&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t),\\
+&d I_i(t) = p_i f_{S_i} S_i(t) P(t)dt - \theta_{I_i} I_i(t)dt - \delta I_i(t)dt + I_i(t) d\zeta_{I_i}(t)\\
 \begin{split}
 &dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma_n F(t) \left(S_n(t)+\xi I_n(t)\right) dt - \gamma_i F(t) \left(S_i(t)+\xi I_i(t)\right) dt\\
-&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F,\\
+&\quad \quad \quad- \delta F(t) dt + \mu dt + F(t) d\zeta_F\\
 \end{split}
@@
 &d P(t) = \beta_n \theta_{I_n} I_n(t)dt + \beta_i \theta_{I_i} I_i(t)dt - f_{S_n} S_n(t) P(t)dt - f_{I_n} I_n(t) P(t)dt - f_{S_i} S_i(t) P(t)dt \\
-&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P},\\
+&\quad \quad \quad- f_{\mathrm{I}_i} I_i(t) P(t)dt - \theta_{\mathrm{p}} P(t)dt - \delta P(t) dt + P(t) d\zeta_{P}\\
 \end{split}
@@
 load('all_shared.RData')
+mif.estimate = as.data.frame(t(mif.estimate))
 
@@
 
-parms$K_F = parms$K_F * 1000
+parms$K_F = parms$K_F * 10
 parms$ri = parms$ri * 10
@@
 parms$probn = parms$probi * 10
-parms$f_Sn = parms$f_Sn * 10000
-parms$f_Si = parms$f_Si * 10000
-parms$f_In = parms$f_In * 1000
-parms$f_Ii = parms$f_Ii * 1000000
-parms$alpha = parms$alpha * 10000
-parms$gamman = parms$gamman * 1e8
-parms$gammai = parms$gammai * 1e4
-
-parms$sigSi = parms$sigSi * 1e4
-parms$sigIn = parms$sigIn * 1e5
-parms$sigIi = parms$sigIi * 1e5
-parms$sigF = parms$sigF * 10
-parms$sigP = parms$sigP * 100
+parms$f_Sn = parms$f_Sn * 1e5
+parms$f_Si = parms$f_Si * 1e4
+parms$f_In = parms$f_In * 1e5
+parms$f_Ii = parms$f_Ii * 1e4
+parms$alpha = parms$alpha * 1e2
+parms$gamman = parms$gamman * 1e7
+parms$gammai = parms$gammai * 1e6
+parms$sigSn = parms$sigSn * 1e2
+parms$sigSi = parms$sigSi * 1e1
+parms$sigIn = parms$sigIn * 1e6
+parms$sigIi = parms$sigIi * 1e1
+parms$sigF = parms$sigF * 1e2
+parms$sigP = parms$sigP * 1e6
+
 @
@@
 \hline\hline
-$S_j$ & Susceptible host density for species $j$  &$individuals \cdot L ^{-1}$ & Variable\\
-$I_j$ & Infected host density for species $j$ & $ individuals \cdot L ^{-1}$& Variable\\
-$F$& Alga density& $10^5 cells \cdot L ^{-1}$& Variable\\
-$P$ & Spore density & $10^3 spores \cdot L ^{-1}$& Variable\\
-$K_F$ & Alga carrying capacity&$10^2\cdot cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
-$r_n$ & Native growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
-$r_i$ & Invasive growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
-$\theta_{I_n}$ & Native infected host mortality rate&$ 10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
-$\theta_{I_i}$ & Invasive infected host mortality rate&$10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
-$\theta_{S_i}$ & Invasive susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
-$\theta_{S_n}$ & Native susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
-$\theta_{P}$ & Spore degradation rate&$10^{-1} \cdot  day^{-3}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
-$p_n$ &Native probability of infection per spore &$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
-$p_i$ & Invasive probability of infection per spore&$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
-$f_{S_n}$ & Native susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
-$f_{S_i}$ & Invasive susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
-$f_{I_n}$ & Native infected host filtering rate&$10^{-3} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
-$f_{I_i}$ & Invasive infected host filtering rate&$10^{-6} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
+$S_j$ & Susceptible host density for species $j$  &$\mathrm{individual} \cdot L ^{-1}$ & Variable\\
+$I_j$ & Infected host density for species $j$ & $ \mathrm{individual} \cdot L ^{-1}$& Variable\\
+$F$& Alga density& $10^6 \mathrm{cell} \cdot L ^{-1}$& Variable\\
+$P$ & Spore density & $10^3 \mathrm{spore} \cdot L ^{-1}$& Variable\\
+$K_F$ & Alga carrying capacity&$10^{-1}\cdot \mathrm{cell} \cdot \mathrm{L} ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
+$r_n$ & Native growth rate &$10^{-7} \cdot  \mathrm{L} \cdot \mathrm{cell}^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
+$r_i$ & Invasive growth rate &$10^{-7} \cdot  \mathrm{L} \cdot  \mathrm{cell}^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
+$\theta_{I_n}$ & Native infected host mortality rate&$ 10^{-1} \cdot \mathrm{day}^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
+$\theta_{I_i}$ & Invasive infected host mortality rate&$10^{-1} \cdot \mathrm{day}^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
+$\theta_{S_i}$ & Invasive susceptible host mortality rate&$10^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
+$\theta_{S_n}$ & Native susceptible host mortality rate&$10^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
+$\theta_{P}$ & Spore degradation rate&$10^{-3} \cdot  \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
+$p_n$ &Native probability of infection per spore &$10^{-4} \cdot \mathrm{individual} \cdot \mathrm{spore}^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
+$p_i$ & Invasive probability of infection per spore&$10^{-4} \cdot \mathrm{individual} \cdot \mathrm{spore}^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
+$f_{S_n}$ & Native susceptible host filtering rate&$10^{-5} \cdot \mathrm{L}  \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$f_Sn,2)}\\
+$f_{S_i}$ & Invasive susceptible host filtering rate&$10^{-4} \cdot \mathrm{L} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
+$f_{I_n}$ & Native infected host filtering rate&$10^{-5} \cdot \mathrm{L} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
+$f_{I_i}$ & Invasive infected host filtering rate&$10^{-4} \cdot \mathrm{L} \cdot \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\
 $\delta$ & Sampling rate&$day ^{-1} $ & 0.013\\
-$\alpha$ & Alga maximum exponential growth rate&$10^{-4} \cdot day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
-$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
-$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot individual^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
-$\mu$ & Alag refilling rate&$10^5 cells \cdot L ^{-1} \cdot day^{-1}$ & 0.37\\
-$\xi$ &Reduction in infected host reproduction and consumption& $Unitless$ & \Sexpr{mysignif(parms$xi,2)}\\
-$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betai,2)}\\
-$\beta_n$ & Spores produced per infected native species&$10^3 \cdot spores \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$betan,2)}\\
-$\sigma_{S_n}$ & Standard deviation of Brownian motion of susceptible native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
-$\sigma_{S_i}$ & Standard deviation of Brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
-$\sigma_{I_n}$ & Standard deviation of Brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
-$\sigma_{I_i}$ & Standard deviation of Brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
-$\sigma_{F}$ & Standard deviation of Brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
-$\sigma_{P}$ &Standard deviation of Brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
-$k_{S_n}$ & Overdispersion parameter& $Unitless$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
-$k_{S_i}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_Si,2)}\\
-$k_{I_n}$ &Overdispersion parameter &$Unitless$ & \Sexpr{mysignif(parms$k_In,2)}\\
-$k_{I_i}$ & Overdispersion parameter&$Unitless$ & \Sexpr{mysignif(parms$k_Ii,2)}\\[1ex]
+$\alpha$ & Alga maximum exponential growth rate&$10^{-2} \cdot \mathrm{day} ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
+$\gamma_n$ & Native host consumption rate&$10^{-7} \cdot \mathrm{L} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day} ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
+$\gamma_i$ & Invasive host consumption rate&$10^{-6} \cdot \mathrm{L} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day} ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\
+$\mu$ & Alag refilling rate&$10^6 \mathrm{cell} \cdot \mathrm{L} ^{-1} \cdot \mathrm{day}^{-1}$ & 0.37\\
+$\xi$ &Reduction in infected host reproduction and consumption& $\mathrm{Unitless}$ & \Sexpr{mysignif(parms$xi,2)}\\
+$\beta_i$ & Spores produced per infected invasive species&$10^3 \cdot \mathrm{spore} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$betai,2)}\\
+$\beta_n$ & Spores produced per infected native species&$10^3 \cdot \mathrm{spore} \cdot \mathrm{individual}^{-1} \cdot \mathrm{day}^{-1}$ & \Sexpr{mysignif(parms$betan,2)}\\
+$\sigma_{S_n}$ & Standard deviation of Brownian motion of susceptible native&$\sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigSn,2)}\\
+$\sigma_{S_i}$ & Standard deviation of Brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
+$\sigma_{I_n}$ & Standard deviation of Brownian motion of infected native&$10^{-5} \cdot \sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
+$\sigma_{I_i}$ & Standard deviation of Brownian motion of infected invasive&$10^{-5} \cdot \sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
+$\sigma_{F}$ & Standard deviation of Brownian motion of alga&$10^{-1} \cdot \sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
+$\sigma_{P}$ &Standard deviation of Brownian motion of parasite &$10^{-2} \cdot \sqrt{\mathrm{individual} \cdot \mathrm{day}^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\
+$k_{S_n}$ & Overdispersion parameter& $\mathrm{Unitless}$ & \Sexpr{mysignif(parms$k_Sn,2)}\\
+$k_{S_i}$ &Overdispersion parameter &$\mathrm{Unitless}$ & \Sexpr{mysignif(parms$k_Si,2)}\\
+$k_{I_n}$ &Overdispersion parameter &$\mathrm{Unitless}$ & \Sexpr{mysignif(parms$k_In,2)}\\
+$k_{I_i}$ & Overdispersion parameter&$\mathrm{Unitless}$ & \Sexpr{mysignif(parms$k_Ii,2)}\\[1ex]
  \hline
@@
 \begin{align}
-&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
-&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
-&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n(t) \cdot dt - \gamma_i F(t) S_i(t) \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
+&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n} \\
+&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i} \\
+\begin{split}
+&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n(t) \cdot dt - \gamma_i F(t) S_i(t) \cdot dt \\
+& \quad \quad \quad \quad -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F\\
+\end{split}
+\\
 &\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
@@
 \label{sec:res}
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Result2, include=FALSE,echo=FALSE,results=hide>>=
  load('SIRPF_model_estimate.RData')
@@
 
- save(SF_invasive_estimate=SF_invasive_estimate,SF_native_estimate=SF_native_estimate,SF2_estimate=SF2_estimate,SIRPF_invasive_estimate=SIRPF_invasive_estimate,SIRPF_native_estimate=SIRPF_native_estimate,SIRPF2_estimate=SIRPF2_estimate,file = "SIRPF_model_estimate.RData")
 @
@@
 The parameters are estimated via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models, which are presented in the Supplementary Material.
-Our analysis demonstrates that the 'All shared parameter' model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics.
-In cases like the Non-parasitized dynamics of single species, where this model's AIC is slightly higher (but within 2 points of the lowest AIC), it doesn't necessarily imply inferiority of model.
+Our analysis demonstrates that the "All shared parameter" model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics.
+In cases like the Non-parasitized dynamics of single species, where this model's AIC is slightly higher, it doesn't necessarily imply inferiority of model.
 AIC balances fit and complexity but doesn't consider stability or unseen data prediction.
@@
 This highlights the importance of looking beyond AIC for model assessment.
-As a result, in our case, the 'All shared parameter' model can be the best choice for all dynamics.\\
+As a result, in our case, the "All shared parameter" model can be the best choice for all dynamics, which has a AIC of \Sexpr{mysignif(pf.loglik.of.mif.estimate,3)} for the dynamics with both species of \textit{Daphnia} and parasite under SIRPF2 model.\\
 

### 2024-04-17 ms.Rnw
Update
Update(unfinished)

@@
 \begin{abstract}
-In ecological research, panel data, often comprising multi-magnitude time series, presents unique analytical challenges due to its high dimensionality.
-This study employs the PanelPOMP framework and the Panel Iterated Filtering (PIF) method to explore the practicality and advancement of mechanistic model development for such complex datasets.
-Mechanistic models, especially valuable in ecological research, allow for quantitative analysis and deeper understanding of dynamic systems.
-However, they are often constrained by the non-Gaussian nature of ecological data.
-Our study applies these methods to the \textit{Daphnia} panel data, consisting of independent time series with 10 data points each, to derive parameters that minimize the Akaike Information Criterion (AIC).
-The comparison of AIC values across models reveals the significant impact of alga, the prey of \textit{Daphnia}, and parasite on their population dynamics which is observed in the SIRPF model highlights this relationship.
-The panelPOMP framework's ability to handle complex panel data analysis is demonstrated, emphasizing its utility in uncovering latent processes not captured by traditional models.
-However, the study also discusses limitations and interpretations of mechanistic models and PanelPOMP framework in ecological data analysis.
+In ecological research, panel data usually consists of multiple time series, which brings challenges for analysis due to their high dimensionality.
+This study applies the PanelPOMP framework and the Panel Iterated Filtering (PIF) method to explore the utility and complexity of mechanistic models in complex system.
+Mechanistic models are particularly important in ecological research for quantitative analysis and deeper understanding of dynamic systems.
+However, they are often limited by the non-Gaussian nature of ecological data.
+Our study applies these methods to \textit{Daphnia} panel data (consisting of independent time series with 10 data points each) to derive parameters that minimize the Akaike information criterion (AIC).
+By utilizing likelihood ratio test and comparing the AIC values of the different models, a significant effect of algae (prey of \textit{Daphnia}) and parasites on their population dynamics was observed, while the SIRPF model highlighted this relationship.
+The study demonstrates the ability of the panelPOMP framework to handle complex panel data analyses, emphasizing its utility in revealing underlying processes not captured by traditional models.
+However, this study also discusses the limitations and interpretations of mechanistic models and the panelPOMP framework in ecological data analysis.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:}  Panel Iterated Filtering; SIRPF model; PanelPOMP; Daphnia.
+{\it Keywords:  Panel Iterated Filtering; SIRPF model; PanelPOMP; Daphnia}
 \vfill
@@
 
-Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary.
-As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \citep{Stalidzans2020}.
-Utilizing the established process, they elucidate the actions of hidden mechanisms \citep{Duarte2003}.
-The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems.
-This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters.
-Mechanistic modeling has been widely employed in social sciences, industrial engineering, and physics for an extended period.
-Improved computational capabilities now make it more effective.
-Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information.
-Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task.
-As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \citep{levin1992problem}.\\
+Ecological experiments contain many elements or processes that are difficult to observe, so models that can describe both observable and unobservable processes are needed to explain the entire ecological experiment.
+As mathematical descriptions of the elements that make up a system, mechanistic models can describe the interactions between the elements of the system and their interactions with the environment.
+Using established processes, they can elucidate the role of hidden mechanisms \citep{Duarte2003}.
+The purpose of mechanistic modeling in ecology is to use mathematical models to describe the development and interactions of dynamic systems.
+This quantitative tool allows researchers to assess the explanatory potential of various conceptual models and to measure the magnitude of relevant parameters.
+Mechanistic modeling has been widely used in the social sciences, industrial engineering and physics.
+Now, increased computational power has made it even more effective.
+Because of the non-smooth, non-linear and stochastic nature of ecological information, traditional time-series analysis offers limited explanatory power.
+Despite creating theoretical models for the data, linking the theory to the data remains a challenging task.
+As a result, eco-mechanical modeling is unable to address most problems on a large scale, but only a limited range of problems \citep{levin1992problem}.\\
 

### 2024-04-17 ms.Rnw
UPdate

@@
 \newcommand{\blind}{1}
-
+\usetikzlibrary{positioning}
+\usetikzlibrary {arrows.meta}
+\usetikzlibrary{shapes.geometric}
 % DON'T change margins - should be 1 inch all around.
@@
 \label{sec:intro}
-
-Ecological experiments contain many elements or processes that are difficult to observe, so models that can describe both observable and unobservable processes are needed to explain the entire ecological experiment.
-As mathematical descriptions of the elements that make up a system, mechanistic models can describe the interactions between the elements of the system and their interactions with the environment.
-Using established processes, they can elucidate the role of hidden mechanisms \citep{Duarte2003}.
-The purpose of mechanistic modeling in ecology is to use mathematical models to describe the development and interactions of dynamic systems.
-This quantitative tool allows researchers to assess the explanatory potential of various conceptual models and to measure the magnitude of relevant parameters.
-Mechanistic modeling has been widely used in the social sciences, industrial engineering and physics.
-Now, increased computational power has made it even more effective.
-Because of the non-smooth, non-linear and stochastic nature of ecological information, traditional time-series analysis offers limited explanatory power.
-Despite creating theoretical models for the data, linking the theory to the data remains a challenging task.
-As a result, eco-mechanical modeling is unable to address most problems on a large scale, but only a limited range of problems \citep{levin1992problem}.\\
+Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary.
+As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \citep{Stalidzans2020}.
+Utilizing the established process, they elucidate the actions of hidden mechanisms  The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems.
+This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters.
+Mechanistic modeling has been widely employed in social sciences, industrial engineering, and physics for an extended period.
+Improved computational capabilities now make it more effective.
+Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information.
+Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task.
+As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \\
 
@@
 Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
-
-The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the Figure 2 and model show:
-
-\usetikzlibrary{positioning}
-\usetikzlibrary {arrows.meta}
-\usetikzlibrary{shapes.geometric}
-
-\begin{figure}
-\begin{center}
-%%%%% SEAIR diagram
-\resizebox{8cm}{!}{
-\begin{tikzpicture}[
-  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
-  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1,inner sep=0cm},
-  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
-  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
-  bendy/.style={bend left=10},
-  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
-  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
-]
-\tikzset{>={}}; % this is needed to implement the arrow redefinition
-
-\node (S) at (-0.5,-0.25) [square] {$S$};
-\node (I) at (4.5,-0.25) [square] {$I$};
-\node (P) at (2.5,0.75)  [rhombus] {$P$};
-\node (R) at (3.5,2)[travel] {$R$};
-
-\node (F) at (4.5,4.5) [report] {$F$};
-
-\draw [->] (F) --  (R);
-\draw [->] (P) --  (R);
-\draw [->] (S) -- (R);
-\draw [->] (I) -- (R);
-
-\draw [->] (F) -- (S);
-\draw [->] (F) -- (I);
-
-\draw (S) edge[loop left] (S);
-\draw (F) edge[loop above] (F);
-
-\draw [->, bendy] (S) to  (I);
-\draw [->, bendy] (I) to  (S);
-\draw [->, bendy] (P) to (I);
-% \path [->] (S) edge (I)
-%            (P) edge (I);
-\draw [->, bendy] (I) to (P);
-\end{tikzpicture}
-}
-\end{center}
-\vspace{-5mm}
-\caption{A flow diagram for the SIRPF model.
-Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
-For \textit{Daphnia} either dentifera or lumholtzi, susceptible, infected \textit{Daphnia} and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$.
-Some susceptible \textit{Daphnia} get infected by parasites, transitioning from $S$ to $I$.
-Both susceptible and infected \textit{Daphnia} contribute to the $R$ state upon death, releasing surviving parasites back into the environment.
-And the parasites continue to impact other individuals and get into another cycle.
-Food is consumed by both $S$ and $I$ individuals.
-Over time, elements in $F$ and $P$ also transit to the $R$ state.}
-\label{fig:flow_SIRPF}
-\end{figure}
 
@@
 By isolating the food variable in the absence of parasites, these models can provide clearer insights into how food resource availability or scarcity impacts the population dynamics of native and invasive species.
-
-\begin{figure}[H]
-\begin{center}
-%%%%% SEAIR diagram
-\resizebox{8cm}{!}{
-\begin{tikzpicture}[
-  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
-  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
-  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
-  bendy/.style={bend left=10},
-  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
-  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
-]
-\tikzset{>={}}; % this is needed to implement the arrow redefinition
-
-\node (Sn) at (-0.5,0) [square] {$S_n$};
-\node (F) at (2.5,0) [report] {$F$};
-\node (Si) at (5.5,0) [square] {$S_i$};
-\node (R) at (2.5,2) [square] {$R$};
-
-\draw [->] (F) --  (Sn);
-\draw [->] (F) --  (Si);
-\draw [->] (F) --  (R);
-\draw [->] (Sn) --  (R);
-\draw [->] (Si) --  (R);
-\draw (Sn) edge[loop above] (Sn);
-\draw (Si) edge[loop above] (Si);
-\draw (F) edge[loop below] (F);
-
-\draw [dotted, thick, red] (3.5, -0.75) rectangle (6.5, 1.5);
-\end{tikzpicture}
-}
-\end{center}
-\vspace{-5mm}
-\caption{A flow diagram for the SF and SF2 model.
-In this system, both types of \textit{Daphnia} and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$.
-As time progresses, individuals in $F$, $S_n$, and $S_i$ transition to the $R$ state, signifying mortality.
-While the whole graph show SF2 model, if we remove the node in the dashed rectangle, the graph will represent the SF model.}
-\label{fig:SF&SF2}
-\end{figure}
-
 \begin{align}

### 2024-04-17 si/si.Rnw
UPdate

@@
 \section{Model Diagram}
-
+\subsection{SIRPF Model}
+
+The SIRPF model describes the dynamics of the Susceptible($S$), Infected($I$), Recovered, Food($F$), and Parasite($P$) populations as the Figure 2 and model show:
+
+\usetikzlibrary{positioning}
+\usetikzlibrary {arrows.meta}
+\usetikzlibrary{shapes.geometric}
+
+\begin{figure}
+\begin{center}
+\resizebox{8cm}{!}{
+\begin{tikzpicture}[
+  square/.style={rectangle, draw=black, minimum width=0.5cm, minimum height=0.5cm, rounded corners=.1cm, fill=blue!8},
+  rhombus/.style={diamond, draw=black, minimum width=0.05cm, minimum height=0.05cm, fill=purple!8,aspect = 1,inner sep=0cm},
+  travel/.style={circle, draw=black, minimum width=0.5cm, minimum height=0.5cm, fill=green!8},
+  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=0.6cm,inner sep=0cm},
+  bendy/.style={bend left=10},
+  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
+  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
+]
+\tikzset{>={}}; % this is needed to implement the arrow redefinition
+
+\node (S) at (-0.5,-0.25) [square] {$S$};
+\node (I) at (4.5,-0.25) [square] {$I$};
+\node (P) at (2.5,0.75)  [rhombus] {$P$};
+\node (R) at (3.5,2)[travel] {$R$};
+
+\node (F) at (4.5,4.5) [report] {$F$};
+
+\draw [->] (F) --  (R);
+\draw [->] (P) --  (R);
+\draw [->] (S) -- (R);
+\draw [->] (I) -- (R);
+
+\draw [->] (F) -- (S);
+\draw [->] (F) -- (I);
+
+\draw (S) edge[loop left] (S);
+\draw (F) edge[loop above] (F);
+
+\draw [->, bendy] (S) to  (I);
+\draw [->, bendy] (I) to  (S);
+\draw [->, bendy] (P) to (I);
+\draw [->, bendy] (I) to (P);
+\end{tikzpicture}
+}
+\end{center}
+\vspace{-5mm}
+\caption{A flow diagram for the SIRPF model.
+Apart from the quantified states above, the graph includes $R$ state, which represents deadth status.
+For \textit{Daphnia} either dentifera or lumholtzi, susceptible, infected \textit{Daphnia} and food retain the ability to produce offspring, which is represented in the graph as a self-recycling of $S$ and $F$, as well as transaction from $I$ to $S$.
+Some susceptible \textit{Daphnia} get infected by parasites, transitioning from $S$ to $I$.
+Both susceptible and infected \textit{Daphnia} contribute to the $R$ state upon death, releasing surviving parasites back into the environment.
+And the parasites continue to impact other individuals and get into another cycle.
+Food is consumed by both $S$ and $I$ individuals.
+Over time, elements in $F$ and $P$ also transit to the $R$ state.}
+\label{fig:flow_SIRPF}
+\end{figure}
+
+\subsection{SF2 Model}
+
+\begin{figure}[H]
+\begin{center}
+%%%%% SEAIR diagram
+\resizebox{8cm}{!}{
+\begin{tikzpicture}[
+  square/.style={rectangle, draw=black, minimum width=1cm, minimum height=1cm, rounded corners=.1cm, fill=blue!8},
+  report/.style={shape=regular polygon, regular polygon sides=8, draw, fill=red!8,minimum size=1cm,inner sep=0cm},
+  rhombus/.style={diamond, draw=black, minimum width=0.1cm, minimum height=0.1cm, fill=purple!8,aspect = 1},
+  bendy/.style={bend left=10},
+  >/.style={shorten >=0.25mm}, % redefine arrow to stop short of node
+  >/.tip={Stealth[length=1.5mm,width=1.5mm]} % redefine arrow style
+]
+\tikzset{>={}}; % this is needed to implement the arrow redefinition
+
+\node (Sn) at (-0.5,0) [square] {$S_n$};
+\node (F) at (2.5,0) [report] {$F$};
+\node (Si) at (5.5,0) [square] {$S_i$};
+\node (R) at (2.5,2) [square] {$R$};
+
+\draw [->] (F) --  (Sn);
+\draw [->] (F) --  (Si);
+\draw [->] (F) --  (R);
+\draw [->] (Sn) --  (R);
+\draw [->] (Si) --  (R);
+\draw (Sn) edge[loop above] (Sn);
+\draw (Si) edge[loop above] (Si);
+\draw (F) edge[loop below] (F);
+
+\draw [dotted, thick, red] (3.5, -0.75) rectangle (6.5, 1.5);
+\end{tikzpicture}
+}
+\end{center}
+\vspace{-5mm}
+\caption{A flow diagram for the SF and SF2 model.
+In this system, both types of \textit{Daphnia} and the food resource contribute to reproduction, depicted as self-recycling loops for $S_n$, $S_i$, and $F$.
+As time progresses, individuals in $F$, $S_n$, and $S_i$ transition to the $R$ state, signifying mortality.
+While the whole graph show SF2 model, if we remove the node in the dashed rectangle, the graph will represent the SF model.}
+\label{fig:SF&SF2}
+\end{figure}
+
+\subsection{SF Model}
 Dynamic models addressed in this manuscript are comprehensively delineated through the descriptions provided within the main text, augmented by the supplementary details furnished in this section.

### 2024-04-17 ms.Rnw
Update of gitignor

@@
 \usepackage{hyphenat}
-\usepackage{Sweave}
+
 \bibliographystyle{apalike}
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+% \SweaveOpts{concordance=TRUE}
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 
-<<Setup, include=FALSE,echo=FALSE,results=hide>>=
+<<Setup, include=FALSE,echo=FALSE,results='hide'>>=
  library(plyr)
@@
 And based on this result, the correlation between native host and invasive host in parasitized dynamics will be further explored.
+% 
+\begin{figure}[H]
+% \centering
+<<echo=FALSE, fig.width=8, fig.height=6,out.width="2in">>=
+ plot(x = 1:10)
+@
+\end{figure}
+%
 
@@
 \centering
-<<echo=FALSE, fig.cap="Plot1", fig=TRUE, width=25, height=10,out.width='\\textwidth'>>=
+<<DataVis,echo=FALSE, fig.width=9,fig.height=6,out.width="2in">>=
 mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
@@
   facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
-  theme(axis.title.x = element_blank(),
+  theme(axis.title.x = element_blank( ),
         axis.title.y = element_text(size = 10),  # Y-axis title
@@
         strip.background = element_blank(),
-        strip.placement = "outside") +
+        strip.placement = "outside",
+        plot.margin = c(.5,.5,.5,.5)) +
   ylab('Square root of Daphnia density') +
@@
 grid.arrange(p1, p2, ncol = 1)
+
 @
 % \includegraphics[width=1\textwidth]{"Density_plot.png"}
-% \label{fig:plot1}
+\label{fig:plot1}
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}.
@@
 
-<<Results, include=FALSE,echo=FALSE,results=hide>>=
+<<Results, include=FALSE,echo=FALSE,results='hide'>>=
 load('all_shared.RData')
@@
 \label{sec:res}
-<<Result2, include=FALSE,echo=FALSE,results=hide>>=
+<<Result2, include=FALSE,echo=FALSE,results='hide'>>=
  load('SIRPF_model_estimate.RData')

### 2024-04-24 ms.Rnw
Update

@@
  library(patchwork)
+ library(dplyr)
+ library(gtable)
 
@@
 
-                      if (Sn < 0.0) {
+                      if (Sn < 0.0 || Sn > 1e5) {
                         Sn = 0.0;
@@
                       }
-                      if (Si < 0.0) {
+                      if (Si < 0.0 || Si > 1e5) {
                         Si = 0.0;
@@
                       }
-                      if (F < 0) {
-                        F = 0;
+                      if (F < 0.0 || F > 1e20) {
+                        F = 0.0;
                         error_count += 1000;
                       }
-                      if (In < 0) {
-                        In = 0;
+                      if (In < 0.0 || In > 1e5) {
+                        In = 0.0;
                         error_count += 0.001;
                       }
-                      if (Ii < 0) {
-                        Ii = 0;
+                      if (Ii < 0.0 || Ii > 1e5) {
+                        Ii = 0.0;
                         error_count += 0.000000001;
                       }                      
-                      if (P < 0 && t > 3.9) {
-                        P = 0;
+                      if ((P < 0.0 || P > 1e20)&& t > 3.9) {
+                        P = 0.0;
                         error_count += 0.000001;
@@
 
-                      T_Sn += delta * Sn * dt;
-                      T_In += delta * In * dt;
-                      T_Si += delta * Si * dt;
-                      T_Ii += delta * Ii * dt;
+                      T_Sn = delta * Sn * dt;
+                      T_In = delta * In * dt;
+                      T_Si = delta * Si * dt;
+                      T_Ii = delta * Ii * dt;
+
+                      T_Sn = fabs(T_Sn);
+                      T_In = fabs(T_In);
+                      T_Si = fabs(T_Si);
+                      T_Ii = fabs(T_Ii);
 
@@
                      F = 16.667;
-                     T_Sn = 0;
-                     T_Si = 0;
-                     T_In = 0;
-                     T_Ii = 0;
-                     In = 0;
-                     Ii = 0;
+
+                     T_Sn = 0.0;
+                     T_Si = 0.0;
+                     T_In = 0.0;
+                     T_Ii = 0.0;
+                     In = 0.0;
+                     Ii = 0.0;
 
@@
                  if (error_count > 0.0) {
-                   lik = -100;
+                   lik = -150;
                   }
@@
 Ecological experiments contain many elements or processes that are difficult to observe, models that can describe both observable and unobservable processes to explain the whole ecological experiment are necessary.
-As a mathematical description of the elements forming a system, mechanistic modeling can describe the mutual interactions between elements in the system and their interactions with the environment \citep{Stalidzans2020}.
-Utilizing the established process, they elucidate the actions of hidden mechanisms  The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems.
+As a mathematical description of the elements forming a system, mechanistic modeling can describe the interactions between elements in the dynamic system and their interactions with the environment \citep{Stalidzans2020}.
+Utilizing the established process by mechanistic models, we can understand the hidden mechanisms in dynamic systems better.
 This quantitative tool allows researchers to evaluate the explanatory potential of various conceptual models and measure the magnitude of pertinent parameters.
@@
 Improved computational capabilities now make it more effective.
+The purpose of mechanistic modeling in ecology involves employing mathematical models to depict the development and interaction of dynamic systems.
 Traditional time series analysis offers limited explanatory power due to the nonstationary, nonlinear, and stochastic nature of ecological information.
 Despite the creation of theoretical models for the data, connecting theories with data continues to be a challenging task.
-As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems \\
+As a result, ecological mechanical modeling cannot solve most of the problems on a wide scale, but rather, only a limited range of the problems.
+\\
 
@@
 In ecology, panel data is a traditional data form that researchers gather to mitigate the impact of random variations throughout experiments, so that it can more accurately measure changes in experimental objectives and control the variables needed to achieve experimental goals.
+The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \citep{King2016}.
+This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering.
+Stemming from iterated filtering \citep{ionides2011iterated} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood.
+This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.\\
+
 Due to the non-linear characteristics of ecological data, the analysis of panel data entails a high-dimensional structure.
@@
 The goal of this study is to investigate the practicality of employing mechanistic models on ecological panel data using the panelPOMP structure.\\
-
-The PanelPOMP approach modifies the partially observable Markov Process model to accommodate panel data by constructing a POMP model for each individual component \citep{King2016}.
-This method enables the determination of whether the parameters within the selected mathematical model are shared or distinct throughout the panel, by computing the highest log likelihood for all potential combinations through the use of panel iterated filtering.
-Stemming from iterated filtering \citep{King2016} applied to a singular nonlinear stochastic time series, panel iterated filtering not only filters within each series but also cycles across the panel to achieve the best likelihood.
-This enables the application of mechanistic models to panel data while preserving its intricate dimensionality.\\
 
@@
 For a more in-depth examination of the concealed processes in the panel data, this research incorporates the Susceptible-Infected-Recovered-Food-Parasite (SIRPF) model - within the panelPOMP framework.
-After applying the model to the \textit{Daphnia} panel data, the maximum log-likelihood and best parameter estimates are determined via an iterative filtering approach.
+After applying the model to the \textit{Daphnia} panel data, the best parameter estimates are determined through maximum log-likelihood estimation via the iterative filtering approach.
 These findings are subsequently compared to discover a more accurate and appropriate model for the given data.
-The superior performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is \textit{Daphnia}'s prey, impact the dynamics of both species of \textit{Daphnia}.
+The good performance of the SIRPF model implies that the population density of alga \textit{Ankistrodesmus falcatus}, which is \textit{Daphnia}'s prey, plays an important role in the dynamics of both species of \textit{Daphnia}.
 It is worth noting that the population density of alga is neither observable nor recorded in the experiment.
 That implies the mechanistic models under panelPOMP potentially better explain the latent process and show the importance of unobservable processes \citep{Edwards2012}.
-At the same time, the panelPOMP framework provides an estimation of the parameters in the model, which helps to understand the hidden mechanisms in the ecological system \citep{Carles2020}.
-Based on the estimation, a detailed interpretation of relationship between invasive host and native host with presence of parasite is explaned.\\
-
-Our objective is to evaluate the explanatory capabilities of two traditional ecological models when applied to the \textit{Daphnia} panel data, and to estimate the parameter sets for a more accurate depiction of the hidden mechanisms within the ecological system.
-By analyzing \textit{Daphnia} panel data, we hope to show the limitations and interpretations of mechanistic models in dealing with ecological data.
-Additionally, the later section will further discuss how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models.
+Nevertheless, the panelPOMP framework provides an estimation of the parameters for each dynamic systems included, which helps to understand the hidden mechanisms in the ecological system \citep{Carles2020}.
+Based on the estimation, a detailed interpretation of relationship between each component in the ecological system is explained.\\
+
+Our objective is to evaluate the explanatory capabilities of SIRPF model when applied to the \textit{Daphnia} panel data, and to estimate the parameter sets for a more accurate interpretation of the hidden mechanisms within the ecological system comparing to classical model.
+By analyzing \textit{Daphnia} panel data, we also show the interpretations and limitations of mechanistic models when dealing with ecological data.
+Additionally, how the panelPOMP framework and PIF contribute to overcoming the inherent challenges associated with fitting ecological models will be discussed in detail.
 
@@
 The dynamics of our study are derived from the Searle's comparison experiment, which focuses on two \textit{Daphnia} species: \textit{Daphnia} \textit{dentifera Forbes}, natively prevalent in North American stratified lakes \citep{searle2016salinization}, and \textit{Daphnia} \textit{lumholtzi Sars}, an invasive species originating from Africa, Asia, and Australia.
-D.
-lumholtzi, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \citep{havel1993daphnia}.
+\textit{D.
+lumholtzi}, first observed in North America in the 1990s, has rapidly expanded across the United States, impacting native species and ecosystem dynamics \citep{havel1993daphnia}.
 These \textit{Daphnia} species, referred to as the native and invasive hosts, are bred in controlled laboratory environments using isofemale lines.
@@
 The dynamics also involves \textit{Metschnikowia bicuspidata}, a parasitic fungus predominantly infecting the native host.
-This parasite, ingested as spores by \textit{Daphnia}, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
+This species of parasite is strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \citep{Searle2016}.
+Searle's study utilized the \textit{M.
+bicuspidata}, which is ingested as spores by \textit{Daphnia}, aggressively affects the host's fecundity and longevity \citep{duffy2010temporal}.
 The parasite's presence is identifiable by the opaqueness it causes in normally transparent hosts, with spore release upon host death.
-Searle's study utilized a M.
-bicuspidata strain from Baker Lake, Michigan, with consistent infectivity and virulence characteristics in different samples \citep{Searle2016}.
-The final species in our study is \textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for \textit{Daphnia}.
+\textit{Ankistrodesmus falcatus}, a green alga that serves as the primary food source for \textit{Daphnia} also presents in the experiment.
 \textit{Ankistrodesmus falcatus} is a globally distributed freshwater chlorophyte, is known for its role in aquatic communities and has been utilized in ecological research due to its nutritional properties \citep{arnold1971ingestion}.
@@
 
-Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two \textit{Daphnia} host species exhibited comparable resource acquisition rates, the invasive species demonstrated a more rapid reproductive cycle \citep{Searle2016}.
+Drawing upon the empirical findings from Searle's laboratory, it was observed that while the two \textit{Daphnia} host species exhibited comparable resource acquisition rates, the invasive species is a strong interspecific competitor \citep{Searle2016}.
 This observation suggests a potential competitive advantage for the invasive species.
@@
 Their approach involved a mesocosm experiment that varied species composition and parasite presence.
-The experimental design was a 3x2x1 factorial arrangement, incorporating three species combinations (native only, invasive only, or both existing species), two parasite exposure scenarios (exposed or unexposed), and a single algal food source.
+The experimental design was a 3 $\times$ 2 $\times$ 1 factorial arrangement, incorporating three species combinations (native only, invasive only, or both existing species), two parasite exposure scenarios (exposed or unexposed), and a single algal food source.
 This design enabled an in-depth analysis of the invasive host's impact on native populations, manifesting either directly through competition for resources, indirectly via the amplification of a mutual parasite, or through a reduction in parasite effects.\\
@@
 This effect is quantifiable through the ratio of natality to mortality in the native host population.
-Conversely, should the invasive species primarily influence the native host indirectly via interactions with parasites, a different pattern emerges.
-Here, the survivability of the native host is expected to be lower in scenarios involving both the parasite and the invasive host, as opposed to situations where the native host is only affected by the parasite.
+Apart from food, disease affect species' survivability.
+The invasive host dilutes disease when it is a strong interspecific competitor, while a weak interspecific invasive species amplifies disease.
 This effect is quantifiable through the ratio of natality to mortality in the native host population as well.\\
@@
 
-In the Non-parasitized level, indoor mesocosms were used \citep{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media.
+At the non-parasitized level, indoor mesocosms were used \citep{Searle2016}, each comprising a 18.9 L bucket with 15 L of high-hardness COMBO media.
 The experimental design involved 10 replicates for single-species treatments, and 9 replicates for mixed-species treatment.
@@
 As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both non-parasitized and prarasitized dynamics to minimize information bias.
-Figures below illustrates these dynamics, with the first row of Figure 1 depicting non-parasitized dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
+Figure 1 illustrates these dynamics, with the first row of Figure 1 depicting non-parasitized dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}.
 The second row of Figure 1 shows parasitized dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation.
@@
 And based on this result, the correlation between native host and invasive host in parasitized dynamics will be further explored.
-% 
-\begin{figure}[H]
-% \centering
-<<echo=FALSE, fig.width=8, fig.height=6,out.width="2in">>=
- plot(x = 1:10)
-@
-\end{figure}
-%
 
@@
 \centering
-<<DataVis,echo=FALSE, fig.width=9,fig.height=6,out.width="2in">>=
+<<DataVis,echo=FALSE, fig.width=9,fig.height=6,out.width="7in">>=
 mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
@@
 p1 <- ggplot(mixed_nopara_data_combined, aes(x = day, y = sqrt(dent.adult))) +
-  geom_line(aes(color = "Dent Adult"), size = 1) +
-  geom_line(aes(y = sqrt(lum.adult), color = "Lum Adult"), size = 1) +
+  geom_line(aes(color = "Dent Susceptible"), linewidth = 1) +
+  geom_line(aes(y = sqrt(lum.adult), color = "Lum Susceptible"), linewidth = 1) +
   facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
@@
         plot.margin = c(.5,.5,.5,.5)) +
-  ylab('Square root of Daphnia density') +
+  ylab('Square root of Daphnia density per liter') +
   scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
   labs(title = "Daphnia Density in Non-parasitized Dynamics", color = "Species") +
-  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue")) +
+  scale_color_manual(values = c("Dent Susceptible" = "black", "Lum Susceptible" = "blue")) +
   theme_minimal() +
@@
 p2 <- ggplot(mixed_para_data_combined, aes(x = day, y = sqrt(dentadult))) +
-  geom_line(aes(color = "Dent Adult"), size = 1) +
-  geom_line(aes(y = sqrt(lumadult), color = "Lum Adult"), size = 1) +
-  geom_line(aes(y = sqrt(luminf), color = "Lum Inf"), size = 1) +
-  geom_line(aes(y = sqrt(dentinf), color = "Dent Inf"), size = 1) +
+  geom_line(aes(color = "Dent Susceptible"), linewidth = 1) +
+  geom_line(aes(y = sqrt(lumadult), color = "Lum Susceptible"), linewidth = 1) +
+  geom_line(aes(y = sqrt(luminf), color = "Lum Inf"), linewidth = 1) +
+  geom_line(aes(y = sqrt(dentinf), color = "Dent Inf"), linewidth = 1) +
   facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
@@
         plot.margin = c(.5,.5,.5,.5)) +
-  ylab('Square root of Daphnia density') +
+  ylab('Square root of Daphnia density per liter') +
   scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
   labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
-  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red")) +
+  scale_color_manual(values = c("Dent Susceptible" = "black", "Lum Susceptible" = "blue", "Lum Inf" = "green", "Dent Inf" = "red")) +
   theme_minimal() +
@@
 % \includegraphics[width=1\textwidth]{"Density_plot.png"}
-\label{fig:plot1}
+\label{fig:data_vis}
 \caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}.
@@
 
-Methodologically, POMP models are analyzed using a suite of computational techniques.
+Methodologically, POMP models are analyzed using combination of various computational techniques.
 Sequential Monte Carlo methods, such as particle filters, are commonly employed for state estimation and parameter inference in these models \citep{bhadra2011malaria}.
@@
 
-Working with POMP models can prove difficult due to the complexity of formulating both the likelihood function and transition density in intricate settings.
+Working with POMP models can prove difficulty due to the complexity of formulating both the likelihood function and transition density in intricate settings.
 To address this issue, several techniques have been devised, including the particle filter.
@@
 
-To implement the particle filter, we start by initializing $M$ particles at time 1, where $\{ X_1^m; m \in 1:M \}$.
+To implement the particle filter, we start by initializing $M$ particles at time 1, where $\{ X_1^m; m \in (1,2,\cdots,M) \}$.
 All with the required latent state detail and a vector of parameter values, $\theta$.
-As a result, for every time step n, where $n \in $ [$N$], we advance the each particle by doing the one-step random walk based on transition density, resulting in a set of particles that represents the prediction distribution at time n, which can be represented as $f_{X_n|X_{n-1}}(\cdot | X_{n-1}^m; \theta)$.
+As a result, for every time step n, where $n \in $ [$1:N$], we advance the each particle by doing the one-step random walk based on transition density, resulting in a set of particles that represents the prediction distribution at time n, which can be represented as $f_{X_n|X_{n-1}}(\cdot | X_{n-1}^m; \theta)$.
 Next, we assign weights to the particles based on the measurement density to obtain an ensemble that represents the filtering distribution at time n, where the new weight is calculated via $w_{n,m} = f_{Y_n|X_n}(y_n | x_n^m)$.
-Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \citep{5}.
+Finally, we reassign the particles based on their weights to generate a new set of particles that represents the filtering distribution, which can then be used to estimate model parameters swarm based on the maximum likelihood \citep{Carles2020}.
 The particle filter is particularly useful in situations where both the likelihood function and transition density are difficult to calculate analytically.\\
@@
 The SIRPF model incorporates these factors as quantitative latent state variables, enabling a more comprehensive understanding of the system's dynamics.
-In the meantime, the ecological processes of \textit{Daphnia} can be viewed as a Markov process, where the population of Susceptible and Infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
+In the meantime, the ecological processes of \textit{Daphnia} can be viewed as a Markov process, where the population of susceptible and infected individuals is observable, while the amount of food and parasites represent unobservable latent states.
 Utilizing panel data on \textit{Daphnia}, the PanelPOMP framework becomes a valuable analytic tool \citep{Carles2020}.
-Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.\\
-
+Despite the lack of direct measurement data for these latent states, PanelPOMP, which has proven efficiency in analyzing ecological panel data, facilitates a robust analysis of these dynamics, effectively handling the complexities inherent in ecological systems.
 {
@@
 &d S(t) = r F(t)\left(S(t)+\xi I(t)\right)dt - \theta_{S}S(t) dt - p f_{S} S(t) Pdt - \delta S(t)dt + S(t) d\zeta_{S}\\
-&d I(t) = p f_{S} S(t) Pdt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}\\
+&d I(t) = p f_{S} S(t) P(t)dt - \theta_{I} I(t)dt - \delta I(t)dt + I(t) d\zeta_{I}\\
 &dF(t) = \alpha F(t)\left(1-\frac{F(t)}{K_F}\right)dt - \gamma F(t) \left(S(t)+\xi I(t)\right) dt - \delta F(t) dt + \mu dt + F(t) d\zeta_F\\
@@
 Building on the SIRPF model, the SIRPF2 model further adapts to a mixed-race parasitized system by introducing another \textit{Daphnia} species.
-This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$)hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows.
-In the SIRPF2 model, the subscript n is used to denote parameters and variables related to the native host, i for those pertaining to the invasive host, and p for the parasite spores.
+This advanced model incorporates both invasive($S_i , I_i$) and native($S_n , I_n$) hosts, along with food($F$) and parasites($P$), and delineates their specific interactions within the ecosystem as the following graph and model shows.
+In the SIRPF2 model, the subscript $n$ is used to denote parameters and variables related to the native host, $i$ for those pertaining to the invasive host, and $p$ for the parasite spores.
 The SIRPF2 model thus offers a more detailed representation of the ecological dynamics by accounting for the complex interplay between different species and environmental factors in a mechanistic framework.
@@
 \caption{A flow diagram for the SIRPF2 model.
-Apart from the quantified states above, the graph includes $R$ state, which represents death status.
+The graph includes $R$ state, which represents death status.
 The flow diagram shows the how two species of \textit{Daphnia} interact with each other indirectly through $F$ and $P$.
@@
 
-In the SIRPF2 model, detailed in Table 1, parameters and variables are distinctly annotated for native hosts (subscript $n$), invasive hosts (subscript $i$), and parasite spores (subscript $p$).
-Intraspecific competition among hosts is conceptualized via the Lotka-Volterra framework, with $r_n$ and $r_i$ denoting respective reproductive rates.
+In the SIRPF2 model, detailed in Table 1, parameters and variables are distinctly annotated for native, invasive hosts, food and parasite spores.
+Intraspecific competition among hosts is conceptualized via limitation of food and environment resources, with $r_n$ and $r_i$ denoting respective reproductive rates.
 Host density growth, influenced by algal density, is posited to be linear, indicating a direct relationship between food availability and \textit{Daphnia} reproduction.
 Mortality rates of susceptible hosts are denoted as $\theta_{S_n}$ and $\theta_{S_i}$.
-The model encapsulates the dynamics of infection through \textit{Metschnikowia bicuspidata} intake, affecting susceptible hosts.
+The model represents the dynamics of infection through \textit{Metschnikowia bicuspidata} intake, affecting susceptible hosts.
 This interaction consequently modulates reproductive rates, represented as $\xi r_n$.
@@
 Upon death, infected hosts ($\theta_{I_n}$, $\theta_{I_i}$) release spores, which are subsequently subject to ingestion and environmental degradation.
+After death of host, the parasite spores are released with rate of $\beta_i$ and $\beta_n$.
 The consumption dynamics of food by susceptible and infected hosts are captured by $\gamma_n$ and $\gamma_i$, and food growth is inversely proportional to its carrying capacity ($K_F$), reflective of resource limitation.
@@
 The integration of these stochastic terms within the SIRPF2 framework, as detailed in Table 1, enhances the model's fidelity in simulating ecological interactions.
-The model, grounded in the Lotka-Volterra competition framework and augmented with stochastic components, provides a nuanced understanding of the interplay between native and invasive hosts, and their interaction with the surrounding resources and parasites.
-This incorporation of stochasticity, particularly through the lens of Brownian motion, is pivotal in transcending the limitations of deterministic models, thereby offering a more robust and comprehensive tool for ecological analysis.\\
+The model, grounded in the Lotka-Volterra competition framework and augmented with stochastic components, provides a detailed understanding of the interplay between native and invasive hosts, and their interaction with the surrounding resources and parasites.
+This incorporation of Brownian motion, is pivotal in transcending the limitations of deterministic models, thereby offering a more robust and comprehensive tool for ecological analysis.\\
 
@@
 This distribution effectively models count data where the variance surpasses the mean, typical in ecological data due to the correlated nature of biological events \citep{ning2021systemic}.
-In our model, the parameter $F$ represents the number of successful \textit{Daphnia} reproductions before observing a fixed number of food scarcity events, and $S / \delta$ reflects average number of \textit{Daphnia} per liter per sampling time in each trial.
+In our model, the parameter $F$ represents the number of successful \textit{Daphnia} reproductions before observing a fixed number of food, and $S / \delta$ reflects average number of \textit{Daphnia} per liter per sampling time in each trial.
 This framework aligns with the observed dynamics in \textit{Daphnia} populations, where variability in reproduction rates, influenced by fluctuating food availability, leads to overdispersed population counts.
@@
 
-The model's effectiveness in capturing the \textit{Daphnia} population dynamics hinges on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
+The model's effectiveness in capturing the \textit{Daphnia} population dynamics depends on its ability to account for overdispersion, a critical factor often overlooked in simpler models like the Poisson distribution \citep{breto2011compound}.
 As such, the negative binomial model stands as a more accurate representation of the complex ecological processes governing \textit{Daphnia} populations, particularly under varying food conditions.
@@
 When constructing statistical models to analyze ecological data, the specification of parameters plays a crucial role in shaping the model's complexity and interpretability.
-In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has significant implications for the model's dimensionality.
-For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimensionality by the number of units, which in this case results in an increase by 10 for single species Non-parasitized dynamics, 9 for mixed species Non-parasitized dynamics and Parasitized dynamics with the invasive species of \textit{Daphnia}, and 8 for the rest dynamics.\\
+In a scenario with 10 units of observation, the decision to assign a parameter as shared across units or unique to each one has strong implications for the model's dimension.
+For instance, designating a parameter as shared across all units incrementally increases the model dimension by one, whereas categorizing a parameter as unique to each unit multiplies the dimension by the number of units, which in this case results in an increase by 10 for single species non-parasitized dynamics, 9 for mixed species non-parasitized dynamics and parasitized dynamics with the invasive species of \textit{Daphnia}, and 8 for the rest dynamics.\\
 
 Permitting parameters to be either shared or unique leads to varying dimensions for each model.
-With respective amount of units $d$ in the panel, adding a shared parameter increases the model's dimension by 1, and a specific parameter will add dimensions by $d$.
 We encountered a balance between fitting accuracy and simplicity when selecting models.
-Akaike information criterion (AIC) is a tool for model selection that balances model complexity against goodness of fit \citep{bozdogan1987model}.
-It is rooted in information theory, particularly the concept of entropy, and provides a relative measure of the information lost when a given model is used to represent the process that generated the data.
+Likelihood ratio test and Akaike Information Criterion (AIC) are tools for model selection that balances model complexity against goodness of fit \citep{bozdogan1987model} \citep{vuong1989likelihood}.
+While we start with permitting all parameters are unit-specific to fit the model, more parameters are unified as shared among all units during the fitting process.
+As a result, a shared parameter model will be the nested model of the model which permits some parameters to be unit-specific.
+Hypothesis test is a valid option to determine if unit-specifc parameters are statisticsally significant.
+We apply likelihood ratio test for all hypothesis and validate the essences of shared parameters, where the detail is shown in the supplementary material.
+Also, it is rooted in information theory, particularly the concept of entropy, and provides a relative measure of the information lost when a given model is used to represent the process that generated the data.
 A lower AIC value indicates a model with a better balance between fit and simplicity.
@@
 
-The parameters are estimated via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models, which are presented in the Supplementary Material.
-Our analysis demonstrates that the "All shared parameter" model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics.
-In cases like the Non-parasitized dynamics of single species, where this model's AIC is slightly higher, it doesn't necessarily imply inferiority of model.
+The parameters are estimated via searching the maximum log-likelihood using panel iterated filtering and calculated AIC for all models, which are presented in the supplementary material.
+Our analysis demonstrates that the "all shared parameter" model, using consistent parameters across each unit, aligns with the principle of experimental parallelism and yields the best (near-optimal) AIC for all six dynamics.
+In cases like the non-parasitized dynamics of single species, where this model's AIC is slightly higher, it doesn't necessarily imply inferiority of model.
 AIC balances fit and complexity but doesn't consider stability or unseen data prediction.
-Here, Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \citep{nylund2007deciding}.
+While Monte Carlo simulations can show that a model with marginally higher AIC might generalize better or offer more robust predictions, possibly due to the lower AIC model overfitting the data \citep{nylund2007deciding}.
 This highlights the importance of looking beyond AIC for model assessment.
-As a result, in our case, the "All shared parameter" model can be the best choice for all dynamics, which has a AIC of \Sexpr{mysignif(pf.loglik.of.mif.estimate,3)} for the dynamics with both species of \textit{Daphnia} and parasite under SIRPF2 model.\\
-
-By leveraging the estimated parameters from the most suitable model, we can conduct a nuanced numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species.
+As a result, in our case, the "All shared parameter" model can be the best choice for all dynamics, which has a likelihood of \Sexpr{mysignif(pf.loglik.of.mif.estimate,3)}(with standard error of \Sexpr{mysignif(s.e.of.pf.loglik.of.mif.estimate,3)}) and AIC of \Sexpr{mysignif(AIC,3)} for the dynamics with both species of \textit{Daphnia} and parasite under SIRPF2 model.
+The SIRPF(2) model's performance with shared parameters points to a commonality in the factors affecting both native and invasive species, a result that could have implications for the management of these populations and their associated ecological interactions.\\
+
+By leveraging the estimated parameters from the most suitable model, we can conduct a numerical analysis of the impact of food resources and parasitism on the interactions between local and invasive host species.
 The survivability ratio($ \lambda = \frac{r}{\theta}$) of growth to mortality rates is a useful metric for assessing the health of a population, as it provides a measure of the population's ability to sustain itself in the face of environmental pressures \citep{ma2021unified}.
 This approach, grounded in ecological principles, enables a detailed examination of the survival strategies of the hosts in response to external challenge.
-The survivability ratio of native and invasive \textit{Daphnia} in single species Non-parasitized Dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectivly.
-While if we mix two species of \textit{Daphnia} into a Non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native \textit{Daphnia} and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species.
+The survivability ratio of native and invasive \textit{Daphnia} in single species non-parasitized dynamics are \Sexpr{mysignif(ratio_SF_n,3)} and \Sexpr{mysignif(ratio_SF_i,3)} respectively.
+While if we mix two species of \textit{Daphnia} into a non-parasitized dynamics, the ratio drops down to \Sexpr{mysignif(ratio_SF2_n,3)} for native \textit{Daphnia} and to \Sexpr{mysignif(ratio_SF2_i,3)} for invasive species.
 In such scenario, we observe significant fluctuations in the population density of \textit{Daphnia}.
@@
 This suggests that while environmental factors like food scarcity impact the survivability of both species, they do so in a proportionally similar manner, maintaining the relative size of survivability between them.
-And similar results can be found in the Parasitized dynamics of invasive species of \textit{Daphnia}, where the survivability ratio of susceptible native and invasive \textit{Daphnia} are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.\\
-
-These findings underscore the importance of parameter sharing in ecological modeling when multiple species or units are involved.
-Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between \textit{Daphnia} populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones.
-The results further reinforce the utility of the AIC in guiding the selection of models that not only provide a good fit to the data but also maintain a degree of simplicity, avoiding overfitting.
-The SIRPF model's performance with shared parameters points to a commonality in the factors affecting both native and invasive species, a result that could have implications for the management of these populations and their associated ecological interactions.\\
+And it can be found that in the parasitized dynamics of invasive species of \textit{Daphnia}, where the survivability ratio of susceptible native and invasive \textit{Daphnia} are \Sexpr{mysignif(ratio_SIRPF_ns,3)} and \Sexpr{mysignif(ratio_SIRPF_is,3)} for single species dynamics, and \Sexpr{mysignif(ratio_SIRPF2_ns,3)}, \Sexpr{mysignif(ratio_SIRPF2_is,3)} for native and invasive population in mixed species dynamics respectively.
+These suggest large reductions in native population survivability, which will cause the invasive host to become an amplifier of disease for the native host.
+As a result, the invasive host is viewed as a stronger competitor comparing to the native hosts.
+\\
+
+\subsection{Simulations}
+\begin{figure}[H]
+\centering
+<<SimulationPlot,echo=FALSE, fig.width=9,fig.height=6,out.width="7in">>=
+load('all_shared.RData')
+coef(mix_par_panelfood) <- mif.estimate
+all_params <- mif.estimate
+
+foreach(u = names(mix_par_panelfood), .combine = rbind) %do% {
+  unit_model <- unitobjects(mix_par_panelfood)[[u]]
+  shared <- all_params
+  pomp::simulate(
+    unit_model, 
+    nsim = 10000,
+    format = "data.frame",
+    params = c(shared)
+  ) -> sims
+
+  sims$unit <- u
+  sims
+} -> all_sims
+
+mixed_para_data_df <- mixed_para_data[[1]]
+mixed_para_data_df$unit <- "u1"
+for (i in 2:length(mixed_para_data)) {
+  mixed_para_data_unit <- mixed_para_data[[i]]
+  mixed_para_data_unit$unit <- paste0('u', i)
+
+  mixed_para_data_df <- rbind(mixed_para_data_df, mixed_para_data_unit)
+}
+
+# all_sims[all_sims$error_count == 0,] -> large_removed
+all_sims[!all_sims$.id %in% unique(all_sims[all_sims$error_count != 0,]$.id),] -> large_removed
+
+quantile_data_dentadult <- large_removed %>%
+  group_by(day) %>%
+  summarize(
+    lower = quantile(dentadult, probs = 0.025),
+    upper = quantile(dentadult, probs = 0.975),
+    .groups = 'drop'
+  )
+
+quantile_data_lumadult <- large_removed %>%
+  group_by(day) %>%
+  summarize(
+    lower = quantile(lumadult, probs = 0.01),
+    upper = quantile(lumadult, probs = 0.985),
+    .groups = 'drop'
+  )
+
+quantile_data_luminf <- large_removed %>%
+  group_by(day) %>%
+  summarize(
+    lower = quantile(luminf, probs = 0.05),
+    upper = quantile(luminf, probs = 0.995),
+    .groups = 'drop'
+  )
+
+quantile_data_dentinf <- large_removed %>%
+  group_by(day) %>%
+  summarize(
+    lower = quantile(dentinf, probs = 0.08),
+    upper = quantile(dentinf, probs = 0.998),
+    .groups = 'drop'
+  )
+
+y_range_row1 <- range(sqrt(max(quantile_data_dentadult$upper,quantile_data_lumadult$upper)),
+                      sqrt(min(quantile_data_dentadult$lower,quantile_data_lumadult$lower)), na.rm = TRUE)
+y_range_row2 <- range(sqrt(max(quantile_data_dentinf$upper,quantile_data_luminf$upper)),
+                      sqrt(min(quantile_data_dentinf$lower,quantile_data_luminf$lower)), na.rm = TRUE)
+
+color_scale <- scale_color_brewer(palette = "Dark2", guide = FALSE)
+# First row plots
+p3 <- suppressWarnings(
+  ggplot() +
+    geom_ribbon(data = quantile_data_dentadult, aes(x = day, ymin = sqrt(lower), ymax = sqrt(upper)), fill = "#ADD8E6", alpha = 0.5) +
+    geom_line(data = mixed_para_data_df, aes(x = day, y =  sqrt(dentadult), group = unit, color = unit)) +
+    scale_y_continuous(limits = y_range_row1) +
+    theme_minimal() +
+    color_scale +
+    labs(x = NULL, y = expression(paste("Sqrt of susceptible ", italic("D. dentifera"), " density"))) +
+    theme(axis.title = element_text(size = 10), axis.text.x = element_blank())
+)
+
+p4 <- suppressWarnings(
+  ggplot() +
+    geom_ribbon(data = quantile_data_lumadult, aes(x = day, ymin = sqrt(lower), ymax = sqrt(upper)), fill = "#ADD8E6", alpha = 0.5) +
+    geom_line(data = mixed_para_data_df, aes(x = day, y =  sqrt(lumadult), group = unit, color = unit)) +
+    scale_y_continuous(limits = y_range_row1) +
+    theme_minimal() +
+    color_scale +
+    labs(x = NULL, y = expression(paste("Sqrt of susceptible ", italic("D. lumholtzi"), " density"))) +
+    theme(axis.title = element_text(size = 10), axis.text.x = element_blank())
+)
+
+p5 <- suppressWarnings(
+  ggplot() +
+    geom_ribbon(data = quantile_data_dentinf, aes(x = day, ymin = sqrt(lower), ymax = sqrt(upper)), fill = "#ADD8E6", alpha = 0.5) +
+    geom_line(data = mixed_para_data_df, aes(x = day, y =  sqrt(dentinf), group = unit, color = unit)) +
+    scale_y_continuous(limits = y_range_row2) +
+    theme_minimal() +
+    color_scale +
+    labs(x = "Day", y = expression(paste("Sqrt of infected ", italic("D. dentifera"), " density"))) +
+    theme(axis.title = element_text(size = 10))
+)
+
+p6 <- suppressWarnings(
+  ggplot() +
+    geom_ribbon(data = quantile_data_luminf, aes(x = day, ymin = sqrt(lower), ymax = sqrt(upper)), fill = "#ADD8E6", alpha = 0.5) +
+    geom_line(data = mixed_para_data_df, aes(x = day, y =  sqrt(luminf), group = unit, color = unit)) +
+    scale_y_continuous(limits = y_range_row2) +
+    theme_minimal() +
+    color_scale +
+    labs(x = "Day", y = expression(paste("Sqrt of infected ", italic("D. lumholtzi"), " density"))) +
+    theme(axis.title = element_text(size = 10))
+)
+
+suppressWarnings(
+  {
+    grid_arrange <- grid.arrange(p3, p4, p5, p6, ncol = 2, nrow = 2, newpage = TRUE)
+  }
+)
+# legend <- gtable_filter(ggplot_gtable(ggplot_build(p5 + scale_color_brewer(palette = "Dark2"))), "guide-box")
+# grid.arrange(grid_arrange, legend, ncol = 2, widths = c(7, 1))
+@
+% \includegraphics[width=1\textwidth]{"Density_plot.png"}
+\label{fig:sim_plot}
+\caption{Simulations of susceptible and infected \textit{Daphnia} in the mixed species parasitized dynamics.
+The each of the four blocks contains a simulation band and real data curves.
+The real data curves are obtained from each bucket in the mixed species parasitized dynamics with various color.
+And simulation band is obtained by the estimated parameter swarm based on the SIRPF2 model.
+The x-axis in in unit of day while the y-axis is in unit of square root of the types density.}
+\end{figure}
+
+Based on likelihood ratio tests and AIC comparisons, we identified the optimal parameter swarm that most accurately fits the data.
+To validate the performance of the estimated parameters and the robustness of the SIRPF2 model, we conducted a simulation study.
+Specifically, we generated 10000 simulation runs for each unit within the model framework and drop the outliers, as depicted in Figure 3.
+As we used shared parameters for each buckets, we combine the simulation bands and real data curves in one block to explore the common pattern of the dynamics.
+These simulations allow us to assess the variability and reliability of the model outputs under the influence of the optimized parameters.
+At the same time, they provide a robust framework to test whether the model can consistently replicate the observed trends across multiple stochastic realizations.
+
+The simulation bands presented in Figure 3 closely align with the empirical data curves from the actual experiment, underscoring the model's capability to replicate the observed dynamics.
+This alignment is not merely about capturing general trends but also about reflecting the variability inherent in the real-world data.
+The close correspondence between the simulation bands and the actual data curves suggests that the observations are not only plausible but also statistically probable realizations of the fitted model.
+
+Such findings are indicative of a goodness of fit in terms of visualization, affirming that the all-shared parameter swarms effectively encapsulate the essential dynamics of the process under study.
+The convergence of the simulated data with the empirical data provides strong evidence that the SIRPF2 model, with the derived parameters, accurately represents the underlying ecological dynamic processes.
+This combination of mechanistic modeling and empirical data validation through simulation offers robust support for the model's predictive accuracy and reliability in ecological researches.
 
@@
 
-The SIRPF model is a powerful tool for analyzing ecological data, as it can be used to estimate parameters and predict future population dynamics.
+The SIRPF model extends the traditional SIR framework by including food and parasites, which is crucial in ecological models where these factors significantly influence host dynamics.
+This approach reflects the understanding that the dynamics of pathogens, including coinfections and community composition, are significantly influenced by ecological factors such as food availability and parasitic interactions.
+The complexity of these interactions is highlighted by studies emphasizing the role of community ecology in infectious disease research and the phylogenetic structure of host-parasite relationships.
+The SIRPF model's inclusion of these aspects offers a more comprehensive understanding of ecological dynamics, particularly in systems where traditional SIR models may fall short.
+As a result, the SIRPF model is a powerful tool for analyzing ecological data, as it can be used to estimate parameters and predict future population dynamics.
 The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of \textit{Daphnia}.
@@
 The model's flexibility allows for the incorporation of a variety of factors, including food resources and parasitism, that can be used to assess the impact of these factors on the population dynamics of \textit{Daphnia}.\\
-
-The SIRPF model extends the traditional SIR framework by including food and parasites, which is crucial in ecological models where these factors significantly influence host dynamics.
-This approach reflects the understanding that the dynamics of pathogens, including coinfections and community composition, are significantly influenced by ecological factors such as food availability and parasitic interactions.
-The complexity of these interactions is highlighted by studies emphasizing the role of community ecology in infectious disease research and the phylogenetic structure of host-parasite relationships.
-The SIRPF model's inclusion of these aspects offers a more comprehensive understanding of ecological dynamics, particularly in systems where traditional SIR models may fall short.\\
 
@@
 
+These findings also underscore the importance of parameter sharing in ecological modeling when multiple species or units are involved.
+Specifically, the superior AIC values obtained for shared-parameter models suggest that the interactions between \textit{Daphnia} populations, their invasive counterparts, and parasites are more homogeneously influenced by shared ecological factors than by unit-specific ones.
+The results further reinforce the utility of the AIC in guiding the selection of models that not only provide a good fit to the data but also maintain a degree of simplicity, avoiding overfitting.
+In the meantime, hypothesis test via likelihood ratio test is a vital method for parameter selection.
+Incorporating with AIC, these evaluation methods helps making decision of mechanistic models.\\
+
 \bibliography{reference}
